---
title: "Supplement codes 1: Dental-related consultations in general practice in England: a population-based retrospective study of national trends and inequalities"
author: "CG"
output: html_document
---

# restart the working environment 
```{r}
rm(list = ls())
```


# library
```{r setup, include=FALSE}
library(readxl)
library(writexl)
library(tidyverse)
library(tidyr)
library(popEpi)
library(epitools)

#install.packages(c("sf","ggplot2","dplyr","viridis", "rgeos", "rgdal", "maptool"))
library(sf)
library(ggplot2)
library(dplyr)
library(viridis)
library(epiR)
library(PHEindicatormethods)
library(broom)
library(dplyr)
library(purrr)
library(tidyverse)
library(lubridate)
library(MASS)
library(sandwich)
library(lmtest)
library(segmented)
# install.packages("forecast")
library(forecast)
library(MASS)     # glm.nb
library(splines)  # ns()

library(dplyr)
library(ggplot2)
library(patchwork)
library(grid)   # for arrow()
#library(rgeos)
#library(rgdal)
#library(maptools)
library(forcats)
library(cowplot)


# frequent used path
tables_figures <- "Analysis_results_CG/tables_figures"
processed_data <- "Analysis_results_CG/processed_data"
raw_data <- "Analysis_results_CG/raw_data"
final_curated_data <- "Analysis_results_CG/processed_data/final_curated_data"

code_lists <- "code_list"


# 2. define study period
study_start <- as.Date("1998-01-01")

study_end <- as.Date("2023-12-31")
 

```


# Section 2: overall ir
## Step 1: create overall summary table (dcat and age will be bind later)
```{r}
# 1. load the ir overall function 
source("IR_summary_table_v3.R")
# 2. calculate overall IR without any grouping 
ir_summary_overall <- make_ir_table(
  data  = sample_final,
  py    = "py_ind_total",
  cases = "dental_visit_count"
)

# 3. calcualte overall IR for each groups 
group_vars <- c("region", "gender","final_ethnicity2", "final_ethnicity_groups2","imd_quintile")

ir_summary_vars <- dplyr::tibble()
for (g in group_vars) {
  tmp <- make_ir_table(
    data  = sample_final,
    py    = "py_ind_total",
    cases = "dental_visit_count",
    var   = g
  )
  ir_summary_vars <- dplyr::bind_rows(ir_summary_vars, tmp)  # accumulate here
}



```

## Step 2: confidence interval for all ir (dcat and age will be bind later)
```{r}

ir_summary_table <- ir_summary_table %>%
  rowwise() %>%
  mutate(
    # Run epi.conf() on each row and scale to 1000 person-years
    tmp = list(epi.conf(cbind(total_cases, total_person_year),
                        ctype = "inc.rate", method = "exact") * 1000),
    lower95_1000  = tmp$lower,
    upper95_1000  = tmp$upper
  ) %>%
  select(-tmp) %>%
  ungroup()

```
# Section 3: annual incidence 

## Step 1: overall annual incidence 
```{r}
#

source("IR_summary_table_v3.R")
# All visits per year (1998–2023)

annual_cases <- dental_records_long_final %>%
  filter(between(obsyear, 1998, 2023)) %>%
  count(year = obsyear, name = "cases")


# calculate py by year 
annual_py <- annual_person_time(
  data  = sample_final,
  id    = patid,
  start = fu_start_date,
  end   = censor_date,
  years = 1998:2023,
  aggregate = TRUE
)
# Annual summary table

annual_summary_table <- annual_py %>% left_join( annual_cases, by = "year")


# calculate ir1000
annual_summary_table$ir1000 <- (annual_summary_table$cases/annual_summary_table$total_person_years)*1000

# plot overall figure 
source("DEEP_lineplots_annual_ir_v1.R")

lp_annual_overall  <-make_lineplot(
  data    = annual_summary_table,
  x       = factor(year),
  y       = ir1000,
  title   = "Annual incidence rate",
  xlab    = "Calendar year",
  ylab    = "Incidence rate per 1,000 PY",
  outfile = paste0(tables_figures, "/objective 1/figure_1a_annual_ir_lineplot.pdf")
)


```


## Step 2: annual IR by 1 subgroup

### Step 2.1: gender 
```{r}
# 1. count the event number 
annual_cases_gender <- annual_cases_subgroups(
  events_df  = dental_records_long_final,
  event_year = obsyear,
  years      = 1998:2023,
  group_vars = "gender"
)

# 2. count the annual py 
annual_py_gender <- annual_py_subgroups(
  sample_df = sample_final,
  id        = patid,
  start     = fu_start_date,
  end       = censor_date,
  years     = 1998:2023,
  group_vars= "gender"
)

# 3. calculate IR
annual_ir_gender <- annual_cases_gender %>%
  left_join(annual_py_gender, by = c("gender","year"))

annual_ir_gender$ir1000 <- annual_ir_gender$cases/annual_ir_gender$total_person_years*1000


# 4. correct variable name for plot 
annual_ir_gender <- annual_ir_gender %>%
  mutate(gender = factor(gender, levels = rev(sort(unique(gender))))) %>%
  rename(Gender = gender)%>%
  rename(Sex = Gender)


# 5. plot annual gender ir
make_lineplot(
  data    = annual_ir_gender,
  x       = factor(year),
  y       = ir1000,
  color   = Sex,
  title   = "Annual incidence rate by gender",
  xlab    = "Calendar year",
  ylab    = "Incidence rate per 1,000 PY",
  outfile = paste0(tables_figures, "/objective 1/figure_1a_annual_ir_gender_lineplot.pdf")
)

```

### Step 2.2: ethnicity
```{r}
# 1. count the event number 
annual_cases_6eth <- annual_cases_subgroups(
  events_df  = dental_records_long_final,
  event_year = obsyear,
  years      = 1998:2023,
  group_vars = "final_ethnicity_groups"
)

# 2. count the annual py 
annual_py_6eth <- annual_py_subgroups(
  sample_df = sample_final,
  id        = patid,
  start     = fu_start_date,
  end       = censor_date,
  years     = 1998:2023,
  group_vars= "final_ethnicity_groups"
)

# 3. calculate IR
annual_ir_6eth <- annual_cases_6eth %>%
  left_join(annual_py_6eth, by = c("final_ethnicity_groups","year"))

annual_ir_6eth$ir1000 <- annual_ir_6eth$cases/annual_ir_6eth$total_person_years*1000


# 4.correct plot name 
annual_ir_6eth <- annual_ir_6eth %>%
  mutate(final_ethnicity_groups = factor(final_ethnicity_groups, levels = rev(sort(unique(final_ethnicity_groups))))) %>%
  rename(Ethnicity = final_ethnicity_groups)

# 5. plot annual 6eth ir
make_lineplot(
  data    = annual_ir_6eth,
  x       = factor(year),
  y       = ir1000,
  color   = Ethnicity,
  title   = "Annual incidence rate by 6eth",
  xlab    = "Calendar year",
  ylab    = "Incidence rate per 1,000 PY",
  outfile = paste0(tables_figures, "/objective 1/figure_1a_annual_ir_6eth_lineplot.pdf")
)

```


### Step 2.3: region

```{r}
# 1. count the event number 
annual_cases_region <- annual_cases_subgroups(
  events_df  = dental_records_long_final,
  event_year = obsyear,
  years      = 1998:2023,
  group_vars = "region"
)

# 2. count the annual py 
annual_py_region <- annual_py_subgroups(
  sample_df = sample_final,
  id        = patid,
  start     = fu_start_date,
  end       = censor_date,
  years     = 1998:2023,
  group_vars= "region"
)

# 3. calculate IR
annual_ir_region <- annual_cases_region %>%
  left_join(annual_py_region, by = c("region","year"))

annual_ir_region$ir1000 <- annual_ir_region$cases/annual_ir_region$total_person_years*1000

# 4.correct plot name 
annual_ir_region <- annual_ir_region %>%
  mutate(region = factor(region, levels = rev(sort(unique(region))))) %>%
  rename(Region = region)


# 5. plot annual region ir
lp_annual_ir_region <- make_lineplot(
  data    = annual_ir_region,
  x       = factor(year),
  y       = ir1000,
  color   = Region,
  title   = "Annual incidence rate by region",
  xlab    = "Calendar year",
  ylab    = "Incidence rate per 1,000 PY",
  outfile = paste0(tables_figures, "/objective 1/figure_1a_annual_ir_region_lineplot.pdf")
)

```

### Step 2.4: imd
```{r}
# 1. count the event number 
annual_cases_imd <- annual_cases_subgroups(
  events_df  = dental_records_long_final,
  event_year = obsyear,
  years      = 1998:2023,
  group_vars = "imd_quintile"
)

# 2. count the annual py 
annual_py_imd <- annual_py_subgroups(
  sample_df = sample_final,
  id        = patid,
  start     = fu_start_date,
  end       = censor_date,
  years     = 1998:2023,
  group_vars= "imd_quintile"
)

# 3. calculate IR
annual_ir_imd <- annual_cases_imd %>%
  left_join(annual_py_imd, by = c("imd_quintile","year"))

annual_ir_imd$ir1000 <- annual_ir_imd$cases/annual_ir_imd$total_person_years*1000


# 4. correct variable name for plot 
library(dplyr)

annual_ir_imd <- annual_ir_imd %>%
  mutate(
    # Recode only 1 and 5, keep others as they are
    imd_quintile = case_when(
      imd_quintile == 1 ~ "1 (Least Deprived)",
      imd_quintile == 5 ~ "5 (Most Deprived)",
      TRUE              ~ as.character(imd_quintile)   # keep other values
    ),
    # Turn into factor with reverse alphabetical order
    imd_quintile = factor(imd_quintile,
                          levels = rev(sort(unique(imd_quintile))))
  ) %>%
  rename(IMD = imd_quintile)


# 5. plot annual imd ir
lp_annual_ir_imd <- make_lineplot(
  data    = annual_ir_imd,
  x       = factor(year),
  y       = ir1000,
  color   = IMD,
  title   = "Annual incidence rate by imd",
  xlab    = "Calendar year",
  ylab    = "Incidence rate per 1,000 PY",
  outfile = paste0(tables_figures, "/objective 1/figure_1a_annual_ir_imd_lineplot.pdf")
)

```

### Step 2.5: age 
```{r}

# 1. count the event number 
annual_cases_age <- annual_cases_subgroups(
  events_df  = dental_records_long_final,
  event_year = obsyear,
  years      = 1998:2023,
  group_vars = "age_band"
)

# 2. count the annual py 
annual_py_age <- annual_py_by_age(
  sample_df = sample_final,
  id        = patid,
  start     = fu_start_year,
  end       = censor_year,
  years     = 1998:2023,
  yob       = yob
)

 


# 3. calculate IR
annual_ir_age <- annual_cases_age %>%
  left_join(annual_py_age, by = c("age_band","year"))

annual_ir_age$ir1000 <- annual_ir_age$cases/annual_ir_age$total_person_years*1000

# 4.correct plot name 
annual_ir_age <- annual_ir_age %>%
  mutate(`Age group` = factor(age_band, levels = c("Older adults (60+)", "Adults (18–59)", "Children (3–17)")))



# 5. plot annual age ir
make_lineplot(
  data    = annual_ir_age,
  x       = factor(year),
  y       = ir1000,
  color   = `Age group`,
  title   = "Annual incidence rate by age",
  xlab    = "Calendar year",
  ylab    = "Incidence rate per 1,000 PY",
  outfile = paste0(tables_figures, "/objective 1/figure_1a_annual_ir_age_lineplot.pdf")
)

# confidence interval
annual_ir_age <- annual_ir_age %>%
  rowwise() %>%
  mutate(
    # Run epi.conf() on each row and scale to 1000 person-years
    tmp = list(epi.conf(cbind(cases, total_person_years),
                        ctype = "inc.rate", method = "exact") * 1000),
    lower95_1000  = tmp$lower,
    upper95_1000  = tmp$upper
  ) %>%
  select(-tmp) %>%
  ungroup()
```
## Step 4: 2x2 ir 
### Step 4.1:  ethnicity xregion
```{r}
# 1. create summary table
ir_eth_region <- make_ir_table_combo(
  data  = sample_final,
  py    = "py_ind_total",
  cases = "dental_visit_count",
  vars  = c("final_ethnicity_groups","region")
)


# 2.calculate IR

ir_eth_region$ir1000 <- ir_eth_region$total_cases/ir_eth_region$total_person_year*1000


# 3. correct variable name for plot 
ir_eth_region <- ir_eth_region %>% 
  mutate(Ethnicity = factor(final_ethnicity_groups, levels = rev(sort(unique(final_ethnicity_groups))))) %>% 
  mutate(Region = factor(region, levels = rev(sort(unique(region)))))


# 4. plot annual 6eth_rgn ir
make_heatmap(
  data    = ir_eth_region,
  x       = Ethnicity,
  y       = Region,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by ethnicity and region",
  xlab        = "Ethnicity",
  ylab        = "Region",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_eth_rgn.pdf"),
  width       = 12,
  height      = 6,
  dpi         = 600
)




```


### Step 4.4:  ethnicity ximd
```{r}
# 1. create summary table
ir_eth_imd <- make_ir_table_combo(
  data  = sample_final,
  py    = "py_ind_total",
  cases = "dental_visit_count",
  vars  = c("final_ethnicity_groups","imd_quintile")
)


# 2.calculate IR

ir_eth_imd$ir1000 <- ir_eth_imd$total_cases/ir_eth_imd$total_person_year*1000


# 3. correct variable name for plot 
ir_eth_imd <- ir_eth_imd %>% 
  mutate(Ethnicity = factor(final_ethnicity_groups, levels = rev(sort(unique(final_ethnicity_groups))))) %>% mutate(
    # Recode only 1 and 5, keep others as they are
    IMD = case_when(
      imd_quintile == 1 ~ "1 (Least Deprived)",
      imd_quintile == 5 ~ "5 (Most Deprived)",
      TRUE              ~ as.character(imd_quintile)   # keep other values
    ),
    # Turn into factor with reverse alphabetical order
    imd_quintile = factor(imd_quintile,
                          levels = rev(sort(unique(imd_quintile)))))

# 4. plot annual 6eth_rgn ir
make_heatmap(
  data    = ir_eth_imd,
  x       = Ethnicity,
  y       = IMD,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by ethnicity and IMD",
  xlab        = "Ethnicity",
  ylab        = "IMD",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_eth_imd.pdf"),
  width       = 12,
  height      = 6,
  dpi         = 600
)




```
### Step 4.5: all combination
```{r}
#############################################
## Remaining 2x2 IR heatmaps (NO age, NO dcat)
## IMD x Sex; IMD x Region; Ethnicity x Sex; Region x Sex
#############################################

## 4.X: IMD x Sex
ir_imd_sex <- make_ir_table_combo(
  data  = sample_final,
  py    = "py_ind_total",
  cases = "dental_visit_count",
  vars  = c("imd_quintile","gender")
)

ir_imd_sex$ir1000 <- ir_imd_sex$total_cases / ir_imd_sex$total_person_year * 1000

ir_imd_sex <- ir_imd_sex %>%
  mutate(
    Sex = factor(gender, levels = rev(sort(unique(gender)))),
    IMD = case_when(
      imd_quintile == 1 ~ "1 (Least Deprived)",
      imd_quintile == 5 ~ "5 (Most Deprived)",
      TRUE              ~ as.character(imd_quintile)
    ),
    imd_quintile = factor(imd_quintile, levels = rev(sort(unique(imd_quintile))))
  )

make_heatmap(
  data    = ir_imd_sex,
  x       = IMD,
  y       = Sex,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by sex and IMD",
  xlab        = "IMD",
  ylab        = "Sex",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_imd_sex.pdf"),
  width       = 12,
  height      = 6,
  dpi         = 600
)


## 4.X: IMD x Region
ir_imd_region <- make_ir_table_combo(
  data  = sample_final,
  py    = "py_ind_total",
  cases = "dental_visit_count",
  vars  = c("imd_quintile","region")
)

ir_imd_region$ir1000 <- ir_imd_region$total_cases / ir_imd_region$total_person_year * 1000

ir_imd_region <- ir_imd_region %>%
  mutate(
    Region = factor(region, levels = rev(sort(unique(region)))),
    IMD = case_when(
      imd_quintile == 1 ~ "1 (Least Deprived)",
      imd_quintile == 5 ~ "5 (Most Deprived)",
      TRUE              ~ as.character(imd_quintile)
    ),
    imd_quintile = factor(imd_quintile, levels = rev(sort(unique(imd_quintile))))
  )

make_heatmap(
  data    = ir_imd_region,
  x       = IMD,
  y       = Region,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by region and IMD",
  xlab        = "IMD",
  ylab        = "Region",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_imd_region.pdf"),
  width       = 12,
  height      = 6,
  dpi         = 600
)


## 4.X: Ethnicity x Sex
ir_eth_sex <- make_ir_table_combo(
  data  = sample_final,
  py    = "py_ind_total",
  cases = "dental_visit_count",
  vars  = c("final_ethnicity_groups","gender")
)

ir_eth_sex$ir1000 <- ir_eth_sex$total_cases / ir_eth_sex$total_person_year * 1000

ir_eth_sex <- ir_eth_sex %>%
  mutate(
    Ethnicity = factor(final_ethnicity_groups,
                       levels = rev(sort(unique(final_ethnicity_groups)))),
    Sex = factor(gender, levels = rev(sort(unique(gender))))
  )

make_heatmap(
  data    = ir_eth_sex,
  x       = Ethnicity,
  y       = Sex,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by ethnicity and sex",
  xlab        = "Ethnicity",
  ylab        = "Sex",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_eth_sex.pdf"),
  width       = 12,
  height      = 6,
  dpi         = 600
)


## 4.X: Region x Sex
ir_region_sex <- make_ir_table_combo(
  data  = sample_final,
  py    = "py_ind_total",
  cases = "dental_visit_count",
  vars  = c("region","gender")
)

ir_region_sex$ir1000 <- ir_region_sex$total_cases / ir_region_sex$total_person_year * 1000

ir_region_sex <- ir_region_sex %>%
  mutate(
    Region = factor(region, levels = rev(sort(unique(region)))),
    Sex = factor(gender, levels = rev(sort(unique(gender))))
  )

make_heatmap(
  data    = ir_region_sex,
  x       = Region,
  y       = Sex,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by region and sex",
  xlab        = "Region",
  ylab        = "Sex",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_region_sex.pdf"),
  width       = 12,
  height      = 6,
  dpi         = 600
)

```


## Step 5: dental category IR
### Step 5.2: overall dent cat py
```{r}
# 1. load the ir overall function 
source("IR_summary_table_v3.R")


# 2. calculate overall IR for each dent cat

ir_summary_dcat <- list()  # initialize an empty list to store results

for (i in dent_cat_list) {
  ir_summary_dcat[[i]] <- make_ir_table(
    data  = sample_final,
    py    = "py_ind_total",
    cases = paste0("dcount_",i)  # assuming your columns are like A_count, B_count, etc.
  )
}


ir_summary_dcat <- bind_rows(ir_summary_dcat, .id = "level") %>%
  relocate(variable, .before = 1) %>%     # optional: put it first
  mutate(variable= "Dental problem category") %>%
  relocate(variable, .before = 1)   # make it the first column
```

### Step 5.4: overall ir and ci 
```{r}
# 3. calculate ir and ci 

ir_summary_dcat <- ir_summary_dcat %>%
  rowwise() %>%
  mutate(
    # Run epi.conf() on each row and scale to 1000 person-years
    tmp = list(epi.conf(cbind(total_cases, total_person_year),
                        ctype = "inc.rate", method = "exact") * 1000),
    ir1000   = (total_cases / total_person_year) * 1000,
    lower95_1000  = tmp$lower,
    upper95_1000  = tmp$upper
  ) %>%
  select(-tmp) %>%
  ungroup()

# 4. combine with ir summary table 

ir_summary_table <- bind_rows(ir_summary_table, ir_summary_dcat
                              )
# 5. explort the ir_summary table
write_xlsx(ir_summary_table, paste0(tables_figures, "/objective 1/ir_summary_table.xlsx"))

```

### Step 5.5: dcat x region

```{r}
# Create empty lists to store results
ir_summary_dcat_region <- list()

# Names to remove
remove_items <- c("Onward referral",
                           "Examination", 
                           "Non-specific dental disease")

# Filter the list
dent_cat_list <- dent_cat_list[!sapply(dent_cat_list, function(x) x %in% remove_items)]

# Loop through each dental category
for (i in c(paste0("dcount_", dent_cat_list))) {

  # ---- region-specific IR ----
  ir_dcat_region <- make_ir_table(
    data  = sample_final,
    py    = "py_ind_total",
    cases = i,
    var   = "region"
  )

  # ---- Store results in lists ----
  ir_summary_dcat_region[[i]] <- ir_dcat_region
}

# Combine results across categories
ir_summary_dcat_region <- bind_rows(ir_summary_dcat_region, .id = "variable") %>%
  relocate(variable, .before = 1) 

  # Add IR per 1000 PY + exact Poisson CI for both tables
ir_summary_dcat_region <- ir_summary_dcat_region %>%
    rowwise() %>%
    mutate(
      ir1000 = (total_cases / total_person_year) * 1000,
      tmp = list(
        epi.conf(cbind(total_cases, total_person_year),
                 ctype = "inc.rate", method = "exact") * 1000
      ),
      lower95 = tmp$lower,
      upper95 = tmp$upper
    ) %>%
    select(-tmp) %>%
    ungroup() 


####################
# 3. correct variable name for plot 
# 3) Clean the category label from "dcount_*" and make a factor (ordered) 
ir_summary_dcat_region <- ir_summary_dcat_region %>% 
  mutate( `Dental problem categories` = gsub("^dcount_", "", variable))%>%
  mutate( `Dental problem categories` = factor( `Dental problem categories`, levels = rev(sort(unique(`Dental problem categories`))) ) )%>%
  mutate(Region = factor(level, levels = rev(sort(unique(level)))))

# 4. plot annual 6eth_rgn ir
make_heatmap(
  data    = ir_summary_dcat_region,
  x       = Region,
  y       =  `Dental problem categories`,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by region and dental problem categories",
  xlab        = "Region",
  ylab        = "Dental problem categories",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_dcat_region.pdf"),
  width       = 12,
  height      = 6,
  dpi         = 600
)

```

### Step 5.6: dcat x imd

```{r}
# Create empty lists to store results
ir_summary_dcat_imd <- list()



# Loop through each dental category
for (i in c(paste0("dcount_", dent_cat_list))) {

  # ---- imd-specific IR ----
  ir_dcat_imd <- make_ir_table(
    data  = sample_final,
    py    = "py_ind_total",
    cases = i,
    var   = "imd_quintile"
  )

  # ---- Store results in lists ----
  ir_summary_dcat_imd[[i]] <- ir_dcat_imd
}

# Combine results across categories
ir_summary_dcat_imd <- bind_rows(ir_summary_dcat_imd, .id = "variable") %>%
  relocate(variable, .before = 1) 

  # Add IR per 1000 PY + exact Poisson CI for both tables
ir_summary_dcat_imd <- ir_summary_dcat_imd %>%
    rowwise() %>%
    mutate(
      ir1000 = (total_cases / total_person_year) * 1000,
      tmp = list(
        epi.conf(cbind(total_cases, total_person_year),
                 ctype = "inc.rate", method = "exact") * 1000
      ),
      lower95 = tmp$lower,
      upper95 = tmp$upper
    ) %>%
    select(-tmp) %>%
    ungroup() 


####################
# 3. correct variable name for plot 
# 3) Clean the category label from "dcount_*" and make a factor (ordered) 
ir_summary_dcat_imd <- ir_summary_dcat_imd %>% 
  mutate( `Dental problem categories` = gsub("^dcount_", "", variable))%>%
  mutate( `Dental problem categories` = factor( `Dental problem categories`, levels = rev(sort(unique(`Dental problem categories`))) ) )%>%
  mutate(imd = factor(level, levels = rev(sort(unique(level)))))

# 4. plot annual 6eth_rgn ir
make_heatmap(
  data    = ir_summary_dcat_imd,
  x       = imd,
  y       =  `Dental problem categories`,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by IMD and dental problem categories",
  xlab        = "IMD",
  ylab        = "Dental problem categories",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_dcat_imd.pdf"),
  width       = 12,
  height      = 6,
  dpi         = 600
)

```

### Step 5.7: dcat x ethnicity

```{r}
# Create empty lists to store results
ir_summary_dcat_ethnicity <- list()



# Loop through each dental category
for (i in c(paste0("dcount_", dent_cat_list))) {

  # ---- ethnicity-specific IR ----
  ir_dcat_ethnicity <- make_ir_table(
    data  = sample_final,
    py    = "py_ind_total",
    cases = i,
    var   = "final_ethnicity_groups"
  )

  # ---- Store results in lists ----
  ir_summary_dcat_ethnicity[[i]] <- ir_dcat_ethnicity
}

# Combine results across categories
ir_summary_dcat_ethnicity <- bind_rows(ir_summary_dcat_ethnicity, .id = "variable") %>%
  relocate(variable, .before = 1) 

  # Add IR per 1000 PY + exact Poisson CI for both tables
ir_summary_dcat_ethnicity <- ir_summary_dcat_ethnicity %>%
    rowwise() %>%
    mutate(
      ir1000 = (total_cases / total_person_year) * 1000,
      tmp = list(
        epi.conf(cbind(total_cases, total_person_year),
                 ctype = "inc.rate", method = "exact") * 1000
      ),
      lower95 = tmp$lower,
      upper95 = tmp$upper
    ) %>%
    select(-tmp) %>%
    ungroup() 


####################
# 3. correct variable name for plot 
# 3) Clean the category label from "dcount_*" and make a factor (ordered) 
ir_summary_dcat_ethnicity <- ir_summary_dcat_ethnicity %>% 
  mutate( `Dental problem categories` = gsub("^dcount_", "", variable))%>%
  mutate( `Dental problem categories` = factor( `Dental problem categories`, levels = rev(sort(unique(`Dental problem categories`))) ) )%>%
  mutate(Ethnicity = factor(level, levels = rev(sort(unique(level)))))


# 4. plot annual 6eth_rgn ir
make_heatmap(
  data    = ir_summary_dcat_ethnicity,
  x       = Ethnicity,
  y       =  `Dental problem categories`,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by ethnicity and dental problem categories",
  xlab        = "Ethnicity",
  ylab        = "Dental problem categories",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_dcat_ethnicity.pdf"),
  width       = 12,
  height      = 6,
  dpi         = 600
)

```

### Step 5.7: dcat x sex
```{r}
# Create empty lists to store results
ir_summary_dcat_sex <- list()



# Loop through each dental category
for (i in c(paste0("dcount_", dent_cat_list))) {

  # ---- sex-specific IR ----
  ir_dcat_sex <- make_ir_table(
    data  = sample_final,
    py    = "py_ind_total",
    cases = i,
    var   = "gender"
  )

  # ---- Store results in lists ----
  ir_summary_dcat_sex[[i]] <- ir_dcat_sex
}

# Combine results across categories
ir_summary_dcat_sex <- bind_rows(ir_summary_dcat_sex, .id = "variable") %>%
  relocate(variable, .before = 1) 

  # Add IR per 1000 PY + exact Poisson CI for both tables
ir_summary_dcat_sex <- ir_summary_dcat_sex %>%
    rowwise() %>%
    mutate(
      ir1000 = (total_cases / total_person_year) * 1000,
      tmp = list(
        epi.conf(cbind(total_cases, total_person_year),
                 ctype = "inc.rate", method = "exact") * 1000
      ),
      lower95 = tmp$lower,
      upper95 = tmp$upper
    ) %>%
    select(-tmp) %>%
    ungroup() 


####################
# 3. correct variable name for plot 
# 3) Clean the category label from "dcount_*" and make a factor (ordered) 
ir_summary_dcat_sex <- ir_summary_dcat_sex %>% 
  mutate( `Dental problem categories` = gsub("^dcount_", "", variable))%>%
  mutate( `Dental problem categories` = factor( `Dental problem categories`, levels = rev(sort(unique(`Dental problem categories`))) ) )%>%
  mutate(sex = factor(level, levels = rev(sort(unique(level)))))


# 4. plot annual 6eth_rgn ir
make_heatmap(
  data    = ir_summary_dcat_sex,
  x       = sex,
  y       =  `Dental problem categories`,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by sex and dental problem categories",
  xlab        = "Sex",
  ylab        = "Dental problem categories",
  
  
  
  
  
  
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_dcat_sex.pdf"),
  width       = 12,
  height      = 6,
  dpi         = 600
)

```

### Step 5.4: annual dcat ir
```{r}
source("IR_summary_table_v3.R")


# All visits per year (1998–2023)
annual_cases_dcat <- annual_cases_subgroups(
  events_df  = dental_records_long_final,
  event_year = obsyear,
  years      = 1998:2023,
  group_vars = c("dental_category2")
)

# annual py stay the same as before so no need to calculate again.
#annual_py <- annual_person_time(
#  data  = sample_final,
#  id    = patid,
#  start = fu_start_date,
#  end   = censor_date,
#  years = 1998:2023,
#  aggregate = TRUE
#)
# Annual summary table

annual_summary_dcat <- annual_cases_dcat %>% left_join(annual_py, by = "year")


# calculate ir1000
annual_summary_dcat$ir1000 <- (annual_summary_dcat$cases/annual_summary_dcat$total_person_years)*1000


# format dental category 
annual_summary_dcat <- annual_summary_dcat %>% 
  mutate(`Dental problem categories` = factor(`dental_category2`, levels = rev(sort(unique(`dental_category2`)))))
# plot overall figure 
source("DEEP_lineplots_annual_ir_v1.R")

lp_annual_dcat <- make_lineplot(
  data    = annual_summary_dcat,
  x       = year,
  y       = ir1000,
  facet =   `Dental problem categories`,
  title   = "Annual incidence rate",
  xlab    = "Calendar year",
  ylab    = "Incidence rate per 1,000 PY",
  outfile = paste0(tables_figures, "/objective 1/figure_1a_dcat_annual_ir_lineplot.pdf")
)
```


# Section 4: standardised rate to 2013 ESP and ESP age band ir
## Step 1: by age
### Step 1.1: create more detailed age band based on 2013 ESP
```{r}

# 1. define age group based on dob_age

# library(epiR)   # for epi.conf()
# library(writexl) # for write_xlsx()

dental_records_long_final <- dental_records_long_final %>%
  mutate(
    age_band2 = case_when(
      dob_age < 5   ~ "0-4",
      dob_age >= 5  & dob_age < 10  ~ "5-9",
      dob_age >= 10 & dob_age < 15  ~ "10-14",
      dob_age >= 15 & dob_age < 20  ~ "15-19",
      dob_age >= 20 & dob_age < 25  ~ "20-24",
      dob_age >= 25 & dob_age < 30  ~ "25-29",
      dob_age >= 30 & dob_age < 35  ~ "30-34",
      dob_age >= 35 & dob_age < 40  ~ "35-39",
      dob_age >= 40 & dob_age < 45  ~ "40-44",
      dob_age >= 45 & dob_age < 50  ~ "45-49",
      dob_age >= 50 & dob_age < 55  ~ "50-54",
      dob_age >= 55 & dob_age < 60  ~ "55-59",
      dob_age >= 60 & dob_age < 65  ~ "60-64",
      dob_age >= 65 & dob_age < 70  ~ "65-69",
      dob_age >= 70 & dob_age < 75  ~ "70-74",
      dob_age >= 75 & dob_age < 80  ~ "75-79",
      dob_age >= 80 & dob_age < 85  ~ "80-84",
      dob_age >= 85 & dob_age < 90  ~ "85-89",
      dob_age >= 90                 ~ "90+",
      TRUE ~ NA_character_   # fallback for negatives/missing
    ),
    age_band2 = factor(
      age_band2,
      levels = c("0-4","5-9","10-14","15-19","20-24","25-29","30-34",
                 "35-39","40-44","45-49","50-54","55-59","60-64","65-69",
                 "70-74","75-79","80-84","85-89","90+")
    )
  )

table(dental_records_long_final$age_band2, useNA = "ifany")

# 2. create new data frame document number of cases
age_band2_cases <- dental_records_long_final %>%
  group_by(age_band2) %>%
  summarise(
    total_cases = n(),       # number of records in each age band
    .groups = "drop"
  ) %>%
  arrange(age_band2)
```



### Step 1.2: get py for them

```{r}
# Add date cutpoints for each age boundary relative to DOB
age_band2_py_cut <- sample_final %>%
  mutate(
    dob_0  = dob,
    dob_5  = dob %m+% years(5),
    dob_10 = dob %m+% years(10),
    dob_15 = dob %m+% years(15),
    dob_20 = dob %m+% years(20),
    dob_25 = dob %m+% years(25),
    dob_30 = dob %m+% years(30),
    dob_35 = dob %m+% years(35),
    dob_40 = dob %m+% years(40),
    dob_45 = dob %m+% years(45),
    dob_50 = dob %m+% years(50),
    dob_55 = dob %m+% years(55),
    dob_60 = dob %m+% years(60),
    dob_65 = dob %m+% years(65),
    dob_70 = dob %m+% years(70),
    dob_75 = dob %m+% years(75),
    dob_80 = dob %m+% years(80),
    dob_85 = dob %m+% years(85),
    dob_90 = dob %m+% years(90)
  )

# 2) Recompute band DAYS exactly (your logic), filling NAs with 0
age_band2_py_cut <- age_band2_py_cut %>%
  mutate(
    py_0_4_days = case_when(
      dob_5  <= fu_start_date ~ 0L,
      dob_0  >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_5) - pmax(fu_start_date, dob_0))
    ),
    py_5_9_days = case_when(
      dob_10 <= fu_start_date ~ 0L,
      dob_5  >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_10) - pmax(fu_start_date, dob_5))
    ),
    py_10_14_days = case_when(
      dob_15 <= fu_start_date ~ 0L,
      dob_10 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_15) - pmax(fu_start_date, dob_10))
    ),
    py_15_19_days = case_when(
      dob_20 <= fu_start_date ~ 0L,
      dob_15 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_20) - pmax(fu_start_date, dob_15))
    ),
    py_20_24_days = case_when(
      dob_25 <= fu_start_date ~ 0L,
      dob_20 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_25) - pmax(fu_start_date, dob_20))
    ),
    py_25_29_days = case_when(
      dob_30 <= fu_start_date ~ 0L,
      dob_25 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_30) - pmax(fu_start_date, dob_25))
    ),
    py_30_34_days = case_when(
      dob_35 <= fu_start_date ~ 0L,
      dob_30 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_35) - pmax(fu_start_date, dob_30))
    ),
    py_35_39_days = case_when(
      dob_40 <= fu_start_date ~ 0L,
      dob_35 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_40) - pmax(fu_start_date, dob_35))
    ),
    py_40_44_days = case_when(
      dob_45 <= fu_start_date ~ 0L,
      dob_40 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_45) - pmax(fu_start_date, dob_40))
    ),
    py_45_49_days = case_when(
      dob_50 <= fu_start_date ~ 0L,
      dob_45 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_50) - pmax(fu_start_date, dob_45))
    ),
    py_50_54_days = case_when(
      dob_55 <= fu_start_date ~ 0L,
      dob_50 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_55) - pmax(fu_start_date, dob_50))
    ),
    py_55_59_days = case_when(
      dob_60 <= fu_start_date ~ 0L,
      dob_55 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_60) - pmax(fu_start_date, dob_55))
    ),
    py_60_64_days = case_when(
      dob_65 <= fu_start_date ~ 0L,
      dob_60 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_65) - pmax(fu_start_date, dob_60))
    ),
    py_65_69_days = case_when(
      dob_70 <= fu_start_date ~ 0L,
      dob_65 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_70) - pmax(fu_start_date, dob_65))
    ),
    py_70_74_days = case_when(
      dob_75 <= fu_start_date ~ 0L,
      dob_70 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_75) - pmax(fu_start_date, dob_70))
    ),
    py_75_79_days = case_when(
      dob_80 <= fu_start_date ~ 0L,
      dob_75 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_80) - pmax(fu_start_date, dob_75))
    ),
    py_80_84_days = case_when(
      dob_85 <= fu_start_date ~ 0L,
      dob_80 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_85) - pmax(fu_start_date, dob_80))
    ),
    py_85_89_days = case_when(
      dob_90 <= fu_start_date ~ 0L,
      dob_85 >= censor_date   ~ 0L,
      TRUE ~ as.integer(pmin(censor_date, dob_90) - pmax(fu_start_date, dob_85))
    ),
    py_90_plus_days = case_when(
      dob_90 >= censor_date ~ 0L,
      TRUE ~ as.integer(censor_date - pmax(fu_start_date, dob_90))
    ),

    # Convert days to person-years
    across(ends_with("_days"), ~ .x / 365.25, .names = "{.col}_py"),

    # Check total PY consistency
    py_sum_check = (
      py_0_4_days_py + py_5_9_days_py + py_10_14_days_py + py_15_19_days_py +
      py_20_24_days_py + py_25_29_days_py + py_30_34_days_py + py_35_39_days_py +
      py_40_44_days_py + py_45_49_days_py + py_50_54_days_py + py_55_59_days_py +
      py_60_64_days_py + py_65_69_days_py + py_70_74_days_py + py_75_79_days_py +
      py_80_84_days_py + py_85_89_days_py + py_90_plus_days_py
    ) - py_ind_total
  )

summary(age_band2_py_cut$py_sum_check)

# 3. calculate overall py

# Your age-band label order
band_levels <- c("0-4","5-9","10-14","15-19","20-24","25-29","30-34",
                 "35-39","40-44","45-49","50-54","55-59","60-64","65-69",
                 "70-74","75-79","80-84","85-89","90+")

# Sum per-band PY columns → long df with your band labels
age_band2_py <- age_band2_py_cut %>%
  dplyr::summarise(dplyr::across(stringr::str_subset(names(.), "^py_.*days_py$"),
                                ~ sum(.x, na.rm = TRUE))) %>%
  tidyr::pivot_longer(dplyr::everything(),
                      names_to = "band_col",
                      values_to = "total_person_year") %>%
  dplyr::mutate(
    age_band2 = band_col %>%
      stringr::str_remove("^py_") %>%
      stringr::str_remove("_days_py$") %>%
      stringr::str_replace_all("_", "-") %>%
      stringr::str_replace("-plus$", "+")
  ) %>%
  dplyr::select(age_band2, total_person_year) %>%
  dplyr::mutate(age_band2 = factor(age_band2, levels = band_levels)) %>%
  dplyr::arrange(age_band2)



# 4. ir_age_band2
ir_age_band2 <- age_band2_cases %>% left_join(age_band2_py, by = "age_band2")
```
### Step 1.3: general ir

```{r}

ir_age_band2 <- ir_age_band2 %>%
  rowwise() %>%
  mutate(
    # Run epi.conf() on each row and scale to 1000 person-years
    tmp = list(epi.conf(cbind(total_cases, total_person_year),
                        ctype = "inc.rate", method = "exact") * 1000),
    ir1000   = (total_cases / total_person_year) * 1000,
    lower95_1000  = tmp$lower,
    upper95_1000  = tmp$upper
  ) %>%
  dplyr::select(-tmp) %>%
  ungroup()

# 4. combine with ir summary table
ir_age_band2$variable <- "age_band2"


ir_summary_table <- bind_rows(ir_summary_table, ir_age_band2)

# 5. export the ir_summary table
write_xlsx(ir_summary_table, paste0(tables_figures, "/objective 1/ir_summary_table.xlsx"))
```

### Step 1.4: direct standardised IR
```{r}
# 1. bind esp2013 weight
ir_age_band2$esp2013 <- esp2013

# 2. calculate age specific IR
ir_age_band2$asir <- sum((ir_age_band2$ir1000/1000) * ir_age_band2$esp2013) / sum(ir_age_band2$esp2013)

# 3. get ci
asir <- calculate_dsr(
  ir_age_band2,
  total_cases,
  total_person_year,
  esp2013
)



```

### Step 1.4: indirect standardised
```{r}
# 2. calculate age specific indirect incidence
ir_age_band2$asi <- (ir_age_band2$ir1000/1000) * ir_age_band2$esp2013
```

### Step 1.5: non standardised age_band2 x dcat
```{r}
# 1. create summary table
age2_dcat_cases <- dental_records_long_final %>%
  group_by(age_band2, dental_category2) %>%
  summarise(total_cases = n(), .groups = "drop") %>%
  arrange(age_band2)

# 2. combine with py
ir_age2_dcat <- age2_dcat_cases %>%
  left_join(age_band2_py, by = "age_band2")

# 3. calculate IR
ir_age2_dcat$ir1000 <- ir_age2_dcat$total_cases / ir_age2_dcat$total_person_year * 1000

# 4. correct variable name for plot
ir_age2_dcat <- ir_age2_dcat %>%
  mutate(`Dental problem categories` = factor(dental_category2, levels = rev(sort(unique(dental_category2))))) %>%
  mutate(`Age group` = factor(age_band2))

# 5. plot annual 6eth_rgn ir
hm_ir_age2_dcat <- make_heatmap(
  data    = ir_age2_dcat,
  x       = `Dental problem categories`,
  y       = `Age group`,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by age group and dental problem categories",
  xlab        = "Dental problem categories",
  ylab        = "Age group",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_age2_dcat.pdf"),
  width       = 12,
  height      = 6,
  dpi         = 600
)
```

## Step 2: gender specific age standardised IR
### Step 2.1: get cases for age band2 by gender
```{r}
age_band2_cases2 <- dental_records_long_final %>%
  group_by(age_band2, gender) %>%
  summarise(total_cases = n(), .groups = "drop") %>%
  arrange(age_band2)
```

### Step 2.2: get py for age band2 by gender
```{r}
# 1) Named character vector: names are your bands, values are column names in the data
py_days <- c(
  "0-4"   = "py_0_4_days_py",
  "5-9"   = "py_5_9_days_py",
  "10-14" = "py_10_14_days_py",
  "15-19" = "py_15_19_days_py",
  "20-24" = "py_20_24_days_py",
  "25-29" = "py_25_29_days_py",
  "30-34" = "py_30_34_days_py",
  "35-39" = "py_35_39_days_py",
  "40-44" = "py_40_44_days_py",
  "45-49" = "py_45_49_days_py",
  "50-54" = "py_50_54_days_py",
  "55-59" = "py_55_59_days_py",
  "60-64" = "py_60_64_days_py",
  "65-69" = "py_65_69_days_py",
  "70-74" = "py_70_74_days_py",
  "75-79" = "py_75_79_days_py",
  "80-84" = "py_80_84_days_py",
  "85-89" = "py_85_89_days_py",
  "90+"   = "py_90_plus_days_py"
)

# 2) Loop one band at a time, pass the NAME of that PY column to make_ir_table()
age_band2_gender_py <- vector("list", length(py_days))
names(age_band2_gender_py) <- names(py_days)

for (band in names(py_days)) {
  colname <- py_days[[band]]

  tmp <- make_ir_table(
    data  = age_band2_py_cut,
    py    = colname,                 # pass the NAME, not the vector
    cases = "dental_visit_count",    # name as string
    var   = "gender"                 # name as string
  )

  age_band2_gender_py[[band]] <- dplyr::mutate(tmp, age_band2 = band)
}

ir_ab2_gender <- dplyr::bind_rows(age_band2_gender_py)
```

### Step 2.3: combine cases and py get ir
```{r}

# 1. clean py data frame
ir_ab2_gender <- ir_ab2_gender %>%
  dplyr::select(-total_cases, -variable) %>%   # drop columns
  dplyr::rename(gender = level)                # rename 'level' -> 'gender'

# 2. combine with case counts
ir_ab2_gender <- ir_ab2_gender %>%
  left_join(age_band2_cases2, by = c("age_band2", "gender"))

# 3. calculate ir per 1000
ir_ab2_gender <- ir_ab2_gender %>%
  rowwise() %>%
  dplyr::mutate(
    tmp = list(epi.conf(cbind(total_cases, total_person_year),
                        ctype = "inc.rate", method = "exact") * 1000),
    ir1000   = (total_cases / total_person_year) * 1000,
    lower95_1000  = tmp$lower,
    upper95_1000  = tmp$upper
  ) %>%
  dplyr::select(-tmp) %>%
  ungroup()

# 4. combine esp2013
ir_ab2_gender <- ir_ab2_gender %>%
  left_join(ir_age_band2 %>% dplyr::select(age_band2, esp2013) ,
            by = "age_band2")

# 5. combined asir (split esp weight by 2 for gender)
ir_ab2_gender$asir <- sum((ir_ab2_gender$ir1000/1000) * (ir_ab2_gender$esp2013/2)) / sum(ir_ab2_gender$esp2013/2)

# 6 get ci
ir_ab2_gender$esp2013_gender <- ir_ab2_gender$esp2013

asir_gender <- calculate_dsr(
  ir_ab2_gender,
  total_cases,
  total_person_year,
  esp2013_gender
)

##########################################
ir_ab2_male <- subset(ir_ab2_gender, gender == "Male")

ir_ab2_male$asir <- sum((ir_ab2_male$ir1000/1000) * (ir_ab2_male$esp2013/2)) / sum(ir_ab2_male$esp2013/2)

asir_male <- calculate_dsr(
  ir_ab2_male,
  total_cases,
  total_person_year,
  esp2013_gender
)

#########################
ir_ab2_female <- subset(ir_ab2_gender, gender == "Female")

ir_ab2_female$asir <- sum((ir_ab2_female$ir1000/1000) * (ir_ab2_female$esp2013/2)) / sum(ir_ab2_female$esp2013/2)

asir_female <- calculate_dsr(
  ir_ab2_female,
  total_cases,
  total_person_year,
  esp2013_gender
)


```

## Step 3: combine asir

```{r}
# 1. add row name
asir$variable <- "Age standardised IR"
asir_gender$variable <- "Age and gender standardised IR"
asir_female$variable <- "Gender-specific age standardised (Female) IR"
asir_male$variable <- "Gender-specific age standardised (Male) IR"

# 2. combine
asir_combined <- bind_rows(asir, asir_gender, asir_female, asir_male)
```

# Section 5: age 2x2 ir
### Step 2.1: age2 x dcat
```{r}
# 1. create summary table
age2_dcat_cases <- dental_records_long_final %>%
  group_by(age_band2, dental_category2) %>%
  summarise(total_cases = n(), .groups = "drop") %>%
  arrange(age_band2)

# 2. combine with py
ir_age2_dcat <- age2_dcat_cases %>%
  left_join(age_band2_py, by = "age_band2")

# 3. calculate IR
ir_age2_dcat$ir1000 <- ir_age2_dcat$total_cases / ir_age2_dcat$total_person_year * 1000

# 4. correct variable name for plot
ir_age2_dcat <- ir_age2_dcat %>%
  mutate(`Dental problem categories` = factor(dental_category2, levels = rev(sort(unique(dental_category2))))) %>%
  mutate(`Age group` = factor(age_band2))

# 5. plot annual 6eth_rgn ir
hm_ir_age2_dcat <- make_heatmap(
  data    = ir_age2_dcat,
  x       = `Dental problem categories`,
  y       = `Age group`,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by age group and dental problem categories",
  xlab        = "Dental problem categories",
  ylab        = "Age group",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_age2_dcat.pdf"),
  width       = 12,
  height      = 6,
  dpi         = 600
)

save(ir_age2_dcat, file = paste0(processed_data, "/final_curated_data/obj1_results/ir_age2_dcat.RData"))
```


### Step 2.2: age2 ximd 
```{r}
# 1. create summary table
age_band2_imd_cases2 <- dental_records_long_final %>%
  group_by(age_band2, imd_quintile) %>%
  summarise(total_cases = n(), .groups = "drop") %>%
  arrange(age_band2)

# Loop one band at a time, pass the NAME of that PY column to make_ir_table()
age_band2_imd_py <- vector("list", length(py_days))
names(age_band2_imd_py) <- names(py_days)

for (band in names(py_days)) {
  colname <- py_days[[band]]

  tmp <- make_ir_table(
    data  = age_band2_py_cut,
    py    = colname,                 # pass the NAME, not the vector
    cases = "dental_visit_count",    # name as string
    var   = "imd_quintile"                 # name as string
  )

  age_band2_imd_py[[band]] <- dplyr::mutate(tmp, age_band2 = band)
}

ir_ab2_imd <- dplyr::bind_rows(age_band2_imd_py)


# 1. clean py data frame
ir_ab2_imd <- ir_ab2_imd %>%
  dplyr::select(-total_cases, -variable) %>%   # drop columns
  dplyr::rename(imd_quintile = level)                # rename 'level' -> 'gender'

# 2. combine with case counts
ir_ab2_imd <- ir_ab2_imd %>%
  left_join(age_band2_imd_cases2, by = c("age_band2", "imd_quintile"))

# 3. correct variable name for plot 
ir_ab2_imd <- ir_ab2_imd %>% mutate(
    # Recode only 1 and 5, keep others as they are
    IMD = case_when(
      imd_quintile == 1 ~ "1 (Least Deprived)",
      imd_quintile == 5 ~ "5 (Most Deprived)",
      TRUE              ~ as.character(imd_quintile)   # keep other values
    ),
    # Turn into factor with reverse alphabetical order
    IMD = factor(IMD,
                          levels = c("1 (Least Deprived)", "2","3","4","5 (Most Deprived)","Unknown"))) %>%
  mutate(age_band2 = factor(age_band2,
                          levels = c("0-4","5-9","10-14","15-19","20-24","25-29","30-34",
                 "35-39","40-44","45-49","50-54","55-59","60-64","65-69",
                 "70-74","75-79","80-84","85-89","90+")))


# 4. calculate ir1000
ir_ab2_imd$ir1000 <- ir_ab2_imd$total_cases / ir_ab2_imd$total_person_year * 1000

# 5. plot annual 6eth_rgn ir
hm_ir_age2_imd <- make_heatmap(
  data    = ir_ab2_imd %>% filter(IMD != "Unknown"),
  x       = IMD,
  y       = age_band2,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by age and IMD",
  xlab        = "IMD",
  ylab        = "Age Band",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_age2_imd.jpeg"),
  width       = 12,
  height      = 6,
  dpi         = 600
)
save(ir_ab2_imd, file = paste0(processed_data, "/final_curated_data/obj1_results/ir_ab2_imd.RData"))
```
### Step 2.3: age2 xethnicity
```{r}
# 1. create summary table
# 1. create summary table
age_band2_eth_cases2 <- dental_records_long_final %>%
  group_by(age_band2, final_ethnicity_groups) %>%
  summarise(total_cases = n(), .groups = "drop") %>%
  arrange(age_band2) %>%
  dplyr::rename(ethnicity = final_ethnicity_groups)  

# Loop one band at a time, pass the NAME of that PY column to make_ir_table()
age_band2_eth_py <- vector("list", length(py_days))
names(age_band2_eth_py) <- names(py_days)

for (band in names(py_days)) {
  colname <- py_days[[band]]

  tmp <- make_ir_table(
    data  = age_band2_py_cut,
    py    = colname,                 # pass the NAME, not the vector
    cases = "dental_visit_count",    # name as string
    var   = "final_ethnicity_groups"                 # name as string
  )

  age_band2_eth_py[[band]] <- dplyr::mutate(tmp, age_band2 = band)
}

ir_ab2_eth <- dplyr::bind_rows(age_band2_eth_py)


# 1. clean py data frame
ir_ab2_eth <- ir_ab2_eth %>%
  dplyr::select(-total_cases, -variable) %>%   # drop columns
  dplyr::rename(ethnicity = level)                # rename 'level' -> 'gender'

# 2. combine with case counts
ir_ab2_eth <- ir_ab2_eth %>%
  left_join(age_band2_eth_cases2, by = c("age_band2", "ethnicity"))

# 3. correct variable name for plot 
ir_ab2_eth <- ir_ab2_eth %>% mutate(
    # Recode only 1 and 5, keep others as they are
    ethnicity = case_when(
      ethnicity == "Black, Black British, Caribbean or African" ~ "Black",
      ethnicity == "Asian or Asian British" ~ "Asian",
      ethnicity == "Other ethnic group" ~ "Other",
      ethnicity =="Mixed or multiple ethnic groups" ~ "Mixed or Multiple",
      TRUE              ~ as.character(ethnicity)   # keep other values
    ),
    # Turn into factor with reverse alphabetical order
    ethnicity = factor(ethnicity,
                          levels = c("White","Asian","Black", "Mixed or Multiple", "Other","Not specified"))) %>%
  mutate(age_band2 = factor(age_band2,
                          levels = c("0-4","5-9","10-14","15-19","20-24","25-29","30-34",
                 "35-39","40-44","45-49","50-54","55-59","60-64","65-69",
                 "70-74","75-79","80-84","85-89","90+")))


# 4. calculate ir1000
ir_ab2_eth$ir1000 <- ir_ab2_eth$total_cases / ir_ab2_eth$total_person_year * 1000

# 5. plot annual 6eth_rgn ir
hm_ir_age2_eth <- make_heatmap(
  data    = ir_ab2_eth %>% filter(ethnicity != "Not specified"),
  x       = ethnicity,
  y       = age_band2,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by ethnicity and eth",
  xlab        = "Ethnicity",
  ylab        = "Age Band",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_age2_eth.jpeg"),
  width       = 12,
  height      = 6,
  dpi         = 600
)
save(ir_ab2_eth, file = paste0(processed_data, "/final_curated_data/obj1_results/ir_ab2_eth.RData"))
```

### Step 2.4: age2 x region

```{r}
# 1. create summary table
# 1. create summary table
age_band2_rgn_cases2 <- dental_records_long_final %>%
  group_by(age_band2, region) %>%
  summarise(total_cases = n(), .groups = "drop") %>%
  arrange(age_band2)

# Loop one band at a time, pass the NAME of that PY column to make_ir_table()
age_band2_rgn_py <- vector("list", length(py_days))
names(age_band2_rgn_py) <- names(py_days)

for (band in names(py_days)) {
  colname <- py_days[[band]]

  tmp <- make_ir_table(
    data  = age_band2_py_cut,
    py    = colname,                 # pass the NAME, not the vector
    cases = "dental_visit_count",    # name as string
    var   = "region"                 # name as string
  )

  age_band2_rgn_py[[band]] <- dplyr::mutate(tmp, age_band2 = band)
}

ir_ab2_rgn <- dplyr::bind_rows(age_band2_rgn_py)


# 1. clean py data frame
ir_ab2_rgn <- ir_ab2_rgn %>%
  dplyr::select(-total_cases, -variable) %>%   # drop columns
  dplyr::rename(region = level)                # rename 'level' -> 'gender'

# 2. combine with case counts
ir_ab2_rgn <- ir_ab2_rgn %>%
  left_join(age_band2_rgn_cases2, by = c("age_band2", "region"))

# 3. correct variable name for plot 
ir_ab2_rgn <- ir_ab2_rgn %>%
  mutate(age_band2 = factor(age_band2,
                          levels = c("0-4","5-9","10-14","15-19","20-24","25-29","30-34",
                 "35-39","40-44","45-49","50-54","55-59","60-64","65-69",
                 "70-74","75-79","80-84","85-89","90+")))


# 4. calculate ir1000
ir_ab2_rgn$ir1000 <- ir_ab2_rgn$total_cases / ir_ab2_rgn$total_person_year * 1000

# 5. plot annual 6rgn_rgn ir
hm_ir_age2_rgn <- make_heatmap(
  data    = ir_ab2_rgn,
  x       = region,
  y       = age_band2,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by region and age",
  xlab        = "Region",
  ylab        = "Age Band",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_age2_rgn.jpeg"),
  width       = 12,
  height      = 6,
  dpi         = 600
)
save(ir_ab2_rgn, file = paste0(processed_data, "/final_curated_data/obj1_results/ir_ab2_rgn.RData"))
```

### Step 2.5: age2 x gender

```{r}

# 3. correct variable name for plot 
ir_ab2_gender <- ir_ab2_gender %>%
  mutate(age_band2 = factor(age_band2,
                          levels = c("0-4","5-9","10-14","15-19","20-24","25-29","30-34",
                 "35-39","40-44","45-49","50-54","55-59","60-64","65-69",
                 "70-74","75-79","80-84","85-89","90+")))



# 5. plot annual 6rgn_rgn ir
hm_ir_age2_gender <- make_heatmap(
  data    = ir_ab2_gender,
  x       = gender,
  y       = age_band2,
  fill        = ir1000,
  facet_nrow  = 1,
  palette     = tailwind_cols,
  legend_name = "IR per 1,000 PY",
  title       = "Incidence heatmap by gender and age",
  xlab        = "Gender",
  ylab        = "Age Band",
  outfile     = paste0(tables_figures, "/objective 1/fig3_ir_hm_age2_gender.jpeg"),
  width       = 12,
  height      = 6,
  dpi         = 600
)

save(ir_ab2_gender, file = paste0(processed_data, "/final_curated_data/obj1_results/ir_ab2_gender.RData"))
```


# Section 6: model the before covid and after covid 

## Step 1: check the distribution of dental count
```{r}

# 1. check the variable
colnames(sample_final)

# 2. distribution
hist(sample_final$dental_visit_count)
table(sample_final$dent_visit_count2)
table(sample_final$had_dental_visit) # use logistic regression
```


## Step 2: define case or not in the new data frame
Note: code ever dental visited but are control in that specific year as 2 = ever dvisited and remove from that year analysis 
```{r}
sample_final <- sample_final %>%
  mutate(fu_start_year = lubridate::year(fu_start_date))

sample_final <- sample_final %>%
  mutate(censor_year = lubridate::year(censor_date))
# 1. store a seperate data frame 
lg_sample_final <- sample_final[,c("patid", "fu_start_date", "fu_start_year", "censor_date", "censor_year", "yob","had_dental_visit","gender","imd_quintile", "final_ethnicity_groups","region")]

lg_sample_final$final_ethnicity_groups <- factor(
  sample_final$final_ethnicity_groups,
  levels = c(
    "White",
    "Asian or Asian British",
    "Black, Black British, Caribbean or African",
    "Mixed or multiple ethnic groups",
    "Not specified",
    "Other ethnic group"
  )
)


# 2. create had dental visit variable for each year 

years <- 1998:2023

# create a matrix of 0/1 visit flags per year
dvisit_mat <- sapply(
  years,
  function(y) {
    as.integer(
      lg_sample_final$patid %in% dental_records_long_final$patid[
        dental_records_long_final$obsyear == y
      ]
    )
  }
)

colnames(dvisit_mat) <- paste0("dvisit_", years)

# bind the new columns to lg_sample_final
lg_sample_final <- dplyr::bind_cols(
  lg_sample_final,
  as_tibble(dvisit_mat)
)
table(lg_sample_final$dvisit_2014 ==1, lg_sample_final$dvisit_2022 ==0)
# 3. code every dental visited as 2
years <- 1998:2023

for (i in 2:length(years)) {
  current_year <- years[i]              # e.g., 2023
  previous_years <- years[1:(i-1)]      # e.g., 1998–2022
  
  current_var <- paste0("dvisit_", current_year)
  previous_vars <- paste0("dvisit_", previous_years)
  
  # change current year to 2 if:
  #   - current == 0
  #   - ANY previous year == 1
  lg_sample_final[[current_var]][
    lg_sample_final[[current_var]] == 0 &
    rowSums(lg_sample_final[previous_vars] == 1, na.rm = TRUE) > 0
  ] <- 2
}


table(lg_sample_final$dvisit_2014 ==1, lg_sample_final$dvisit_2022 ==0)


```


## Step 3: define the sample for each year cohort

```{r}


# 1. loop to get the correct filter on cases from previous year
lg_sample_final_year_list <- list()
years <- 1998:2023

lg_sample_final_year_list<- lapply(years, function(y) {
  dvar <- paste0("dvisit_", y)
  
  # keep only rows where that year's dvisit is NOT 2
  lg_sample_final[ is.na(lg_sample_final[[dvar]]) | lg_sample_final[[dvar]] != 2, ]
})

# Name list elements using the numeric years
names(lg_sample_final_year_list) <- as.character(years)




# 2. loop to filter on the follow up start time and end time 
for (y in years) {
  nm <- as.character(y)          # list element name
  
  df <- lg_sample_final_year_list[[nm]]
  
  df <- df %>%
    filter(
      fu_start_year <= y,
      censor_year   >= y
    )
  
  lg_sample_final_year_list[[nm]] <- df
}


```


## Step 4: define age 
```{r}
source("CPRD_define_variable2_v1.R")
# 1. define age and age band and age group
for (y in years) {
  nm <- as.character(y)
  
  df <- lg_sample_final_year_list[[nm]]
  
  # age in that specific year
  df$age <- y - df$yob
  
  df$age_band <- age_band(df$age)
  
  lg_sample_final_year_list[[nm]] <- df
}



# 2. create adult and children list 
lg_children_year_list <- lapply(lg_sample_final_year_list, function(df) {
  df[df$age_band %in% c("0-4","5-9","10-14",  "15-19"), ]
})

lg_adult_year_list <- lapply(lg_sample_final_year_list, function(df) {
  df[df$age_band %in% c("20-24" , "25-29" , "30-34",  "35-39",  "40-44",  "45-49",
  "50-54",  "55-59"), ]
})

lg_older_adult_year_list <- lapply(lg_sample_final_year_list, function(df) {
  df[df$age_band %in% c( "60-64",  "65-69",  "70-74",  "75-79",
"80-84","85-89","90+"), ]
})


lapply(lg_children_year_list, function(df) {
     c(
       table(df$age_band)
     )
 }) #checked

lapply(lg_adult_year_list, function(df) {
     c(
       table(df$age_band)
     )
 })

lapply(lg_older_adult_year_list, function(df) {
     c(
       table(df$age_band)
     )
 })
```


## Step 4: models for adults and older adults
```{r}
# 1. empty list to store model objects
lg_adult_models_list <- list()

for (i in 1998:2023) {
  
  # correct outcome name dvisit_2019, dvisit_2020, ...
  formula_i <- as.formula(
    paste0("dvisit_", i,
           " ~ gender + imd_quintile + age_band + final_ethnicity_groups + region")
  )
  
  df_i <- lg_adult_year_list[[as.character(i)]]
  
  model_i <- glm(
    formula = formula_i,
    data    = df_i,
    family  = binomial(link = "logit")
  )
  
  lg_adult_models_list[[paste0("lg_adult_", i)]] <- model_i
}

# 1. empty list to store model objects
lg_older_adult_models_list <- list()

for (i in 1998:2023) {
  
  # correct outcome name dvisit_2019, dvisit_2020, ...
  formula_i <- as.formula(
    paste0("dvisit_", i,
           " ~ gender + imd_quintile + age_band + final_ethnicity_groups + region")
  )
  
  df_i <- lg_older_adult_year_list[[as.character(i)]]
  
  model_i <- glm(
    formula = formula_i,
    data    = df_i,
    family  = binomial(link = "logit")
  )
  
  lg_older_adult_models_list[[paste0("lg_older_adult_", i)]] <- model_i
}


```

## Step 5: extract OR and CI
```{r}
# 1. older_adult resuls
lg_older_adult_results <- purrr::map_dfr(
  names(lg_older_adult_models_list),
  function(model_name) {
    
    model <- lg_older_adult_models_list[[model_name]]
    
    broom::tidy(model) %>%
      dplyr::mutate(
        model  = model_name,
        OR     = exp(estimate),
        CI_low = exp(estimate - 1.96 * std.error),
        CI_high= exp(estimate + 1.96 * std.error)
      ) %>%
      dplyr::select(model, term, estimate, OR, CI_low, CI_high, p.value)
  }
)


# 2. list 2
lg_adult_results <- purrr::map_dfr(
  names(lg_adult_models_list),
  function(model_name) {
    
    model <- lg_adult_models_list[[model_name]]
    
    broom::tidy(model) %>%
      dplyr::mutate(
        model  = model_name,
        OR     = exp(estimate),
        CI_low = exp(estimate - 1.96 * std.error),
        CI_high= exp(estimate + 1.96 * std.error)
      ) %>%
      dplyr::select(model, term, estimate, OR, CI_low, CI_high, p.value)
  }
)

# 2. list 2
lg_children_results <- purrr::map_dfr(
  names(lg_children_models_list),
  function(model_name) {
    
    model <- lg_children_models_list[[model_name]]
    
    broom::tidy(model) %>%
      dplyr::mutate(
        model  = model_name,
        OR     = exp(estimate),
        CI_low = exp(estimate - 1.96 * std.error),
        CI_high= exp(estimate + 1.96 * std.error)
      ) %>%
      dplyr::select(model, term, estimate, OR, CI_low, CI_high, p.value)
  }
)
```

## format the results 
```{r}
# 1. format 
lg_adult_results_format <- lg_adult_results %>%
  filter(term != "(Intercept)") %>%
  mutate(
    Year = as.integer(sub("lg_adult_", "", model)),

    Level = case_when(
      grepl("^gender", term) ~
        paste("Sex", sub("^gender", "", term)),

      grepl("^imd_quintile", term) ~
        paste("IMD", sub("^imd_quintile", "", term)),

      grepl("^age_band", term) ~
        paste("Age", sub("^age_band", "", term)),

      grepl("^region", term) ~
        paste("English region", sub("^region", "", term)),

      grepl("^final_ethnicity_groups", term) ~
        paste("Ethnicity", sub("^final_ethnicity_groups", "", term)),

      TRUE ~ term
    ),

    Level = case_when(
      Level == "Ethnicity Asian or Asian British" ~ "Ethnicity Asian",
      Level == "Ethnicity Black, Black British, Caribbean or African" ~ "Ethnicity Black",
      Level == "Ethnicity Mixed or multiple ethnic groups" ~ "Ethnicity Mixed",
      Level == "Ethnicity Other ethnic group" ~ "Ethnicity Other",

      Level == "IMD 1" ~ "IMD 1 (Least Deprived)",
      Level == "IMD 5" ~ "IMD 5 (Most Deprived)",

      TRUE ~ Level
    ),
        P = p.value,

    `OR (95% CI)` = sprintf(
      "%.2f (%.2f–%.2f)",
      OR, CI_low, CI_high
    )
  ) %>%
  dplyr::select(
    Year,
    Level,
    `OR (95% CI)`,
    P
  )





# 2. format 
lg_older_adult_results_format <- lg_older_adult_results %>%
  filter(term != "(Intercept)") %>%
  mutate(
    Year = as.integer(sub("lg_older_adult_", "", model)),

    Level = case_when(
      grepl("^gender", term) ~
        paste("Sex", sub("^gender", "", term)),

      grepl("^imd_quintile", term) ~
        paste("IMD", sub("^imd_quintile", "", term)),

      grepl("^age_band", term) ~
        paste("Age", sub("^age_band", "", term)),

      grepl("^region", term) ~
        paste("English region", sub("^region", "", term)),

      grepl("^final_ethnicity_groups", term) ~
        paste("Ethnicity", sub("^final_ethnicity_groups", "", term)),

      TRUE ~ term
    ),

    Level = case_when(
      Level == "Ethnicity Asian or Asian British" ~ "Ethnicity Asian",
      Level == "Ethnicity Black, Black British, Caribbean or African" ~ "Ethnicity Black",
      Level == "Ethnicity Mixed or multiple ethnic groups" ~ "Ethnicity Mixed",
      Level == "Ethnicity Other ethnic group" ~ "Ethnicity Other",

      Level == "IMD 1" ~ "IMD 1 (Least Deprived)",
      Level == "IMD 5" ~ "IMD 5 (Most Deprived)",

      TRUE ~ Level
    ),

    P = p.value,
    `OR (95% CI)` = sprintf(
      "%.2f (%.2f–%.2f)",
      OR, CI_low, CI_high
    )
  ) %>%
  dplyr::select(
    Year,
    Level,
    `OR (95% CI)`,
    P
  )

# adding crude OR and reference row


```


### events for each year and each level
```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(writexl)


#=========================================================
# 1. Function to format logistic regression results
#=========================================================

format_lg_results <- function(results_data, model_prefix) {

  results_data %>%
    filter(term != "(Intercept)") %>%
    mutate(
      Year = as.integer(
        sub(
          paste0("^", model_prefix),
          "",
          model
        )
      ),

      Level = case_when(
        grepl("^gender", term) ~
          paste("Sex", sub("^gender", "", term)),

        grepl("^imd_quintile", term) ~
          paste("IMD", sub("^imd_quintile", "", term)),

        grepl("^age_band", term) ~
          paste("Age", sub("^age_band", "", term)),

        grepl("^region", term) ~
          paste(
            "English region",
            sub("^region", "", term)
          ),

        grepl("^final_ethnicity_groups", term) ~
          paste(
            "Ethnicity",
            sub("^final_ethnicity_groups", "", term)
          ),

        TRUE ~ term
      ),

      Level = case_when(
        Level == "Ethnicity Asian or Asian British" ~
          "Ethnicity Asian",

        Level ==
          "Ethnicity Black, Black British, Caribbean or African" ~
          "Ethnicity Black",

        Level ==
          "Ethnicity Mixed or multiple ethnic groups" ~
          "Ethnicity Mixed",

        Level == "Ethnicity Other ethnic group" ~
          "Ethnicity Other",

        Level == "IMD 1" ~
          "IMD 1 (Least Deprived)",

        Level == "IMD 5" ~
          "IMD 5 (Most Deprived)",

        TRUE ~ Level
      ),

      P = p.value,

      `OR (95% CI)` = sprintf(
        "%.2f (%.2f–%.2f)",
        OR,
        CI_low,
        CI_high
      )
    ) %>%
    dplyr::select(
      model,
      Year,
      Level,
      `OR (95% CI)`,
      P
    )
}


#=========================================================
# 2. Format the three logistic regression result tables
#=========================================================

lg_children_results_format <- format_lg_results(
  results_data = lg_children_results,
  model_prefix = "lg_children_"
)

lg_adult_results_format <- format_lg_results(
  results_data = lg_adult_results,
  model_prefix = "lg_adult_"
)

lg_older_adult_results_format <- format_lg_results(
  results_data = lg_older_adult_results,
  model_prefix = "lg_older_adult_"
)


#=========================================================
# 3. Function to count events and non-events
#    for every variable level in every list element
#=========================================================

lg_count_events_by_level <- function(
  lg_data_list,
  model_prefix
) {

  purrr::imap_dfr(
    lg_data_list,
    function(lg_data, list_name) {

      # Extract year from the name of each list element
      current_year <- as.integer(
        stringr::str_extract(
          as.character(list_name),
          "\\d{4}"
        )
      )

      # Create model name matching the regression results
      current_model <- paste0(
        model_prefix,
        current_year
      )

      #---------------------------------------------------
      # Sex counts
      #---------------------------------------------------

      sex_counts <- lg_data %>%
        filter(
          !is.na(gender),
          as.character(had_dental_visit) %in% c("Yes", "No")
        ) %>%
        mutate(
          Level = paste(
            "Sex",
            as.character(gender)
          )
        ) %>%
        count(
          Level,
          had_dental_visit,
          name = "N"
        )

      #---------------------------------------------------
      # IMD counts
      #---------------------------------------------------

      imd_counts <- lg_data %>%
        filter(
          !is.na(imd_quintile),
          as.character(had_dental_visit) %in% c("Yes", "No")
        ) %>%
        mutate(
          Level = case_when(
            as.character(imd_quintile) == "1" ~
              "IMD 1 (Least Deprived)",

            as.character(imd_quintile) == "5" ~
              "IMD 5 (Most Deprived)",

            TRUE ~ paste(
              "IMD",
              as.character(imd_quintile)
            )
          )
        ) %>%
        count(
          Level,
          had_dental_visit,
          name = "N"
        )

      #---------------------------------------------------
      # Age counts
      #---------------------------------------------------

      age_counts <- lg_data %>%
        filter(
          !is.na(age_band),
          as.character(had_dental_visit) %in% c("Yes", "No")
        ) %>%
        mutate(
          Level = paste(
            "Age",
            as.character(age_band)
          )
        ) %>%
        count(
          Level,
          had_dental_visit,
          name = "N"
        )

      #---------------------------------------------------
      # Region counts
      #---------------------------------------------------

      region_counts <- lg_data %>%
        filter(
          !is.na(region),
          as.character(had_dental_visit) %in% c("Yes", "No")
        ) %>%
        mutate(
          Level = paste(
            "English region",
            as.character(region)
          )
        ) %>%
        count(
          Level,
          had_dental_visit,
          name = "N"
        )

      #---------------------------------------------------
      # Ethnicity counts
      #---------------------------------------------------

      ethnicity_counts <- lg_data %>%
        filter(
          !is.na(final_ethnicity_groups),
          as.character(had_dental_visit) %in% c("Yes", "No")
        ) %>%
        mutate(
          Level = case_when(
            as.character(final_ethnicity_groups) ==
              "Asian or Asian British" ~
              "Ethnicity Asian",

            as.character(final_ethnicity_groups) ==
              "Black, Black British, Caribbean or African" ~
              "Ethnicity Black",

            as.character(final_ethnicity_groups) ==
              "Mixed or multiple ethnic groups" ~
              "Ethnicity Mixed",

            as.character(final_ethnicity_groups) ==
              "Other ethnic group" ~
              "Ethnicity Other",

            TRUE ~ paste(
              "Ethnicity",
              as.character(final_ethnicity_groups)
            )
          )
        ) %>%
        count(
          Level,
          had_dental_visit,
          name = "N"
        )

      #---------------------------------------------------
      # Combine all variable-level counts
      #---------------------------------------------------

      bind_rows(
        sex_counts,
        imd_counts,
        age_counts,
        region_counts,
        ethnicity_counts
      ) %>%
        mutate(
          model = current_model,

          Status = case_when(
            as.character(had_dental_visit) == "Yes" ~
              "Event N",

            as.character(had_dental_visit) == "No" ~
              "Non-event N",

            TRUE ~ NA_character_
          )
        ) %>%
        filter(!is.na(Status)) %>%
        dplyr::select(
          model,
          Level,
          Status,
          N
        ) %>%
        complete(
          model,
          Level,
          Status = c(
            "Event N",
            "Non-event N"
          ),
          fill = list(N = 0L)
        ) %>%
        pivot_wider(
          names_from = Status,
          values_from = N,
          values_fill = 0
        ) %>%
        dplyr::select(
          model,
          Level,
          `Event N`,
          `Non-event N`
        )
    }
  )
}


#=========================================================
# 4. Count events and non-events for the three lists
#=========================================================

lg_children_event_counts <- lg_count_events_by_level(
  lg_data_list = lg_children_year_list,
  model_prefix = "lg_children_"
)

lg_adult_event_counts <- lg_count_events_by_level(
  lg_data_list = lg_adult_year_list,
  model_prefix = "lg_adult_"
)

lg_older_adult_event_counts <- lg_count_events_by_level(
  lg_data_list = lg_older_adult_year_list,
  model_prefix = "lg_older_adult_"
)


#=========================================================
# 6. Function to join counts and format final tables
#=========================================================

add_level_event_counts <- function(
  results_table,
  event_counts
) {

  results_table %>%
    left_join(
      event_counts,
      by = c(
        "model",
        "Level"
      )
    ) %>%
    mutate(
      `N events` = case_when(
        is.na(`Event N`) ~ NA_character_,

        TRUE ~ formatC(
          `Event N`,
          format = "f",
          digits = 0,
          big.mark = ","
        )
      ),

      `N non-events` = case_when(
        is.na(`Non-event N`) ~ NA_character_,

        TRUE ~ formatC(
          `Non-event N`,
          format = "f",
          digits = 0,
          big.mark = ","
        )
      ),

      `OR (95% CI)` = gsub(
        "\\.",
        "·",
        `OR (95% CI)`
      ),

      P = case_when(
        is.na(P)   ~ NA_character_,
        P < 0.0001 ~ "<0.0001",
        P < 0.001  ~ sprintf("%.5f", P),
        P < 0.01   ~ sprintf("%.4f", P),
        P < 0.1    ~ sprintf("%.3f", P),
        TRUE       ~ sprintf("%.2f", P)
      ),

      P = gsub(
        "\\.",
        "·",
        P
      )
    ) %>%
    dplyr::select(
      Year,
      Level,
      `N events`,
      `N non-events`,
      `OR (95% CI)`,
      P
    ) %>%
    arrange(
      Year,
      Level
    )
}


#=========================================================
# 7. Create final formatted tables
#=========================================================

lg_children_results_format <- add_level_event_counts(
  results_table = lg_children_results_format,
  event_counts = lg_children_event_counts
)

lg_adult_results_format <- add_level_event_counts(
  results_table = lg_adult_results_format,
  event_counts = lg_adult_event_counts
)

lg_older_adult_results_format <- add_level_event_counts(
  results_table = lg_older_adult_results_format,
  event_counts = lg_older_adult_event_counts
)



```
### Children: add ref rows
```{r}
##########################################
# CHILDREN LOGISTIC REGRESSION TABLE
# Add reference rows + crude OR + crude P
##########################################

# ============================================================
# 1. Rename existing OR and P as adjusted results
# ============================================================

lg_children_results_format <- lg_children_results_format %>%
  dplyr::rename(
    `Adjusted OR (95% CI)` = `OR (95% CI)`,
    `Adjusted P` = P
  )


# ============================================================
# 2. Create Variable for EXISTING rows
# ============================================================

lg_children_results_format <- lg_children_results_format %>%
  dplyr::mutate(
    Variable = dplyr::case_when(
      grepl("^Sex ", Level) ~ "Sex",
      grepl("^IMD ", Level) ~ "IMD",
      grepl("^Age ", Level) ~ "Age",
      grepl("^English region ", Level) ~ "English region",
      grepl("^Ethnicity ", Level) ~ "Ethnicity",
      TRUE ~ NA_character_
    )
  )


# ============================================================
# 3. Convert existing N events / N non-events back to numeric
#    e.g. "1,000" -> 1000
# ============================================================

lg_children_results_format <- lg_children_results_format %>%
  dplyr::mutate(
    `N events` = as.numeric(
      gsub(",", "", `N events`)
    ),
    `N non-events` = as.numeric(
      gsub(",", "", `N non-events`)
    )
  )


# ============================================================
# 4. Define reference levels
# ============================================================

reference_levels_children <- tibble::tribble(
  ~variable,                 ~reference_level,   ~Level,                         ~Variable,
  "gender",                  "Male",             "Sex Male",                     "Sex",
  "imd_quintile",            "1",                "IMD 1 (Least Deprived)",       "IMD",
  "age_band",                "0-4",              "Age 0-4",                      "Age",
  "region",                  "East Midlands",    "English region East Midlands", "English region",
  "final_ethnicity_groups",  "White",            "Ethnicity White",              "Ethnicity"
)


# ============================================================
# 5. Extract reference-level Event N / Non-event N
#    from each yearly dataset
# ============================================================

lg_children_reference_rows <- purrr::imap_dfr(
  lg_children_year_list,
  function(lg_data, lg_year_name) {

    purrr::pmap_dfr(
      reference_levels_children,
      function(variable, reference_level, Level, Variable) {

        lg_data %>%
          dplyr::filter(
            as.character(.data[[variable]]) == reference_level
          ) %>%
          dplyr::count(
            had_dental_visit,
            name = "N"
          ) %>%
          tidyr::complete(
            had_dental_visit = c("Yes", "No"),
            fill = list(N = 0L)
          ) %>%
          tidyr::pivot_wider(
            names_from = had_dental_visit,
            values_from = N,
            values_fill = 0
          ) %>%
          dplyr::transmute(
            Year = as.integer(lg_year_name),
            Level = Level,
            `N events` = as.numeric(Yes),
            `N non-events` = as.numeric(No),
            `Adjusted OR (95% CI)` = "Reference",
            `Adjusted P` = NA_character_,
            Variable = Variable
          )
      }
    )
  }
)


# ============================================================
# 6. Bind reference rows back to existing formatted results
# ============================================================

lg_children_results_format <- dplyr::bind_rows(
  lg_children_results_format,
  lg_children_reference_rows
)


# ============================================================
# 6. Calculate crude OR, 95% CI and crude P
#    within each Year + Variable
# ============================================================

lg_children_results_format <- lg_children_results_format %>%
  dplyr::group_by(
    Year,
    Variable
  ) %>%
  dplyr::mutate(

    # reference counts
    ref_event = `N events`[
      `Adjusted OR (95% CI)` == "Reference"
    ][1],

    ref_non_event = `N non-events`[
      `Adjusted OR (95% CI)` == "Reference"
    ][1],

    # crude OR
    crude_OR =
      (`N events` / `N non-events`) /
      (ref_event / ref_non_event),

    # SE of log OR
    crude_SE = sqrt(
      1 / `N events` +
      1 / `N non-events` +
      1 / ref_event +
      1 / ref_non_event
    ),

    # 95% CI
    crude_lower = exp(
      log(crude_OR) - 1.96 * crude_SE
    ),

    crude_upper = exp(
      log(crude_OR) + 1.96 * crude_SE
    ),

    # crude Wald P
    crude_z =
      log(crude_OR) / crude_SE,

    crude_p_numeric =
      2 * stats::pnorm(
        -abs(crude_z)
      )
  ) %>%
  dplyr::ungroup()


# ============================================================
# 7. Format crude OR
# ============================================================

lg_children_results_format <- lg_children_results_format %>%
  dplyr::mutate(

    `Crude OR (95% CI)` = dplyr::case_when(

      `Adjusted OR (95% CI)` == "Reference" ~
        "Reference",

      TRUE ~ sprintf(
        "%.2f (%.2f–%.2f)",
        crude_OR,
        crude_lower,
        crude_upper
      )
    )
  )


# ============================================================
# 8. Format crude P using EXACTLY your original P rule
# ============================================================

lg_children_results_format <- lg_children_results_format %>%
  dplyr::mutate(

    `Crude P` = dplyr::case_when(
      `Adjusted OR (95% CI)` == "Reference" ~ NA_character_,
      is.na(crude_p_numeric) ~ NA_character_,
      crude_p_numeric < 0.0001 ~ "<0.0001",
      crude_p_numeric < 0.001  ~ sprintf("%.5f", crude_p_numeric),
      crude_p_numeric < 0.01   ~ sprintf("%.4f", crude_p_numeric),
      crude_p_numeric < 0.1    ~ sprintf("%.3f", crude_p_numeric),
      TRUE                     ~ sprintf("%.2f", crude_p_numeric)
    )
  ) %>%

  # replace decimal point with middle dot
  dplyr::mutate(
    `Crude P` = gsub(
      "\\.",
      "·",
      as.character(`Crude P`)
    )
  )


# ============================================================
# 9. Re-format N events / N non-events with commas
#    only after crude calculations are finished
# ============================================================

lg_children_results_format <- lg_children_results_format %>%
  dplyr::mutate(

    `N events` = format(
      `N events`,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    ),

    `N non-events` = format(
      `N non-events`,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    )
  )


# ============================================================
# 10. Keep final columns
# ============================================================

lg_children_results_format <- lg_children_results_format %>%
  dplyr::select(
    Year,
    Variable,
    Level,
    `N events`,
    `N non-events`,
    `Crude OR (95% CI)`,
    `Crude P`,
    `Adjusted OR (95% CI)`,
    `Adjusted P`
  )%>%
  dplyr::arrange(Year, Level)
```
### adult: add ref row
```{r}
##########################################
# ADULT LOGISTIC REGRESSION TABLE
# Add reference rows + crude OR + crude P
##########################################

# ============================================================
# 1. Rename existing OR and P as adjusted results
# ============================================================

lg_adult_results_format <- lg_adult_results_format %>%
  dplyr::rename(
    `Adjusted OR (95% CI)` = `OR (95% CI)`,
    `Adjusted P` = P
  )


# ============================================================
# 2. Create Variable for EXISTING rows
# ============================================================

lg_adult_results_format <- lg_adult_results_format %>%
  dplyr::mutate(
    Variable = dplyr::case_when(
      grepl("^Sex ", Level) ~ "Sex",
      grepl("^IMD ", Level) ~ "IMD",
      grepl("^Age ", Level) ~ "Age",
      grepl("^English region ", Level) ~ "English region",
      grepl("^Ethnicity ", Level) ~ "Ethnicity",
      TRUE ~ NA_character_
    )
  )


# ============================================================
# 3. Convert existing N events / N non-events back to numeric
#    e.g. "1,000" -> 1000
# ============================================================

lg_adult_results_format <- lg_adult_results_format %>%
  dplyr::mutate(
    `N events` = as.numeric(
      gsub(",", "", `N events`)
    ),
    `N non-events` = as.numeric(
      gsub(",", "", `N non-events`)
    )
  )


# ============================================================
# 4. Define reference levels
# ============================================================

reference_levels_adult <- tibble::tribble(
  ~variable,                ~reference_level, ~Level,                         ~Variable,
  "gender",                 "Male",           "Sex Male",                     "Sex",
  "imd_quintile",           "1",              "IMD 1 (Least Deprived)",       "IMD",
  "age_band",               "20-24",          "Age 20-24",                    "Age",
  "region",                 "East Midlands",  "English region East Midlands", "English region",
  "final_ethnicity_groups",  "White",          "Ethnicity White",              "Ethnicity"
)


# ============================================================
# 5. Extract reference-level Event N / Non-event N
#    from each yearly dataset
# ============================================================

lg_adult_reference_rows <- purrr::imap_dfr(
  lg_adult_year_list,
  function(lg_data, lg_year_name) {

    purrr::pmap_dfr(
      reference_levels_adult,
      function(variable, reference_level, Level, Variable) {

        lg_data %>%
          dplyr::filter(
            as.character(.data[[variable]]) == reference_level
          ) %>%
          dplyr::count(
            had_dental_visit,
            name = "N"
          ) %>%
          tidyr::complete(
            had_dental_visit = c("Yes", "No"),
            fill = list(N = 0L)
          ) %>%
          tidyr::pivot_wider(
            names_from = had_dental_visit,
            values_from = N,
            values_fill = 0
          ) %>%
          dplyr::transmute(
            Year = as.integer(lg_year_name),
            Level = Level,
            `N events` = as.numeric(Yes),
            `N non-events` = as.numeric(No),
            `Adjusted OR (95% CI)` = "Reference",
            `Adjusted P` = NA_character_,
            Variable = Variable
          )
      }
    )
  }
)


# ============================================================
# 6. Bind reference rows back to existing formatted results
# ============================================================

lg_adult_results_format <- dplyr::bind_rows(
  lg_adult_results_format,
  lg_adult_reference_rows
)


# ============================================================
# 7. Calculate crude OR, 95% CI and crude P
#    within each Year + Variable
# ============================================================

lg_adult_results_format <- lg_adult_results_format %>%
  dplyr::group_by(
    Year,
    Variable
  ) %>%
  dplyr::mutate(

    ref_event = `N events`[
      `Adjusted OR (95% CI)` == "Reference"
    ][1],

    ref_non_event = `N non-events`[
      `Adjusted OR (95% CI)` == "Reference"
    ][1],

    crude_OR =
      (`N events` / `N non-events`) /
      (ref_event / ref_non_event),

    crude_SE = sqrt(
      1 / `N events` +
      1 / `N non-events` +
      1 / ref_event +
      1 / ref_non_event
    ),

    crude_lower = exp(
      log(crude_OR) - 1.96 * crude_SE
    ),

    crude_upper = exp(
      log(crude_OR) + 1.96 * crude_SE
    ),

    crude_z =
      log(crude_OR) / crude_SE,

    crude_p_numeric =
      2 * stats::pnorm(
        -abs(crude_z)
      )
  ) %>%
  dplyr::ungroup()


# ============================================================
# 8. Format crude OR
# ============================================================

lg_adult_results_format <- lg_adult_results_format %>%
  dplyr::mutate(

    `Crude OR (95% CI)` = dplyr::case_when(

      `Adjusted OR (95% CI)` == "Reference" ~
        "Reference",

      TRUE ~ sprintf(
        "%.2f (%.2f–%.2f)",
        crude_OR,
        crude_lower,
        crude_upper
      )
    )
  )


# ============================================================
# 9. Format crude P using EXACTLY original P rule
# ============================================================

lg_adult_results_format <- lg_adult_results_format %>%
  dplyr::mutate(

    `Crude P` = dplyr::case_when(
      `Adjusted OR (95% CI)` == "Reference" ~ NA_character_,
      is.na(crude_p_numeric) ~ NA_character_,
      crude_p_numeric < 0.0001 ~ "<0.0001",
      crude_p_numeric < 0.001  ~ sprintf("%.5f", crude_p_numeric),
      crude_p_numeric < 0.01   ~ sprintf("%.4f", crude_p_numeric),
      crude_p_numeric < 0.1    ~ sprintf("%.3f", crude_p_numeric),
      TRUE                     ~ sprintf("%.2f", crude_p_numeric)
    )
  ) %>%

  dplyr::mutate(
    `Crude P` = gsub(
      "\\.",
      "·",
      as.character(`Crude P`)
    )
  )


# ============================================================
# 10. Re-format N events / N non-events with commas
# ============================================================

lg_adult_results_format <- lg_adult_results_format %>%
  dplyr::mutate(

    `N events` = format(
      `N events`,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    ),

    `N non-events` = format(
      `N non-events`,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    )
  )


# ============================================================
# 11. Keep final columns
# ============================================================

lg_adult_results_format <- lg_adult_results_format %>%
  dplyr::select(
    Year,
    Variable,
    Level,
    `N events`,
    `N non-events`,
    `Crude OR (95% CI)`,
    `Crude P`,
    `Adjusted OR (95% CI)`,
    `Adjusted P`
  ) %>%
  dplyr::arrange(Year, Level)
```

### older adult: add ref row
```{r}
# ============================================================
# 1. Rename existing OR and P as adjusted results
# ============================================================

lg_older_adult_results_format <- lg_older_adult_results_format %>%
  dplyr::rename(
    `Adjusted OR (95% CI)` = `OR (95% CI)`,
    `Adjusted P` = P
  )


# ============================================================
# 2. Create Variable for EXISTING rows
# ============================================================

lg_older_adult_results_format <- lg_older_adult_results_format %>%
  dplyr::mutate(
    Variable = dplyr::case_when(
      grepl("^Sex ", Level) ~ "Sex",
      grepl("^IMD ", Level) ~ "IMD",
      grepl("^Age ", Level) ~ "Age",
      grepl("^English region ", Level) ~ "English region",
      grepl("^Ethnicity ", Level) ~ "Ethnicity",
      TRUE ~ NA_character_
    )
  )


# ============================================================
# 3. Convert existing N events / N non-events back to numeric
# ============================================================

lg_older_adult_results_format <- lg_older_adult_results_format %>%
  dplyr::mutate(
    `N events` = as.numeric(
      gsub(",", "", `N events`)
    ),
    `N non-events` = as.numeric(
      gsub(",", "", `N non-events`)
    )
  )


# ============================================================
# 4. Define reference levels
# ============================================================

reference_levels_older_adult <- tibble::tribble(
  ~variable,                ~reference_level, ~Level,                         ~Variable,
  "gender",                 "Male",           "Sex Male",                     "Sex",
  "imd_quintile",           "1",              "IMD 1 (Least Deprived)",       "IMD",
  "age_band",               "60-64",          "Age 60-64",                    "Age",
  "region",                 "East Midlands",  "English region East Midlands", "English region",
  "final_ethnicity_groups",  "White",          "Ethnicity White",              "Ethnicity"
)


# ============================================================
# 5. Extract reference-level Event N / Non-event N
# ============================================================

lg_older_adult_reference_rows <- purrr::imap_dfr(
  lg_older_adult_year_list,
  function(lg_data, lg_year_name) {

    purrr::pmap_dfr(
      reference_levels_older_adult,
      function(variable, reference_level, Level, Variable) {

        lg_data %>%
          dplyr::filter(
            as.character(.data[[variable]]) == reference_level
          ) %>%
          dplyr::count(
            had_dental_visit,
            name = "N"
          ) %>%
          tidyr::complete(
            had_dental_visit = c("Yes", "No"),
            fill = list(N = 0L)
          ) %>%
          tidyr::pivot_wider(
            names_from = had_dental_visit,
            values_from = N,
            values_fill = 0
          ) %>%
          dplyr::transmute(
            Year = as.integer(lg_year_name),
            Level = Level,
            `N events` = as.numeric(Yes),
            `N non-events` = as.numeric(No),
            `Adjusted OR (95% CI)` = "Reference",
            `Adjusted P` = NA_character_,
            Variable = Variable
          )
      }
    )
  }
)


# ============================================================
# 6. Bind reference rows back to existing formatted results
# ============================================================

lg_older_adult_results_format <- dplyr::bind_rows(
  lg_older_adult_results_format,
  lg_older_adult_reference_rows
)


# ============================================================
# 7. Calculate crude OR, 95% CI and crude P
# ============================================================

lg_older_adult_results_format <- lg_older_adult_results_format %>%
  dplyr::group_by(
    Year,
    Variable
  ) %>%
  dplyr::mutate(

    ref_event = `N events`[
      `Adjusted OR (95% CI)` == "Reference"
    ][1],

    ref_non_event = `N non-events`[
      `Adjusted OR (95% CI)` == "Reference"
    ][1],

    crude_OR =
      (`N events` / `N non-events`) /
      (ref_event / ref_non_event),

    crude_SE = sqrt(
      1 / `N events` +
      1 / `N non-events` +
      1 / ref_event +
      1 / ref_non_event
    ),

    crude_lower = exp(
      log(crude_OR) - 1.96 * crude_SE
    ),

    crude_upper = exp(
      log(crude_OR) + 1.96 * crude_SE
    ),

    crude_z =
      log(crude_OR) / crude_SE,

    crude_p_numeric =
      2 * stats::pnorm(
        -abs(crude_z)
      )
  ) %>%
  dplyr::ungroup()


# ============================================================
# 8. Format crude OR
# ============================================================

lg_older_adult_results_format <- lg_older_adult_results_format %>%
  dplyr::mutate(

    `Crude OR (95% CI)` = dplyr::case_when(

      `Adjusted OR (95% CI)` == "Reference" ~
        "Reference",

      TRUE ~ sprintf(
        "%.2f (%.2f–%.2f)",
        crude_OR,
        crude_lower,
        crude_upper
      )
    )
  )


# ============================================================
# 9. Format crude P using EXACTLY original P rule
# ============================================================

lg_older_adult_results_format <- lg_older_adult_results_format %>%
  dplyr::mutate(

    `Crude P` = dplyr::case_when(
      `Adjusted OR (95% CI)` == "Reference" ~ NA_character_,
      is.na(crude_p_numeric) ~ NA_character_,
      crude_p_numeric < 0.0001 ~ "<0.0001",
      crude_p_numeric < 0.001  ~ sprintf("%.5f", crude_p_numeric),
      crude_p_numeric < 0.01   ~ sprintf("%.4f", crude_p_numeric),
      crude_p_numeric < 0.1    ~ sprintf("%.3f", crude_p_numeric),
      TRUE                     ~ sprintf("%.2f", crude_p_numeric)
    )
  ) %>%

  dplyr::mutate(
    `Crude P` = gsub(
      "\\.",
      "·",
      as.character(`Crude P`)
    )
  )


# ============================================================
# 10. Re-format N events / N non-events with commas
# ============================================================

lg_older_adult_results_format <- lg_older_adult_results_format %>%
  dplyr::mutate(

    `N events` = format(
      `N events`,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    ),

    `N non-events` = format(
      `N non-events`,
      big.mark = ",",
      scientific = FALSE,
      trim = TRUE
    )
  )


# ============================================================
# 11. Keep final columns
# ============================================================

lg_older_adult_results_format <- lg_older_adult_results_format %>%
  dplyr::select(
    Year,
    Variable,
    Level,
    `N events`,
    `N non-events`,
    `Crude OR (95% CI)`,
    `Crude P`,
    `Adjusted OR (95% CI)`,
    `Adjusted P`
  ) %>%
  dplyr::arrange(Year, Level)
```
### export

```{r}
#=========================================================
# 9. Export final tables
#=========================================================

write_xlsx(
  lg_children_results_format,
  paste0(
    "P:/Precision Health Programme/JW Research/DEEP study/",
    "GP Dental/Analysis_results_CG/tables_figures/",
    "lancet scaled/formatted/Table S10.xlsx"
  )
)

write_xlsx(
  lg_adult_results_format,
  paste0(
    "P:/Precision Health Programme/JW Research/DEEP study/",
    "GP Dental/Analysis_results_CG/tables_figures/",
    "lancet scaled/formatted/Table S11.xlsx"
  )
)

write_xlsx(
  lg_older_adult_results_format,
  paste0(
    "P:/Precision Health Programme/JW Research/DEEP study/",
    "GP Dental/Analysis_results_CG/tables_figures/",
    "lancet scaled/formatted/Table S12.xlsx"
  )
)
```

## Step 6: forest plot code 
```{r}
forest_panel <- function(df, year, term_levels, model_prefix, show_y = FALSE) {

  x_limits <- c(0.2, 6)
  x_breaks <- c(0.2, 0.5, 1, 2, 4,6)

  forest_col <- pink_to_red[30]

  df_year <- df %>%
    filter(model == paste0(model_prefix, year),
           term != "(Intercept)") %>%
    mutate(
      term_label = factor(term_label, levels = term_levels),

      ci_low_plot  = pmax(CI_low,  x_limits[1]),
      ci_high_plot = pmin(CI_high, x_limits[2]),

      arrow_left  = CI_low  < x_limits[1],
      arrow_right = CI_high > x_limits[2]
    )

  df_right <- df_year %>% filter(arrow_right)
  df_left  <- df_year %>% filter(arrow_left)

  # ONLY rows with no arrows get caps
  df_caps <- df_year %>% filter(!arrow_left & !arrow_right)

  # cap size (vertical half-height around y)
  cap_h <- 0.30  # increase/decrease to change cap size

  ggplot(df_year, aes(x = OR, y = term_label)) +
    geom_vline(xintercept = 1, linetype = 2, color = "gray50") +

    # point
    geom_point(size = 1.4, colour = forest_col) +

    # CI line (no caps) for everyone
    geom_segment(
      aes(x = ci_low_plot, xend = ci_high_plot,
          y = term_label,  yend = term_label),
      linewidth = 0.30,
      colour = forest_col
    ) +

    # ---- caps ONLY for non-arrow rows (manual vertical segments) ----
    # left cap
    geom_segment(
      data = df_caps,
      aes(x = ci_low_plot, xend = ci_low_plot,
          y = as.numeric(term_label) - cap_h,
          yend = as.numeric(term_label) + cap_h),
      inherit.aes = FALSE,
      linewidth = 0.30,
      colour = forest_col
    ) +
    # right cap
    geom_segment(
      data = df_caps,
      aes(x = ci_high_plot, xend = ci_high_plot,
          y = as.numeric(term_label) - cap_h,
          yend = as.numeric(term_label) + cap_h),
      inherit.aes = FALSE,
      linewidth = 0.30,
      colour = forest_col
    ) +

    # right arrow (no cap)
    geom_segment(
      data = df_right,
      aes(y = term_label, yend = term_label,
          x = x_limits[2] / 1.3, xend = x_limits[2]),
      inherit.aes = FALSE,
      linewidth = 0.3,
      colour = forest_col,
      arrow = arrow(length = unit(0.1, "cm"),
                    ends = "last", type = "closed")
    ) +

    # left arrow (no cap)
    geom_segment(
      data = df_left,
      aes(y = term_label, yend = term_label,
          x = x_limits[1], xend = x_limits[1] * 1),
      inherit.aes = FALSE,
      linewidth = 0.3,
      colour = forest_col,
      arrow = arrow(length = unit(0.1, "cm"),
                    ends = "first", type = "closed")
    ) +

    scale_y_discrete(drop = FALSE) +
    scale_x_log10(limits = x_limits, breaks = x_breaks) +

    labs(x = NULL, y = NULL, title = paste0(year)) +

    theme_classic(base_size = 11) +
    theme(
      plot.title = element_text(size = 10, hjust = 0.5, face = "bold"),
      axis.text.x = element_text(size = 8),
      axis.title.x = element_blank(),
      axis.line.x  = element_line(),

      axis.text.y  = if (show_y) element_text(size = 8) else element_blank(),
      axis.ticks.y = if (show_y) element_line()        else element_blank(),
      axis.line.y  = if (show_y) element_line()        else element_blank(),

      # remove right-hand y artefacts
      axis.text.y.right  = element_blank(),
      axis.ticks.y.right = element_blank(),
      axis.line.y.right  = element_blank(),

      panel.border = element_blank()
    )
}

```

## forest plot during covid 
covid and non covid plot creation use overlapped name should be run in sequence
to recreate any plots need to rerun all codes from step 7 to step 12 in this section
## Step 7: adult COVID forest (UPDATED)
```{r}



## ------------------------------------------------------------
## Adult: Create forest plot using existing forest_panel()
## ------------------------------------------------------------

# 1. Years and model names
adult_years  <- 2019:2023
adult_models <- paste0("lg_adult_", adult_years)

# 2. Clean term labels for adults (NO unknown IMD/ethnicity)
adult_main <- lg_adult_results %>%
  filter(model %in% adult_models) %>%
  mutate(
    term_label = case_when(
      # Sex
      term == "genderFemale" ~ "Female",

      # IMD (removed Unknown)
      term == "imd_quintile2" ~ "IMD 2",
      term == "imd_quintile3" ~ "IMD 3",
      term == "imd_quintile4" ~ "IMD 4",
      term == "imd_quintile5" ~ "IMD 5",

      # Adult age bands (non-reference levels)
      term == "age_band25-29" ~ "Age 25–29",
      term == "age_band30-34" ~ "Age 30–34",
      term == "age_band35-39" ~ "Age 35–39",
      term == "age_band40-44" ~ "Age 40–44",
      term == "age_band45-49" ~ "Age 45–49",
      term == "age_band50-54" ~ "Age 50–54",
      term == "age_band55-59" ~ "Age 55–59",


      # Ethnicity (removed Not specified)
      term == "final_ethnicity_groupsAsian or Asian British"                     ~ "Asian",
      term == "final_ethnicity_groupsBlack, Black British, Caribbean or African" ~ "Black",
      term == "final_ethnicity_groupsMixed or multiple ethnic groups"            ~ "Mixed ethnicity",
      term == "final_ethnicity_groupsOther ethnic group"                         ~ "Other ethnicity",

      # Regions
      term == "regionEast of England"          ~ "East of England",
      term == "regionLondon"                   ~ "London",
      term == "regionNorth East"               ~ "North East",
      term == "regionNorth West"               ~ "North West",
      term == "regionSouth East"               ~ "South East",
      term == "regionSouth West"               ~ "South West",
      term == "regionWest Midlands"            ~ "West Midlands",
      term == "regionYorkshire and The Humber" ~ "Yorkshire and The Humber",

      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(term_label))

# 3. Reference rows for adults (Age 20–24 is the reference)
adult_ref_rows <- expand.grid(
  model = adult_models,
  term_label = c(
    "Male (ref)",
    "Age 20–24 (ref)",
    "White (ref)",
    "IMD 1 (ref)",
    "East Midlands (ref)"
  ),
  stringsAsFactors = FALSE
) %>%
  mutate(
    term    = paste0("ref_", row_number()),
    OR      = 1,
    CI_low  = 1,
    CI_high = 1
  )

# 4. Combine adult dataset
lg_adult_plot <- bind_rows(adult_main, adult_ref_rows)

# 5. Term ordering with spacers (removed unknown IMD/ethnicity)
adult_terms <- c(
  # Sex
  "Male (ref)",
  "Female",
  " ",

  # Age
  "Age 20–24 (ref)",
  "Age 25–29",
  "Age 30–34",
  "Age 35–39",
  "Age 40–44",
  "Age 45–49",
  "Age 50–54",
  "Age 55–59",

  "  ",

  # Ethnicity
  "White (ref)",
  "Asian",
  "Black",
  "Mixed ethnicity",
  "Other ethnicity",
  "   ",

  # IMD
  "IMD 1 (ref)",
  "IMD 2",
  "IMD 3",
  "IMD 4",
  "IMD 5",
  "    ",

  # Region
  "East Midlands (ref)",
  "East of England",
  "London",
  "North East",
  "North West",
  "South East",
  "South West",
  "West Midlands",
  "Yorkshire and The Humber"
)

# 6. Build adult panels using your existing forest_panel() function
adult_panels <- lapply(
  seq_along(adult_years),
  function(i)
    forest_panel(
      df           = lg_adult_plot,
      year         = adult_years[i],
      term_levels  = adult_terms,
      model_prefix = "lg_adult_",
      show_y       = (i == 1)
    )
)

# 7. Combine panels + add Odds Ratio label
p_adult_combined <- wrap_plots(adult_panels, nrow = 1)

xlabel_adult <- ggplot() +
  theme_void() +
  labs(x = "Odds Ratio") +
  theme(
    axis.title.x = element_text(
      size   = 9,
      margin = margin(t = -5)
    )
  )

p_adult_combined <- wrap_plots(
  p_adult_combined,
  xlabel_adult,
  ncol = 1,
  heights = c(14, 0.1)
)

# 8. Save plot
ggsave(
  filename = "P:/Precision Health Programme/JW Research/DEEP study/GP Dental/Analysis_results_CG/tables_figures/lancet scaled/formatted/forest_adult_COVID.pdf",
  plot     = patchwork::patchworkGrob(p_adult_combined),
  width    = 12,
  height   = 6
)

```


## Step 8: older adult COVID forest (UPDATED)
```{r}

## ------------------------------------------------------------
## older_adult: Create forest plot using existing forest_panel()
## ------------------------------------------------------------

# 1. Years and model names
older_adult_years  <- 2019:2023
older_adult_models <- paste0("lg_older_adult_", older_adult_years)


# 2. Clean term labels for older_adults
#    NO unknown IMD / ethnicity
older_adult_main <- lg_older_adult_results %>%
  filter(model %in% older_adult_models) %>%
  mutate(
    term_label = case_when(

      # Sex
      term == "genderFemale" ~ "Female",

      # IMD
      term == "imd_quintile2" ~ "IMD 2",
      term == "imd_quintile3" ~ "IMD 3",
      term == "imd_quintile4" ~ "IMD 4",
      term == "imd_quintile5" ~ "IMD 5",

      # Age bands
      term == "age_band65-69" ~ "Age 65–69",
      term == "age_band70-74" ~ "Age 70–74",
      term == "age_band75-79" ~ "Age 75–79",
      term == "age_band80-84" ~ "Age 80–84",
      term == "age_band85-89" ~ "Age 85–89",
      term == "age_band90+"   ~ "Age 90+",

      # Ethnicity
      term == "final_ethnicity_groupsAsian or Asian British" ~
        "Asian",

      term == "final_ethnicity_groupsBlack, Black British, Caribbean or African" ~
        "Black",

      term == "final_ethnicity_groupsMixed or multiple ethnic groups" ~
        "Mixed ethnicity",

      term == "final_ethnicity_groupsOther ethnic group" ~
        "Other ethnicity",

      # Region
      term == "regionEast of England" ~
        "East of England",

      term == "regionLondon" ~
        "London",

      term == "regionNorth East" ~
        "North East",

      term == "regionNorth West" ~
        "North West",

      term == "regionSouth East" ~
        "South East",

      term == "regionSouth West" ~
        "South West",

      term == "regionWest Midlands" ~
        "West Midlands",

      term == "regionYorkshire and The Humber" ~
        "Yorkshire and The Humber",

      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(term_label))


# 3. Reference rows
#    Age 60–64 is the age reference category
older_adult_ref_rows <- expand.grid(
  model = older_adult_models,
  term_label = c(
    "Male (ref)",
    "Age 60–64 (ref)",
    "White (ref)",
    "IMD 1 (ref)",
    "East Midlands (ref)"
  ),
  stringsAsFactors = FALSE
) %>%
  mutate(
    term    = paste0("ref_", row_number()),
    OR      = 1,
    CI_low  = 1,
    CI_high = 1
  )


# 4. Combine model estimates and reference rows
lg_older_adult_plot <- bind_rows(
  older_adult_main,
  older_adult_ref_rows
)


# 5. Term ordering with spacers
older_adult_terms <- c(

  # Sex
  "Male (ref)",
  "Female",
  " ",

  # Age
  "Age 60–64 (ref)",
  "Age 65–69",
  "Age 70–74",
  "Age 75–79",
  "Age 80–84",
  "Age 85–89",
  "Age 90+",
  "  ",

  # Ethnicity
  "White (ref)",
  "Asian",
  "Black",
  "Mixed ethnicity",
  "Other ethnicity",
  "   ",

  # IMD
  "IMD 1 (ref)",
  "IMD 2",
  "IMD 3",
  "IMD 4",
  "IMD 5",
  "    ",

  # Region
  "East Midlands (ref)",
  "East of England",
  "London",
  "North East",
  "North West",
  "South East",
  "South West",
  "West Midlands",
  "Yorkshire and The Humber"
)


# 6. Build one forest panel for each year
older_adult_panels <- lapply(
  seq_along(older_adult_years),
  function(i) {

    forest_panel(
      df           = lg_older_adult_plot,
      year         = older_adult_years[i],
      term_levels  = older_adult_terms,
      model_prefix = "lg_older_adult_",
      show_y       = (i == 1)
    )

  }
)


# 7. Combine panels
p_older_adult_combined <- wrap_plots(
  older_adult_panels,
  nrow = 1
)


# Add common x-axis label
xlabel_older_adult <- ggplot() +
  theme_void() +
  labs(x = "Odds Ratio") +
  theme(
    axis.title.x = element_text(
      size   = 9,
      margin = margin(t = -5)
    )
  )


p_older_adult_combined <- wrap_plots(
  p_older_adult_combined,
  xlabel_older_adult,
  ncol    = 1,
  heights = c(14, 0.1)
)


# 8. Save plot
ggsave(
  filename = paste0(
    "P:/Precision Health Programme/JW Research/DEEP study/",
    "GP Dental/Analysis_results_CG/tables_figures/",
    "lancet scaled/formatted/",
    "forest_older_adult_COVID.pdf"
  ),
  plot   = patchwork::patchworkGrob(p_older_adult_combined),
  width  = 12,
  height = 6
)
```
## Step 9: patch forest plots
```{r}
library(ggplot2)
library(patchwork)

# 1) Wrap each "combined" forest so it behaves like ONE panel
children_one <- wrap_elements(p_children_combined)
adults_one   <- wrap_elements(p_adult_combined)
older_adults_one   <- wrap_elements(p_older_adult_combined)

# 2) Make label rows
lab_A <- ggplot() +
  annotate("text", x = 0, y = 0, label = "(A) Children",
           hjust = 0, fontface = "bold", size = 5) +
  theme_void() +
  coord_cartesian(xlim = c(0, 1), ylim = c(-1, 1), clip = "off")

lab_B <- ggplot() +
  annotate("text", x = 0, y = 0, label = "(B) Adults",
           hjust = 0, fontface = "bold", size = 5) +
  theme_void() +
  coord_cartesian(xlim = c(0, 1), ylim = c(-1, 1), clip = "off")

lab_C <- ggplot() +
  annotate("text", x = 0, y = 0, label = "(C) Older adults",
           hjust = 0, fontface = "bold", size = 5) +
  theme_void() +
  coord_cartesian(xlim = c(0, 1), ylim = c(-1, 1), clip = "off")

# 3) Build each block (label + plot), then wrap the whole block too
forest_children_named <- wrap_elements(
  lab_A / children_one + plot_layout(ncol = 1, heights = c(0.08, 1))
)

forest_adults_named <- wrap_elements(
  lab_B / adults_one + plot_layout(ncol = 1, heights = c(0.08, 1))
)

forest_older_adults_named <- wrap_elements(
  lab_C / older_adults_one + plot_layout(ncol = 1, heights = c(0.08, 1))
)

# 4) Stack children (top) and adults (bottom)
patched_forest <- forest_children_named / forest_adults_named +
  plot_layout(ncol = 1, heights = c(1, 1.2))


ggsave(
  filename  = paste0(
    "P:/Precision Health Programme/JW Research/DEEP study/",
    "GP Dental/Analysis_results_CG/tables_figures/",
    "lancet scaled/formatted/",
    "Figure 4AB.pdf"
  ),
  plot      = patched_forest,
  width     = 16,
  height    = 12,
  units     = "in",
  dpi       = 1200,
  limitsize = FALSE
)

ggsave(
  filename  = paste0(
    "P:/Precision Health Programme/JW Research/DEEP study/",
    "GP Dental/Analysis_results_CG/tables_figures/",
    "lancet scaled/formatted/",
    "Figure 4C.pdf"
  ),
  plot      = forest_older_adults_named,
  width     = 16,
  height    = 7,
  units     = "in",
  dpi       = 1200,
  limitsize = FALSE
)
```


## forest plot for supplement 
## Step 10: adult forest (UPDATED)
```{r}



## ------------------------------------------------------------
## Adult: Create forest plot using existing forest_panel()
## ------------------------------------------------------------

# 1. Years and model names
adult_years  <- c(1998, 2003, 2008, 2013, 2018)
adult_models <- paste0("lg_adult_", adult_years)

# 2. Clean term labels for adults (NO unknown IMD/ethnicity)
adult_main <- lg_adult_results %>%
  filter(model %in% adult_models) %>%
  mutate(
    term_label = case_when(
      # Sex
      term == "genderFemale" ~ "Female",

      # IMD (removed Unknown)
      term == "imd_quintile2" ~ "IMD 2",
      term == "imd_quintile3" ~ "IMD 3",
      term == "imd_quintile4" ~ "IMD 4",
      term == "imd_quintile5" ~ "IMD 5",

      # Adult age bands (non-reference levels)
      term == "age_band25-29" ~ "Age 25–29",
      term == "age_band30-34" ~ "Age 30–34",
      term == "age_band35-39" ~ "Age 35–39",
      term == "age_band40-44" ~ "Age 40–44",
      term == "age_band45-49" ~ "Age 45–49",
      term == "age_band50-54" ~ "Age 50–54",
      term == "age_band55-59" ~ "Age 55–59",


      # Ethnicity (removed Not specified)
      term == "final_ethnicity_groupsAsian or Asian British"                     ~ "Asian",
      term == "final_ethnicity_groupsBlack, Black British, Caribbean or African" ~ "Black",
      term == "final_ethnicity_groupsMixed or multiple ethnic groups"            ~ "Mixed ethnicity",
      term == "final_ethnicity_groupsOther ethnic group"                         ~ "Other ethnicity",

      # Regions
      term == "regionEast of England"          ~ "East of England",
      term == "regionLondon"                   ~ "London",
      term == "regionNorth East"               ~ "North East",
      term == "regionNorth West"               ~ "North West",
      term == "regionSouth East"               ~ "South East",
      term == "regionSouth West"               ~ "South West",
      term == "regionWest Midlands"            ~ "West Midlands",
      term == "regionYorkshire and The Humber" ~ "Yorkshire and The Humber",

      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(term_label))

# 3. Reference rows for adults (Age 20–24 is the reference)
adult_ref_rows <- expand.grid(
  model = adult_models,
  term_label = c(
    "Male (ref)",
    "Age 20–24 (ref)",
    "White (ref)",
    "IMD 1 (ref)",
    "East Midlands (ref)"
  ),
  stringsAsFactors = FALSE
) %>%
  mutate(
    term    = paste0("ref_", row_number()),
    OR      = 1,
    CI_low  = 1,
    CI_high = 1
  )

# 4. Combine adult dataset
lg_adult_plot <- bind_rows(adult_main, adult_ref_rows)

# 5. Term ordering with spacers (removed unknown IMD/ethnicity)
adult_terms <- c(
  # Sex
  "Male (ref)",
  "Female",
  " ",

  # Age
  "Age 20–24 (ref)",
  "Age 25–29",
  "Age 30–34",
  "Age 35–39",
  "Age 40–44",
  "Age 45–49",
  "Age 50–54",
  "Age 55–59",

  "  ",

  # Ethnicity
  "White (ref)",
  "Asian",
  "Black",
  "Mixed ethnicity",
  "Other ethnicity",
  "   ",

  # IMD
  "IMD 1 (ref)",
  "IMD 2",
  "IMD 3",
  "IMD 4",
  "IMD 5",
  "    ",

  # Region
  "East Midlands (ref)",
  "East of England",
  "London",
  "North East",
  "North West",
  "South East",
  "South West",
  "West Midlands",
  "Yorkshire and The Humber"
)

# 6. Build adult panels using your existing forest_panel() function
adult_panels <- lapply(
  seq_along(adult_years),
  function(i)
    forest_panel(
      df           = lg_adult_plot,
      year         = adult_years[i],
      term_levels  = adult_terms,
      model_prefix = "lg_adult_",
      show_y       = (i == 1)
    )
)

# 7. Combine panels + add Odds Ratio label
p_adult_combined <- wrap_plots(adult_panels, nrow = 1)

xlabel_adult <- ggplot() +
  theme_void() +
  labs(x = "Odds Ratio") +
  theme(
    axis.title.x = element_text(
      size   = 9,
      margin = margin(t = -5)
    )
  )

p_adult_combined <- wrap_plots(
  p_adult_combined,
  xlabel_adult,
  ncol = 1,
  heights = c(14, 0.1)
)

# 8. Save plot
ggsave(
  filename = paste0(
    "P:/Precision Health Programme/JW Research/DEEP study/",
    "GP Dental/Analysis_results_CG/tables_figures/",
    "lancet scaled/formatted/",
    "Figure S3B.pdf"
  ),
  plot     = p_adult_combined,
  width    = 12,
  height   = 8
)

```

## Step 11: older adult forest (UPDATED)
```{r}



## ------------------------------------------------------------
## older_adult: Create forest plot using existing forest_panel()
## ------------------------------------------------------------

# 1. Years and model names
older_adult_years  <- c(1998, 2003, 2008, 2013, 2018)
older_adult_models <- paste0("lg_older_adult_", older_adult_years)

# 2. Clean term labels for older_adults (NO unknown IMD/ethnicity)
older_adult_main <- lg_older_adult_results %>%
  filter(model %in% older_adult_models) %>%
  mutate(
    term_label = case_when(
      # Sex
      term == "genderFemale" ~ "Female",

      # IMD (removed Unknown)
      term == "imd_quintile2" ~ "IMD 2",
      term == "imd_quintile3" ~ "IMD 3",
      term == "imd_quintile4" ~ "IMD 4",
      term == "imd_quintile5" ~ "IMD 5",

      # older_adult age bands (non-reference levels)

      term == "age_band65-69" ~ "Age 65–69",
      term == "age_band70-74" ~ "Age 70–74",
      term == "age_band75-79" ~ "Age 75–79",
      term == "age_band80-84" ~ "Age 80–84",
      term == "age_band85-89" ~ "Age 85–89",
      term == "age_band90+"   ~ "Age 90+",

      # Ethnicity (removed Not specified)
      term == "final_ethnicity_groupsAsian or Asian British"                     ~ "Asian",
      term == "final_ethnicity_groupsBlack, Black British, Caribbean or African" ~ "Black",
      term == "final_ethnicity_groupsMixed or multiple ethnic groups"            ~ "Mixed ethnicity",
      term == "final_ethnicity_groupsOther ethnic group"                         ~ "Other ethnicity",

      # Regions
      term == "regionEast of England"          ~ "East of England",
      term == "regionLondon"                   ~ "London",
      term == "regionNorth East"               ~ "North East",
      term == "regionNorth West"               ~ "North West",
      term == "regionSouth East"               ~ "South East",
      term == "regionSouth West"               ~ "South West",
      term == "regionWest Midlands"            ~ "West Midlands",
      term == "regionYorkshire and The Humber" ~ "Yorkshire and The Humber",

      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(term_label))

# 3. Reference rows for older_adults (Age 20–24 is the reference)
older_adult_ref_rows <- expand.grid(
  model = older_adult_models,
  term_label = c(
    "Male (ref)",
    "Age 60–64 (ref)",
    "White (ref)",
    "IMD 1 (ref)",
    "East Midlands (ref)"
  ),
  stringsAsFactors = FALSE
) %>%
  mutate(
    term    = paste0("ref_", row_number()),
    OR      = 1,
    CI_low  = 1,
    CI_high = 1
  )

# 4. Combine older_adult dataset
lg_older_adult_plot <- bind_rows(older_adult_main, older_adult_ref_rows)

# 5. Term ordering with spacers (removed unknown IMD/ethnicity)
older_adult_terms <- c(
  # Sex
  "Male (ref)",
  "Female",
  " ",

  # Age
  "Age 60–64 (ref)",
  "Age 65–69",
  "Age 70–74",
  "Age 75–79",
  "Age 80–84",
  "Age 85–89",
  "Age 90+",
  "  ",

  # Ethnicity
  "White (ref)",
  "Asian",
  "Black",
  "Mixed ethnicity",
  "Other ethnicity",
  "   ",

  # IMD
  "IMD 1 (ref)",
  "IMD 2",
  "IMD 3",
  "IMD 4",
  "IMD 5",
  "    ",

  # Region
  "East Midlands (ref)",
  "East of England",
  "London",
  "North East",
  "North West",
  "South East",
  "South West",
  "West Midlands",
  "Yorkshire and The Humber"
)

# 6. Build older_adult panels using your existing forest_panel() function
older_adult_panels <- lapply(
  seq_along(older_adult_years),
  function(i)
    forest_panel(
      df           = lg_older_adult_plot,
      year         = older_adult_years[i],
      term_levels  = older_adult_terms,
      model_prefix = "lg_older_adult_",
      show_y       = (i == 1)
    )
)

# 7. Combine panels + add Odds Ratio label
p_older_adult_combined <- wrap_plots(older_adult_panels, nrow = 1)

xlabel_older_adult <- ggplot() +
  theme_void() +
  labs(x = "Odds Ratio") +
  theme(
    axis.title.x = element_text(
      size   = 9,
      margin = margin(t = -5)
    )
  )

p_older_adult_combined <- wrap_plots(
  p_older_adult_combined,
  xlabel_older_adult,
  ncol = 1,
  heights = c(14, 0.1)
)

# 8. Save plot
ggsave(
  filename  = paste0(
    "P:/Precision Health Programme/JW Research/DEEP study/",
    "GP Dental/Analysis_results_CG/tables_figures/",
    "lancet scaled/formatted/",
    "Figure S3C.pdf"
  ),
  plot     = p_adult_combined,
  width    = 12,
  height   = 8
)

```



## Step 12: patch forest plots
```{r}
library(ggplot2)
library(patchwork)

# 1) Wrap each "combined" forest so it behaves like ONE panel
children_one <- wrap_elements(p_children_combined)
adults_one   <- wrap_elements(p_adult_combined)
older_adults_one   <- wrap_elements(p_older_adult_combined)

# 2) Make label rows
lab_A <- ggplot() +
  annotate("text", x = 0, y = 0, label = "(A) Children",
           hjust = 0, fontface = "bold", size = 5) +
  theme_void() +
  coord_cartesian(xlim = c(0, 1), ylim = c(-1, 1), clip = "off")

lab_B <- ggplot() +
  annotate("text", x = 0, y = 0, label = "(B) Adults",
           hjust = 0, fontface = "bold", size = 5) +
  theme_void() +
  coord_cartesian(xlim = c(0, 1), ylim = c(-1, 1), clip = "off")

lab_C <- ggplot() +
  annotate("text", x = 0, y = 0, label = "(C) Older adults",
           hjust = 0, fontface = "bold", size = 5) +
  theme_void() +
  coord_cartesian(xlim = c(0, 1), ylim = c(-1, 1), clip = "off")

# 3) Build each block (label + plot), then wrap the whole block too
forest_children_named <- wrap_elements(
  lab_A / children_one + plot_layout(ncol = 1, heights = c(0.08, 1))
)

forest_adults_named <- wrap_elements(
  lab_B / adults_one + plot_layout(ncol = 1, heights = c(0.08, 1))
)

forest_older_adults_named <- wrap_elements(
  lab_C / older_adults_one + plot_layout(ncol = 1, heights = c(0.08, 1))
)

# 4) Stack children (top) and adults (bottom)
patched_forest <- forest_children_named / forest_adults_named +
  plot_layout(ncol = 1, heights = c(1, 1.2))


ggsave(
  filename  = paste0(
    "P:/Precision Health Programme/JW Research/DEEP study/",
    "GP Dental/Analysis_results_CG/tables_figures/",
    "lancet scaled/formatted/",
    "Figure S3AB.pdf"
  ),
  plot      = patched_forest,
  width     = 16,
  height    = 12,
  units     = "in",
  dpi       = 1200,
  limitsize = FALSE
)

ggsave(
  filename  = paste0(
    "P:/Precision Health Programme/JW Research/DEEP study/",
    "GP Dental/Analysis_results_CG/tables_figures/",
    "lancet scaled/formatted/",
    "Figure S3C.pdf"
  ),
  plot      = forest_older_adults_named,
  width     = 16,
  height    = 7,
  units     = "in",
  dpi       = 1200,
  limitsize = FALSE
)
```


# Section 7: average repeated time measure
a patient have n times of repeated visit for a same problem will be include n times (record level average)
```{r}
library(dplyr)

## -------------------------------------------------
## record-level data using Dental Problem Category
## -------------------------------------------------

rec_use <- dental_records_long_final %>%
  filter(
    !is.na(patid),
    !is.na(obsdate),
    !is.na(`Dental Problem Category`)
  ) %>%
  mutate(
    obsdate = as.Date(obsdate)
  ) %>%
  arrange(patid, `Dental Problem Category`, obsdate)


## =================================================
## 1. First-to-second repeated visit interval only
## One interval per patient per Dental Problem Category
## If one patient has repeated visits for multiple conditions,
## each condition is counted separately.
## =================================================

repeat_interval_first_to_second <- rec_use %>%
  arrange(patid, `Dental Problem Category`, obsdate) %>%
  group_by(patid, `Dental Problem Category`) %>%
  summarise(
    n_problem_visits = n(),
    first_visit_date = first(obsdate),
    second_visit_date = nth(obsdate, 2),
    days_first_to_second_visit = as.numeric(second_visit_date - first_visit_date),
    .groups = "drop"
  ) %>%
  filter(n_problem_visits >= 2)


average_first_to_second <- repeat_interval_first_to_second %>%
  summarise(
    n_intervals = n(),
    mean_days_first_to_second = mean(days_first_to_second_visit, na.rm = TRUE),
    sd_days_first_to_second = sd(days_first_to_second_visit, na.rm = TRUE),
    median_days_first_to_second = median(days_first_to_second_visit, na.rm = TRUE),
    q1_days_first_to_second = quantile(days_first_to_second_visit, 0.25, na.rm = TRUE),
    q3_days_first_to_second = quantile(days_first_to_second_visit, 0.75, na.rm = TRUE),
    min_days_first_to_second = min(days_first_to_second_visit, na.rm = TRUE),
    max_days_first_to_second = max(days_first_to_second_visit, na.rm = TRUE)
  )


## =================================================
## 2. All consecutive repeated visit intervals
## One patient and one condition can contribute multiple intervals:
## second - first, third - second, fourth - third, etc.
## =================================================

repeat_interval_all_consecutive <- rec_use %>%
  arrange(patid, `Dental Problem Category`, obsdate) %>%
  group_by(patid, `Dental Problem Category`) %>%
  mutate(
    visit_order_for_same_condition = row_number(),
    previous_visit_date = lag(obsdate),
    days_between_consecutive_visits = as.numeric(obsdate - previous_visit_date)
  ) %>%
  ungroup() %>%
  filter(!is.na(days_between_consecutive_visits))


average_all_consecutive <- repeat_interval_all_consecutive %>%
  summarise(
    n_intervals = n(),
    mean_days_all_consecutive = mean(days_between_consecutive_visits, na.rm = TRUE),
    sd_days_all_consecutive = sd(days_between_consecutive_visits, na.rm = TRUE),
    median_days_all_consecutive = median(days_between_consecutive_visits, na.rm = TRUE),
    q1_days_all_consecutive = quantile(days_between_consecutive_visits, 0.25, na.rm = TRUE),
    q3_days_all_consecutive = quantile(days_between_consecutive_visits, 0.75, na.rm = TRUE),
    min_days_all_consecutive = min(days_between_consecutive_visits, na.rm = TRUE),
    max_days_all_consecutive = max(days_between_consecutive_visits, na.rm = TRUE)
  )


## =================================================
## 3. Combine the two summaries into one table
## =================================================

average_repeat_visit_time_summary <- bind_rows(
  average_first_to_second %>%
    mutate(interval_definition = "First to second visit only") %>%
    rename(
      mean_days = mean_days_first_to_second,
      sd_days = sd_days_first_to_second,
      median_days = median_days_first_to_second,
      q1_days = q1_days_first_to_second,
      q3_days = q3_days_first_to_second,
      min_days = min_days_first_to_second,
      max_days = max_days_first_to_second
    ),
  
  average_all_consecutive %>%
    mutate(interval_definition = "Interval to next visit for a same dental problem") %>%
    rename(
      mean_days = mean_days_all_consecutive,
      sd_days = sd_days_all_consecutive,
      median_days = median_days_all_consecutive,
      q1_days = q1_days_all_consecutive,
      q3_days = q3_days_all_consecutive,
      min_days = min_days_all_consecutive,
      max_days = max_days_all_consecutive
    )
) %>%
  dplyr::select(
    interval_definition,
    n_intervals,
    mean_days,
    sd_days,
    median_days,
    q1_days,
    q3_days,
    min_days,
    max_days
  ) %>%
  mutate(
    mean_sd_days = paste0(
      round(mean_days, 1),
      " (",
      round(sd_days, 1),
      ")"
    ),
    median_iqr_days = paste0(
      round(median_days, 1),
      " (",
      round(q1_days, 1),
      "-",
      round(q3_days, 1),
      ")"
    )
  )



## =================================================
## 4. Summary by Dental Problem Category
## =================================================

average_first_to_second_by_problem <- repeat_interval_first_to_second %>%
  group_by(`Dental Problem Category`) %>%
  summarise(
    n_intervals = n(),
    mean_days = mean(days_first_to_second_visit, na.rm = TRUE),
    sd_days = sd(days_first_to_second_visit, na.rm = TRUE),
    median_days = median(days_first_to_second_visit, na.rm = TRUE),
    q1_days = quantile(days_first_to_second_visit, 0.25, na.rm = TRUE),
    q3_days = quantile(days_first_to_second_visit, 0.75, na.rm = TRUE),
    min_days = min(days_first_to_second_visit, na.rm = TRUE),
    max_days = max(days_first_to_second_visit, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    interval_definition = "First to second visit only",
    mean_sd_days = paste0(
      round(mean_days, 1),
      " (",
      round(sd_days, 1),
      ")"
    ),
    median_iqr_days = paste0(
      round(median_days, 1),
      " (",
      round(q1_days, 1),
      "-",
      round(q3_days, 1),
      ")"
    )
  )


average_all_consecutive_by_problem <- repeat_interval_all_consecutive %>%
  group_by(`Dental Problem Category`) %>%
  summarise(
    n_intervals = n(),
    mean_days = mean(days_between_consecutive_visits, na.rm = TRUE),
    sd_days = sd(days_between_consecutive_visits, na.rm = TRUE),
    median_days = median(days_between_consecutive_visits, na.rm = TRUE),
    q1_days = quantile(days_between_consecutive_visits, 0.25, na.rm = TRUE),
    q3_days = quantile(days_between_consecutive_visits, 0.75, na.rm = TRUE),
    min_days = min(days_between_consecutive_visits, na.rm = TRUE),
    max_days = max(days_between_consecutive_visits, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    interval_definition = "Interval to next visit for a same dental problem",
    mean_sd_days = paste0(
      round(mean_days, 1),
      " (",
      round(sd_days, 1),
      ")"
    ),
    median_iqr_days = paste0(
      round(median_days, 1),
      " (",
      round(q1_days, 1),
      "-",
      round(q3_days, 1),
      ")"
    )
  )


average_repeat_visit_time_by_problem <- bind_rows(
  average_first_to_second_by_problem,
  average_all_consecutive_by_problem
) %>%
  dplyr::select(
    interval_definition,
    `Dental Problem Category`,
    n_intervals,
    mean_days,
    sd_days,
    median_days,
    q1_days,
    q3_days,
    min_days,
    max_days,
    mean_sd_days,
    median_iqr_days
  )



```

### Step 1: by condition regarless nex condition
```{r}
library(dplyr)

## =================================================
## Time from each condition visit to the next visit
## regardless of the next visit condition
##
## Example:
## Perio -> Caries -> Perio -> Toothache
##
## Intervals counted as:
## Perio   -> Caries     counted under Perio
## Caries  -> Perio      counted under Caries
## Perio   -> Toothache  counted under Perio
## Toothache has no next visit, so not counted
## =================================================

condition_to_next_any_visit_records <- rec_use %>%
  arrange(patid, obsdate) %>%
  group_by(patid) %>%
  mutate(
    current_visit_date = obsdate,
    current_problem = `Dental Problem Category`,
    next_visit_date_any_condition = lead(obsdate),
    next_problem_any_condition = lead(`Dental Problem Category`),
    days_current_condition_to_next_any_visit = as.numeric(
      next_visit_date_any_condition - current_visit_date
    )
  ) %>%
  ungroup() %>%
  filter(!is.na(days_current_condition_to_next_any_visit))


## =================================================
## Summary by current Dental Problem Category
## =================================================

condition_to_next_any_visit_by_problem <- condition_to_next_any_visit_records %>%
  group_by(current_problem) %>%
  summarise(
    n_intervals = n(),
    mean_days = mean(days_current_condition_to_next_any_visit, na.rm = TRUE),
    sd_days = sd(days_current_condition_to_next_any_visit, na.rm = TRUE),
    median_days = median(days_current_condition_to_next_any_visit, na.rm = TRUE),
    q1_days = quantile(days_current_condition_to_next_any_visit, 0.25, na.rm = TRUE),
    q3_days = quantile(days_current_condition_to_next_any_visit, 0.75, na.rm = TRUE),
    min_days = min(days_current_condition_to_next_any_visit, na.rm = TRUE),
    max_days = max(days_current_condition_to_next_any_visit, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(
    `Dental Problem Category` = current_problem
  ) %>%
  mutate(
    interval_definition = "Interval to next visit for dental problems",
    mean_sd_days = paste0(
      round(mean_days, 1),
      " (",
      round(sd_days, 1),
      ")"
    ),
    median_iqr_days = paste0(
      round(median_days, 1),
      " (",
      round(q1_days, 1),
      "-",
      round(q3_days, 1),
      ")"
    )
  ) %>%
  dplyr::select(
    interval_definition,
    `Dental Problem Category`,
    n_intervals,
    mean_days,
    sd_days,
    median_days,
    q1_days,
    q3_days,
    min_days,
    max_days,
    mean_sd_days,
    median_iqr_days
  )



## =================================================
## Optional: combine overall and by-problem summary
## =================================================

condition_to_next_any_visit_summary <- 
  condition_to_next_any_visit_by_problem%>%
  mutate(
    `Dental Problem Category` = case_when(
      `Dental Problem Category` == "Infection" ~ "Dental abscess",
      TRUE ~ `Dental Problem Category`
    )
  ) %>%
  dplyr::select(
    `Dental Problem Category`,
    interval_definition,
    n_intervals,
    mean_days,
    sd_days,
    median_days,
    q1_days,
    q3_days,
    min_days,
    max_days,
    mean_sd_days,
    median_iqr_days
  )



```

## Step 2: repeated visit regardless condition
```{r}
library(dplyr)

## -------------------------------------------------
## New calculation only:
## interval between consecutive visits regardless of condition
## This does NOT overwrite your previous data frames
## -------------------------------------------------

any_problem_visit_gap_records <- dental_records_long_final %>%
  filter(
    !is.na(patid),
    !is.na(obsdate)
  ) %>%
  mutate(
    obsdate_clean_for_any_problem_gap = as.Date(obsdate)
  ) %>%
  arrange(patid, obsdate_clean_for_any_problem_gap) %>%
  group_by(patid) %>%
  mutate(
    any_problem_visit_order = row_number(),
    previous_any_problem_visit_date = lag(obsdate_clean_for_any_problem_gap),
    days_between_any_problem_visits = as.numeric(
      obsdate_clean_for_any_problem_gap - previous_any_problem_visit_date
    )
  ) %>%
  ungroup() %>%
  filter(!is.na(days_between_any_problem_visits))


## -------------------------------------------------
## Summary: mean, SD, median, IQR
## -------------------------------------------------

any_problem_visit_gap_summary <- any_problem_visit_gap_records %>%
  summarise(
    n_intervals = n(),
    mean_days = mean(days_between_any_problem_visits, na.rm = TRUE),
    sd_days = sd(days_between_any_problem_visits, na.rm = TRUE),
    median_days = median(days_between_any_problem_visits, na.rm = TRUE),
    q1_days = quantile(days_between_any_problem_visits, 0.25, na.rm = TRUE),
    q3_days = quantile(days_between_any_problem_visits, 0.75, na.rm = TRUE),
    min_days = min(days_between_any_problem_visits, na.rm = TRUE),
    max_days = max(days_between_any_problem_visits, na.rm = TRUE)
  ) %>%
  mutate(
    interval_definition = "Interval to next visit for dental problems",
    mean_sd_days = paste0(
      round(mean_days, 1),
      " (",
      round(sd_days, 1),
      ")"
    ),
    median_iqr_days = paste0(
      round(median_days, 1),
      " ( ",
      round(q1_days, 1),
      "-",
      round(q3_days, 1),
      ")"
    )
  ) %>%
  dplyr::select(
    interval_definition,
    n_intervals,
    mean_days,
    sd_days,
    median_days,
    q1_days,
    q3_days,
    min_days,
    max_days,
    mean_sd_days,
    median_iqr_days
  )



```

```{r}
#----------------------------------------------------------------
# bind table and correct infection name
#----------------------------------------------------------------
average_repeat_visit_time_summary$`Dental Problem Category`<-"All problems"
any_problem_visit_gap_summary$`Dental Problem Category`<-"All problems"
repeated_time_summary_combined <- bind_rows(average_repeat_visit_time_summary,average_repeat_visit_time_by_problem,any_problem_visit_gap_summary,condition_to_next_any_visit_summary) 

repeated_time_summary_combined <-repeated_time_summary_combined%>%
  mutate(    `Dental Problem Category` = case_when(    
    `Dental Problem Category`=="Infection" ~ "Dental abscess",
    T~    `Dental Problem Category`))%>%
  mutate( `Interval definition` = interval_definition,
          `N repeated records` = n_intervals,
          `Mean (SD)` = mean_sd_days,
          `Median (IQR)` = median_iqr_days)%>%
  dplyr::select(    `Dental Problem Category`, 
                    `N repeated records`,
                    `Interval definition` ,
                    `Median (IQR)`) %>%
  filter(`Interval definition` !="First to second visit only") 

repeated_time_summary_wide <- repeated_time_summary_combined %>%
  dplyr::select(
    `Dental Problem Category`,
    `N repeated records`,
    `Interval definition`,
    `Median (IQR)`
  ) %>%
  distinct() %>%
  pivot_wider(
    names_from = `Interval definition`,
    values_from = c(`N repeated records`, `Median (IQR)`)
  )

write_xlsx(
  repeated_time_summary_wide,
  paste0(tables_figures, "/objective 1/repeated_time_summary_combined.xlsx"
))
```


## Step 3: creat bar plot 
```{r}
library(dplyr)
library(ggplot2)
library(scales)

## -------------------------------------------------
## Prepare plot data
## Group n_visits >= 20 as "20+"
## -------------------------------------------------

visit_count_plot_data <- dental_sample_final %>%
  filter(!is.na(n_visits)) %>%
  mutate(
    n_visits_grouped = case_when(
      n_visits >= 20 ~ "20+",
      TRUE ~ as.character(n_visits)
    ),
    n_visits_grouped = factor(
      n_visits_grouped,
      levels = c(as.character(1:19), "20+")
    )
  ) %>%
  count(n_visits_grouped, name = "n_patients") %>%
  mutate(
    percent = n_patients / sum(n_patients) * 100,
    n_patients_label = scales::comma(n_patients)
  )


## -------------------------------------------------
## Bar chart
## Y-axis = percentage of patients
## Bar label = number of patients
## Horizontal background grid lines only
## -------------------------------------------------

p_visit_count <- ggplot(
  visit_count_plot_data %>% filter(n_visits_grouped != "1"),
  aes(x = n_visits_grouped, y = percent)
) +
  geom_col(
    fill = "#E97F63",
    colour = "#E97F63",
    width = 0.8
  ) +
  geom_text(
    aes(label = n_patients_label),
    vjust = -0.3,
    size = 3
  ) +
  scale_y_continuous(
    labels = function(x) paste0(x, "%"),
    breaks = scales::pretty_breaks(n = 6),
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(
    x = "Number of visits",
    y = "Percentage of patients"
  ) +
  theme_classic() +
  theme(
    panel.grid.major.y = element_line(colour = "grey85", linewidth = 0.3),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

ggsave(
  filename = paste0(tables_figures, "/objective 1/Distribution of number of visits per patient.jpeg"),
  plot = p_visit_count,
  width = 8,
  height = 5,
  dpi = 600
)
```


##Step 4: calculate number of records with 0 interval (at the same date)
```{r}
same_date_count <- dental_records_long_final %>%
  group_by(patid, obsdate) %>%
  summarise(
    n_records = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(n_records))

same_date_multiple <- dental_records_long_final %>%
  count(patid, obsdate, name = "n_records") %>%
  filter(n_records > 1) %>%
  arrange(desc(n_records))


same_date_multiple %>%
  count(n_records)

(sum(same_date_multiple$n_records)/nrow(dental_records_long_final))*100
```


#### update table 1 stratefied by cases 

```{r}

source("tableone_create_export_v1.R")
# Categorical variables
#---------------------------------------------------------
# Create the analysis dataset
#---------------------------------------------------------
table1_sample_data <- sample_final %>%
  left_join(
    dental_sample_final %>%
      dplyr::select(patid, first_dcat),
    by = "patid"
  )

#---------------------------------------------------------
# Run Table 1 stratified by dental-visit status
#---------------------------------------------------------
table1_sample_events <- table1(
  data = table1_sample_data,
  
  vars = c(
    "age_first_dental",
    "gender",
    "final_ethnicity_groups",
    "final_ethnicity",
    "imd_quintile",
    "dent_visit_count2",
    "first_dcat",
    "ever_had_repeated_visit_same_problem",
    "repeated_visit_for_first_problem"
  ),
  
  strata = "had_dental_visit",
  
  factorVars = factorVars,
  
  addOverall = TRUE,
  
  directory = paste0(
    "P:/Precision Health Programme/JW Research/DEEP study/",
    "GP Dental/Analysis_results_CG/tables_figures/",
    "lancet scaled/formatted/"
  ),
  
  filename = "Table 1 stratified by dental visit status.xlsx"
)


```



