---
title: "Supplement codes 2: Dental-related consultations in general practice in England: a population-based retrospective study of national trends and inequalities"
author: "CG"
output: html_document
---
# restart the working environment 
```{r}
rm(list = ls())
```


# library
```{r setup, include=FALSE}
library(readxl)
library(writexl)
library(tidyverse)
library(tidyr)
library(popEpi)
library(epitools)

#install.packages(c("sf","ggplot2","dplyr","viridis", "rgeos", "rgdal", "maptool"))
library(sf)
library(ggplot2)
library(dplyr)
library(viridis)
library(epiR)
library(PHEindicatormethods)
library(broom)
library(dplyr)
library(purrr)
library(tidyverse)
library(lubridate)
library(MASS)
library(sandwich)
library(lmtest)
library(segmented)
# install.packages("forecast")
library(forecast)
library(MASS)     # glm.nb
library(splines)  # ns()

library(dplyr)
library(ggplot2)
library(patchwork)
library(grid)   # for arrow()
#library(rgeos)
#library(rgdal)
#library(maptools)
library(forcats)
library(cowplot)
library(pdftools)
library(patchwork)
library(ggplot2)




# frequent used path
tables_figures <- "Analysis_results_CG/tables_figures/lancet scaled"
processed_data <- "Analysis_results_CG/processed_data_v2"
raw_data <- "Analysis_results_CG/raw_data"
final_curated_data <- "Analysis_results_CG/processed_data_v2/final_curated_data"

code_lists <- "Analysis_results_CG/code_list"


# 2. define study period
study_start <- as.Date("1998-01-01")

study_end <- as.Date("2025-12-31")


# 3. load function
source_folder <- function(folder) {
  files <- sort(list.files(folder, pattern = "\\.R$", full.names = TRUE))
  
  for (f in files) {
    message("Sourcing: ", basename(f))
    try(source(f, local = .GlobalEnv), silent = FALSE)
  }
}

folder <- "ApendixA_supp_functions"

source_folder(folder)

```


#########################################################################
# Section 1: calculate ir scale ratio for all

# Section 1.1: overall ir convert 
## Step 1: weight ir 
```{r}
# 1. create weight variable by n events/ total cases
ir_summary_table_S7$weight <- as.numeric(ir_summary_table_S7$`N events`/ir_summary_table_S7$total_cases)

# 2. use overall weight for those NA 
overall_weight <- as.numeric(ir_summary_table_S7$weight[ir_summary_table_S7$variable == "overall"])
overall_weight

ir_summary_table_S7$weight <- ifelse(is.na(ir_summary_table_S7$weight),overall_weight, ir_summary_table_S7$weight)
table(ir_summary_table_S7$weight == overall_weight, useNA = "ifany")# imd6+ethnicity6+ethnicity14+overall1 =27, correct

# 3: apply weight to ir 1000 and ci
ir_summary_table_scaled <- ir_summary_table_S7 %>%
mutate(`IR per 1000py` = `IR per 1000py` * weight,
`Lower CI` = `Lower CI` * weight,
`upper CI` = `upper CI` * weight
)

```

## step 2: format table to export
```{r}
ir_summary_table_scaled_formatted <- ir_summary_table_scaled %>%
filter(variable != "Ethnicity (14 groups)") %>%
mutate(
Variable = recode(
variable,
overall = "Overall",
Region = "English region",
gender = "Sex",
`Dental problem category` = "Dental problem category",
`Age Band` = "Age",
.default = variable
),
Level = case_when(
variable == "overall" ~ "All participants",
TRUE ~ as.character(level)),
Weight = round(weight, 2),
`IR per 1,000 person-years (95% CI)` = sprintf(
"%.2f (%.2f–%.2f)",
`IR per 1000py`,
`Lower CI`,
`upper CI`)) %>%
dplyr::select(
Variable,
Level,
`Total person years`,
`N events`,
Weight,
`IR per 1,000 person-years (95% CI)`
) %>%
mutate(
Variable = if_else(
duplicated(Variable),
"",Variable))


ir_summary_table_scaled_formatted <- ir_summary_table_scaled_formatted %>%
  mutate(
    `N events` = formatC(`N events`, format = "f", digits = 0, big.mark = ","),
  `Total person years` = formatC(
    `Total person years`,format = "f", digits = 0,big.mark = ","),
    `IR per 1000 PY (95%CI)` = gsub("\\.","·",`IR per 1,000 person-years (95% CI)`)
  )%>%
   dplyr::select(-`IR per 1,000 person-years (95% CI)`,-Weight)
write_xlsx(ir_summary_table_scaled_formatted , paste0(tables_figures, "/formatted/Table S7.Incidece rate of having GP visit for dental problems.xlsx"))
```

# Section 1.2: annual ir convert
## Step 1: combine data
```{r}

# 1. left join these data to create scaled data 
annual_ir_age2_scaled <- annual_ir_age2 %>%
  left_join(annual_ir_age2_overall, by = c("age_band","year"))

annual_ir_gender_scaled <- annual_ir_gender %>%
  left_join(annual_ir_gender_overall, by = c("Sex","year"))

annual_ir_region_scaled <- annual_ir_region %>%
  left_join(annual_ir_region_overall, by = c("Region","year"))

annual_ir_scaled <- annual_summary_table %>%
  left_join(annual_ir_overall, by = c("year"))

annual_ir_dcat_scaled <- annual_summary_dcat %>%
  mutate(`Dental problem categories` = case_when (
    `Dental problem categories` == "Infection" ~"Dental abscess",
    TRUE~ `Dental problem categories`
  ))%>%
  left_join(annual_ir_dcat_overall, by = c("Dental problem categories","year"))


```



## Step 2: weight ir 
```{r}
annual_ir_list_scaled <- list(annual_ir_age2_scaled,annual_ir_gender_scaled, annual_ir_region_scaled,annual_ir_scaled,annual_ir_dcat_scaled)

# 1. get annual weight using actual weight 
annual_ir_list_scaled <- lapply(annual_ir_list_scaled,
function(df) {
  
# 0. create ci 
df <- df %>%
rowwise() %>%
mutate(
tmp = list(epi.conf(cbind(cases, total_person_years),ctype = "inc.rate",method = "exact") * 1000),
lower95 = tmp$lower,
upper95 = tmp$upper
) %>%
dplyr::select(-tmp) %>%
ungroup()
# 1. Create weight
df$weight <- as.numeric(df$`cases` / df$cases_overall)

# 2. Scale IR and CIs
df <- df %>%
mutate(
`ir1000_scaled` = `ir1000` * weight,
 lower95_scaled = lower95*weight,
 upper95_scaled = upper95*weight
)
df})

annual_ir_age2_scaled <- annual_ir_list_scaled[[1]]
annual_ir_gender_scaled <- annual_ir_list_scaled[[2]]
annual_ir_region_scaled <- annual_ir_list_scaled[[3]]
annual_ir_scaled <- annual_ir_list_scaled[[4]]
annual_ir_dcat_scaled <- annual_ir_list_scaled[[5]]

# 3. add annual ir weight to imd and ethnicity 
annual_ir_6eth_scaled <- annual_ir_6eth %>%
  left_join(annual_ir_scaled %>% dplyr::select(year, weight), by = "year")%>%
  rowwise() %>%
mutate(
tmp = list(epi.conf(cbind(cases, total_person_years),ctype = "inc.rate",method = "exact") * 1000),
lower95 = tmp$lower,
upper95 = tmp$upper
) %>%
dplyr::select(-tmp) %>%
ungroup()%>%
  mutate(
`ir1000_scaled` = `ir1000` * weight,
 lower95_scaled = lower95*weight,
 upper95_scaled = upper95*weight
)

annual_ir_imd_scaled <- annual_ir_imd %>%
 left_join(annual_ir_scaled %>% dplyr::select(year, weight), by = "year")%>%
  rowwise() %>%
mutate(
tmp = list(epi.conf(cbind(cases, total_person_years),ctype = "inc.rate",method = "exact") * 1000),
lower95 = tmp$lower,
upper95 = tmp$upper
) %>%
dplyr::select(-tmp) %>%
ungroup()%>%
  mutate(
`ir1000_scaled` = `ir1000` * weight,
 lower95_scaled = lower95*weight,
 upper95_scaled = upper95*weight
)
```


## Step 3: recreate plot

### Step 3.1: update age annual incidence rate 
```{r}
# 5. plot annual age ir
lp_annual_ir_age2 <- make_lineplot(
  data    = annual_ir_age2_scaled %>% mutate(`Age group` =`Age group.x`),
  x       = factor(year),
  y       = ir1000_scaled,
  color   = `Age group`,
  title   = "(B) Incidence rate by age group",
  xlab    = "Year",
  ylab    = "Incidence rate (per 1000 PY)",
  outfile = paste0(tables_figures, "/figure_1a_annual_ir_age2_lineplot_scaled.jpeg")
)

```

### Step 3.2: update sex annual IR plots
```{r}


# 5. plot annual gender ir
lp_annual_ir_sex <- make_lineplot(
  data    = annual_ir_gender_scaled,
  x       = factor(year),
  y       = ir1000_scaled,
  color   = Sex,
  title   = "Incidence rate by sex",
  xlab    = "Year",
  ylab    = "Incidence rate (per 1000 PY)",
  outfile = paste0(tables_figures, "/figure_1a_annual_ir_sex_lineplot_scaled.jpeg")
)


```
### Seto 3.3: update ethnicity plot

```{r}

# 5. plot annual 6eth ir
lp_annual_ir_6eth<- make_lineplot(
  data    = annual_ir_6eth_scaled %>% filter(Ethnicity != "Not specified"),
  x       = factor(year),
  y       = ir1000_scaled,
  color   = Ethnicity,
  title   = "(C) Incidence rate by ethnicity",
  xlab    = "Year",
  ylab    = "Incidence rate (per 1000 PY)",
  outfile = paste0(tables_figures, "/figure_1a_annual_ir_6eth_lineplot_scaled.jpeg")
)

```

### Step 3.4: update Dcat 
```{r}

lp_annual_dcat <- make_lineplot(
  data    = annual_ir_dcat_scaled %>% mutate(`Dental problem` = `Dental problem categories`),
  x       = factor(year),
  y       = ir1000_scaled,
  color =   `Dental problem`,
  title   = "(F) Incidence rate by dental problem",
  xlab    = "Year",
  ylab    = "Incidence rate (per 1000 PY)",
  outfile = paste0(tables_figures, "/figure_1a_dcat_annual_ir_lineplot_scaled.jpeg")
)


```

### Step 3.5: update overall
```{r}
lp_annual_overall  <-make_lineplot(
  data    = annual_ir_scaled,
  x       = factor(year),
  y       = ir1000_scaled,
  title   = "(A) Overall incidence rate",
  xlab    = "Year",
  ylab    = "Incidence rate (per 1000 PY)",
  outfile = paste0(tables_figures, "/figure_1a_annual_ir_lineplot_scaled.jpeg")
)


```


### Step 3.6: updated imd
```{r}
# 5. plot annual imd ir
lp_annual_ir_imd <- make_lineplot(
  data    = annual_ir_imd_scaled %>% filter(IMD != "Unknown"),
  x       = factor(year),
  y       = ir1000_scaled,
  color   = IMD,
  title   = "(D) Incidence rate by IMD",
  xlab    = "Year",
  ylab    = "Incidence rate (per 1000 PY)",
  outfile = paste0(tables_figures, "/figure_1a_annual_ir_imd_lineplot_scaled.jpeg")
)


```


### Step 3.7 updated region
```{r}
# 5. plot annual region ir
lp_annual_ir_region <- make_lineplot(
  data    = annual_ir_region_scaled %>% mutate(`English region` = Region),
  x       = factor(year),
  y       = ir1000_scaled,
  color   = `English region`,
  title   = "(E) Incidence rate by English region",
  xlab    = "Year",
  ylab    = "Incidence rate (per 1000 PY)",
  outfile = paste0(tables_figures, "/figure_1a_annual_ir_region_lineplot_scaled.jpeg")
)


```


### patch the reordered plots

```{r}
library(patchwork)
library(ggplot2)



annual_ir_scaled_filter <- annual_ir_scaled %>%
  mutate(Sex = "Overall") %>%
  dplyr::select(year, Sex, ir1000_scaled)

annual_ir_gender_overall_scaled <- annual_ir_gender_scaled %>%
  dplyr::select(year, Sex, ir1000_scaled) %>%
  bind_rows(annual_ir_scaled_filter)

lp_annual_overall  <-make_lineplot(
  data    = annual_ir_gender_overall_scaled,
  x       = factor(year),
  y       = ir1000_scaled,
  color =  Sex,
  title   = "(A) Overall incidence rate and 
  incidence rate by Sex",
  xlab    = "Year",
  ylab    = "Incidence rate (per 1000 PY)",
  outfile = paste0(tables_figures, "/figure_1a_annual_ir_lineplot_scaled.jpeg")
)



# ---- annual-only objects to avoid overwriting ----
annual_p_list <- list(
  overall = lp_annual_overall,
  eth6    = lp_annual_ir_6eth,
  age2    = lp_annual_ir_age2,
  imd     = lp_annual_ir_imd,
  region  = lp_annual_ir_region,
  dcat    = lp_annual_dcat
)

# ---- Common y-range across all annual plots ----
annual_get_plot_yrange <- function(annual_p) {
  annual_b <- ggplot_build(annual_p)

  annual_yr <- unlist(lapply(annual_b$layout$panel_params, function(annual_pp) {
    if (!is.null(annual_pp$y.range)) annual_pp$y.range else c(NA_real_, NA_real_)
  }))

  if (all(is.na(annual_yr))) {
    annual_yr <- unlist(lapply(annual_b$layout$panel_params, function(annual_pp) {
      if (!is.null(annual_pp$y$range$range)) annual_pp$y$range$range else c(NA_real_, NA_real_)
    }))
  }

  range(annual_yr, na.rm = TRUE)
}

annual_y_ranges <- lapply(annual_p_list, annual_get_plot_yrange)
annual_y_lim <- range(do.call(c, annual_y_ranges), na.rm = TRUE)

# ---- Apply common y-limits (no legend positioning here yet) ----
annual_p_list <- lapply(annual_p_list, function(annual_p) {
  annual_p + coord_cartesian(ylim = annual_y_lim)
})

# ---- Left column: legends on the LEFT; Right column: legends on the RIGHT ----
annual_left_col <-
  (annual_p_list$overall /annual_p_list$eth6  /  annual_p_list$region) &
  theme(legend.position = "left")

annual_right_col <-
  (annual_p_list$age2 / annual_p_list$imd / annual_p_list$dcat) &
  theme(legend.position = "right")

# ---- Combine into 2 columns x 3 rows ----
annual_lplot <- annual_left_col | annual_right_col


ggsave(
  filename = paste0(tables_figures, "/formatted/Figure 3. Annual incidence rate overall and by sociodemographic factors.jpeg"),
  plot     = annual_lplot,
  width    = 14,
  height   = 14,
  units    = "in",
  dpi      = 1200
)
ggsave(
  filename = paste0(tables_figures, "/formatted/Figure 3. Annual incidence rate overall and by sociodemographic factors.pdf"),
  plot     = annual_lplot,
  width    = 14,
  height   = 14,
  units    = "in",
  dpi      = 1200
)
```

## Step 4: extract the maxium IR and minium and their CI 
the format table code has error need to sort 
it doesnot contains correct level... 
```{r}
# ============================================================
# Extract minimum and maximum annual scaled IR
# ============================================================

library(dplyr)
library(purrr)
library(writexl)


# 1. Put all scaled annual IR datasets into a named list
annual_ir_list_scaled <- list(
  ethnicity = annual_ir_6eth_scaled,
  age       = annual_ir_age2_scaled,
  sex       = annual_ir_gender_scaled,
  imd       = annual_ir_imd_scaled,
  region    = annual_ir_region_scaled,
  overall   = annual_ir_scaled,
  dcat      = annual_ir_dcat_scaled
)


# 2. Function to extract the row with:
#    - minimum scaled IR
#    - maximum scaled IR
#    from each dataset
extract_min_max_ir <- function(df, source_name) {
  
  min_row <- df %>%
    filter(!is.na(ir1000_scaled)) %>%
    slice_min(
      order_by = ir1000_scaled,
      n = 1,
      with_ties = FALSE
    ) %>%
    mutate(extreme = "Minimum")
  
  max_row <- df %>%
    filter(!is.na(ir1000_scaled)) %>%
    slice_max(
      order_by = ir1000_scaled,
      n = 1,
      with_ties = FALSE
    ) %>%
    mutate(extreme = "Maximum")
  
  bind_rows(min_row, max_row) %>%
    mutate(
      source = source_name,
      .before = 1
    )
}


# 3. Extract minimum and maximum scaled IR
annual_ir_minmax_table_scaled <- imap_dfr(
  annual_ir_list_scaled,
  extract_min_max_ir
)


# 4. Function to identify the level/group column
get_level_label <- function(df) {
  
  candidate_cols <- c(
    "age_group",
    "age2",
    "sex",
    "gender",
    "ethnicity",
    "eth6",
    "imd",
    "region",
    "dcat",
    "level",
    "group",
    "group_var",
    "group_name"
  )
  
  existing_cols <- intersect(candidate_cols, names(df))
  
  # If no grouping variable exists
  if (length(existing_cols) == 0) {
    return(rep(NA_character_, nrow(df)))
  }
  
  # Create empty character vector
  out <- rep(NA_character_, nrow(df))
  
  # Fill using whichever grouping column contains a value
  for (col in existing_cols) {
    
    vals <- as.character(df[[col]])
    
    rows_to_fill <- is.na(out) & !is.na(vals)
    
    out[rows_to_fill] <- vals[rows_to_fill]
  }
  
  return(out)
}


# 5. Format final minimum/maximum table
annual_ir_minmax_table_scaled_fmt <- annual_ir_minmax_table_scaled %>%
  mutate(

    Variable = case_when(
      source == "overall"   ~ "Overall",
      source == "age"       ~ "Age group",
      source == "sex"       ~ "Sex",
      source == "ethnicity" ~ "Ethnicity",
      source == "imd"       ~ "IMD",
      source == "region"    ~ "English region",
      source == "dcat"      ~ "Dental problem",
      TRUE                  ~ source
    ),

    Level = case_when(
      source == "age"       ~ as.character(`Age group.x`),
      source == "sex"       ~ as.character(Sex),
      source == "dcat"      ~ as.character(`Dental problem categories`),
      source == "imd"       ~ as.character(IMD),
      source == "ethnicity" ~ as.character(Ethnicity),
      source == "region"    ~ as.character(Region),
      source == "overall"   ~ "Overall",
      TRUE                  ~ NA_character_
    ),

    `IR per 1000py (95% CI)` = sprintf(
      "%.2f (%.2f - %.2f)",
      ir1000_scaled,
      lower95_scaled,
      upper95_scaled
    ),

    `Total person years` = as.integer(round(total_person_years)),
    Weight = sprintf("%.2f", weight)
  ) %>%
  
  dplyr::select(
    Variable,
    Level,
    Extreme = extreme,
    Year = year,
    `Total person years`,
    `N events` = cases,
    Weight,
    `IR per 1000py (95% CI)`
  ) %>%

  # Overall first, then remaining variables alphabetically
  arrange(
    Variable != "Overall",
    Variable,
    Extreme
  )


annual_ir_minmax_table_scaled_fmt <- annual_ir_minmax_table_scaled_fmt%>%
    mutate(
  `N events` = formatC(`N events`, format = "f", digits = 0, big.mark = ","),
`Total person years` = formatC(
  `Total person years`,format = "f", digits = 0,big.mark = ","),
    `IR per 1000 PY (95%CI)` = gsub("\\.","·",`IR per 1000py (95% CI)`)
)%>%
  dplyr::select(-`IR per 1000py (95% CI)`, -Weight)

# 7. Export final table
write_xlsx(
  annual_ir_minmax_table_scaled_fmt,
  paste0(
    tables_figures,
    "/formatted/Table S9. Maximum and minimum annual incidence rates by variable.xlsx"
  )
)
```


# Section 1.3: ESP scale
## Step 1: combine data
```{r}
ir_age_gender_scaled <- ir_ab2_gender %>%
  rename(cases = total_cases) %>%
  left_join(
    ir_age_gender_overall,
    by = c("gender", "age_band2")
  )
```


## Step 2: weight ir 
```{r}
# 1. create weight variable by n events/ total cases
ir_age_gender_scaled$weight <- as.numeric(ir_age_gender_scaled$cases/ir_age_gender_scaled$total_cases)



# 2: apply weight to ir 1000 and ci
ir_age_gender_scaled <- ir_age_gender_scaled %>%
mutate(`ir1000_scaled` = `ir1000` * weight,
`lower95_scaled` = `lower95_1000` * weight,
`upper95_scaled` = `upper95_1000` * weight
)

```


## Step 3: asir
```{r}

## ============================================================
## Step 2: Calculate overall age-standardised scaled IR
## ============================================================

asir_age_scaled <- ir_age_overall %>%
  summarise(
    
    value_scaled = sum(
      `IR per 1000py` * esp2013,
      na.rm = TRUE
    ) / sum(
      esp2013,
      na.rm = TRUE
    ),
    
    lowercl_scaled = sum(
      `Lower CI` * esp2013,
      na.rm = TRUE
    ) / sum(
      esp2013,
      na.rm = TRUE
    ),
    
    uppercl_scaled = sum(
      `upper CI` * esp2013,
      na.rm = TRUE
    ) / sum(
      esp2013,
      na.rm = TRUE
    ),
    
    total_count_scaled = sum(
      `N events`,
      na.rm = TRUE
    ),
    
    total_pop_scaled = sum(
      `Total person years`,
      na.rm = TRUE
    )
  )


## ============================================================
## Step 3: Calculate female age-standardised scaled IR
## ============================================================

asir_female_scaled <- ir_age_gender_scaled %>%
  filter(gender == "Female") %>%
  summarise(
    
    value_scaled = sum(
      ir1000_scaled * (esp2013 / 2),
      na.rm = TRUE
    ) / sum(
      esp2013 / 2,
      na.rm = TRUE
    ),
    
    lowercl_scaled = sum(
      lower95_scaled * (esp2013 / 2),
      na.rm = TRUE
    ) / sum(
      esp2013 / 2,
      na.rm = TRUE
    ),
    
    uppercl_scaled = sum(
      upper95_scaled * (esp2013 / 2),
      na.rm = TRUE
    ) / sum(
      esp2013 / 2,
      na.rm = TRUE
    ),
    
    total_count_scaled = sum(
      cases,
      na.rm = TRUE
    ),
    
    total_pop_scaled = sum(
      total_person_year,
      na.rm = TRUE
    )
  )


## ============================================================
## Step 4: Calculate male age-standardised scaled IR
## ============================================================

asir_male_scaled <- ir_age_gender_scaled %>%
  filter(gender == "Male") %>%
  summarise(
    
    value_scaled = sum(
      ir1000_scaled * (esp2013 / 2),
      na.rm = TRUE
    ) / sum(
      esp2013 / 2,
      na.rm = TRUE
    ),
    
    lowercl_scaled = sum(
      lower95_scaled * (esp2013 / 2),
      na.rm = TRUE
    ) / sum(
      esp2013 / 2,
      na.rm = TRUE
    ),
    
    uppercl_scaled = sum(
      upper95_scaled * (esp2013 / 2),
      na.rm = TRUE
    ) / sum(
      esp2013 / 2,
      na.rm = TRUE
    ),
    
    total_count_scaled = sum(
      cases,
      na.rm = TRUE
    ),
    
    total_pop_scaled = sum(
      total_person_year,
      na.rm = TRUE
    )
  )


## ============================================================
## Step 5: Add labels
## ============================================================

asir_age_scaled <- asir_age_scaled %>%
  mutate(
    variable_scaled = "Overall"
  )

asir_female_scaled <- asir_female_scaled %>%
  mutate(
    variable_scaled = "Female"
  )

asir_male_scaled <- asir_male_scaled %>%
  mutate(
    variable_scaled = "Male"
  )


## ============================================================
## Step 6: Combine scaled ASIR results
## ============================================================

asir_combined_scaled <- bind_rows(
  asir_age_scaled,
  asir_female_scaled,
  asir_male_scaled
)


## ============================================================
## Step 7: Format combined table
## ============================================================

asir_combined_formatted_scaled <- asir_combined_scaled %>%
  mutate(
    
    `Standardised IR` = variable_scaled,
    
    `N events` = as.integer(
      total_count_scaled
    ),
    
    `Person years` = as.integer(
      round(total_pop_scaled)
    ),
    
    `IR per 1000py (95% CI)` = sprintf(
      "%.2f (%.2f - %.2f)",
      value_scaled,
      lowercl_scaled,
      uppercl_scaled
    )
  ) %>%
  dplyr::select(
    `Standardised IR`,
    `N events`,
    `Person years`,
    `IR per 1000py (95% CI)`
  )


## ============================================================
## Step 8: View final table
## ============================================================

asir_combined_formatted_scaled <- asir_combined_formatted_scaled %>%
  mutate(
   `N events` = formatC(`N events`, format = "f", digits = 0, big.mark = ","),
 `Total person years` = formatC(
   `Person years`,format = "f", digits = 0,big.mark = ","),
    `IR per 1000 PY (95%CI)` = gsub("\\.","·",`IR per 1000py (95% CI)`)
 )%>%
  dplyr::select(-`IR per 1000py (95% CI)`, -`Person years`)


## ============================================================
## Step 9: Export
## ============================================================

write_xlsx(
  asir_combined_formatted_scaled,
  paste0(
    tables_figures,
    "/formatted/Table S6.Directly standardised incidence rate based on the ESP2013 population, per 1000 person years_scaled.xlsx"
  )
)
```


# Section 1.4: update ITS model
## Step 1: import policy and IR
```{r}


ir_monthly <- records_time_agg %>%
  dplyr::select(year, month, total_records, total_person_years,
                ir1000, lower95_1000, upper95_1000, label)

# 1. fill in 2023-12-31 for those end time NA
policy_time$end_time[is.na(policy_time$end_time)] <- "2023-12-31"


# 2. convert to date variable
policy_time$start_time <- as.Date(policy_time$start_time)
policy_time$end_time <- as.Date(policy_time$end_time)

```


## Step 2: re aggregate the data 
```{r}
## 0. Make sure dates are Date class ---------------------------------------

policy_time <- policy_time %>%
  mutate(
    start_time = as.Date(start_time),
    end_time   = as.Date(end_time)
  ) %>%
  arrange(start_time)

## 1. Monthly ir data ------------------------------------------------------

ir_monthly$date <- as.Date(paste0(ir_monthly$label, "-01"))

records_time_agg <- ir_monthly %>%
  arrange(date)

## 2A. Expand policies into MONTHLY rows for 0/1 (use original end_time) ---

policy_monthly_dummy <- policy_time %>%
  mutate(
    start_month = floor_date(start_time, "month"),
    end_month   = floor_date(end_time, "month")
  ) %>%
  rowwise() %>%
  mutate(date = list(seq.Date(start_month, end_month, by = "month"))) %>%
  unnest(date) %>%
  ungroup()

## 3A. Create 0/1 wide policy-active matrix (unchanged logic) -------------

policy_wide <- policy_monthly_dummy %>%
  mutate(active = 1L) %>%
  dplyr::select(date, policy_label, active) %>%
  pivot_wider(
    names_from  = policy_label,
    values_from = active,
    values_fill = 0L
  )

## 4A. Join 0/1 policy-active to monthly ir -------------------------------

records_time_agg <- records_time_agg %>%
  left_join(policy_wide, by = "date") %>%
  mutate(across(starts_with("policy_"), ~ replace_na(., 0L))) %>%
  mutate(across(starts_with("gap_"), ~ replace_na(., 0L))) %>%
  arrange(date)

## 5. Create time index (1, 2, 3, ..., per month) -------------------------

records_time_agg <- records_time_agg %>%
  mutate(time_point = row_number())

## 2B. Prepare policy table for SLOPES (end at next policy start - 1 month)

policy_for_slope <- policy_time %>%
  arrange(start_time) %>%
  mutate(
    next_start  = lead(start_time),
    start_month = floor_date(start_time, "month"),
    eff_end = 
      # end until policy end
    floor_date(end_time, "month")
  )

## 3B. Expand into MONTHLY rows for slopes (start_month -> eff_end) -------

policy_monthly_slope <- policy_for_slope %>%
  rowwise() %>%
  mutate(date = list(seq.Date(start_month, eff_end, by = "month"))) %>%
  unnest(date) %>%
  ungroup()

## 4B. Create slope variables (0,1,2,... only while active for slope) -----

policy_monthly_slope <- policy_monthly_slope %>%
  group_by(policy_label) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(slope = row_number() - 1L) %>%
  ungroup()

policy_slope_wide <- policy_monthly_slope %>%
  dplyr::select(date, policy_label, slope) %>%
  pivot_wider(
    names_from  = policy_label,
    values_from = slope,
    names_glue  = "timeSince_{policy_label}",
    values_fill = 0L
  )

## 5B. Join slope columns into monthly aggregated data --------------------

records_time_agg <- records_time_agg %>%
  left_join(policy_slope_wide, by = "date") %>%
  mutate(across(starts_with("timeSince_"), ~ replace_na(., 0L)))


```


## Step 3: pre-analysis check 
```{r}


mean(records_time_agg$ir1000_scaled ) 
var(records_time_agg$ir1000_scaled ) 

# var >> mean -- use negative binomial ITS (for count/integer)

```


## Step 4: rerun the model with scaled data 
```{r}

## 1) Fit NB ITS model with spline of time_point

records_time_agg$year <- as.integer(records_time_agg$year )
annual_ir_scaled$year <- as.integer(annual_ir_scaled$year )

records_time_agg <- records_time_agg %>%
  arrange(date) %>%
  mutate(
    year2 = lubridate::year(date)
  ) %>%
  left_join(
    annual_ir_scaled %>%
      dplyr::select(year, weight),
    by = c("year2" = "year")
  ) %>%
  mutate(
    ir1000_scaled = ir1000 * weight,
    lower95_scaled = lower95_1000 * weight,
    upper95_scaled = upper95_1000 * weight
  )
  

# crude
nb_its_crude <- glm.nb(
 ir1000 ~ time_point +
   policy_D + timeSince_policy_D +
  policy_E + timeSince_policy_E +
  policy_F + timeSince_policy_F +
  policy_G + timeSince_policy_G 
  ,data = records_time_agg) 

summary(nb_its_crude)

# spline
m_nb_spline <- glm.nb(
  ir1000_scaled ~
    policy_D + timeSince_policy_D +
  policy_E + timeSince_policy_E +
  policy_F + timeSince_policy_F +
  policy_G + timeSince_policy_G +
    
    ## Flexible spline of time for long-term + seasonal patterns
    ns(time_point, df = 3),
  
  data = records_time_agg
)

summary(m_nb_spline)

```

## Step 5: overall ITS plot reformat
```{r}
#---------------------------------------------------------
# 1) Add fitted values from your NB model
#---------------------------------------------------------


records_time_agg <- records_time_agg %>%
  arrange(date) %>%
  mutate(
    fit_nb = predict(m_nb_spline, type = "response"),
  ) 



#---------------------------------------------------------
# 2) Plot: grey dots + orange fit line + selected policy lines
#---------------------------------------------------------

p_its <- ggplot(records_time_agg, aes(x = date)) +

  geom_vline(
    data = policy_time_filtered,
    aes(xintercept = start_time),
    linetype = "dashed",
    linewidth = 0.6,
    color = "grey30"
  ) +

  geom_text(
    data = policy_time_filtered,
    aes(x = start_time, label = paste0(cut_id)),
    y = Inf,
    angle = 90,
    vjust = -0.5,
    size = 3
  ) +

  geom_point(aes(y = ir1000_scaled), alpha = 0.7, color = "grey50") +

  geom_line(aes(y = fit_nb), linewidth = 1, color = "#E34000") +

  scale_x_date(
    breaks = seq(as.Date("1998-01-01"), as.Date("2023-01-01"), by = "1 year"),
    labels = function(x) {
      yr <- as.integer(format(x, "%Y"))
      ifelse(yr %% 2 == 0, as.character(yr), "")
    }
  ) +

  scale_y_continuous(
    limits = c(0, 15),
    breaks = seq(0, 15, by = 5)
  ) +

  labs(
    x = "Year",
    y = "Incidence rate (per 1000 PY)"
  ) +

  theme_minimal() +

  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 8),

    axis.line = element_line(linewidth = 0.4, colour = "grey30"),
    axis.ticks = element_line(linewidth = 0.4, colour = "grey30"),
    axis.ticks.length = unit(0.7, "mm"),

    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),

    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )


#---------------------------------------------------------
# 4) Save plot
#---------------------------------------------------------
ggsave(
  paste0(tables_figures, "/formatted/its_withoutABC_spline_scaled.jpeg"),
  plot   = p_its,
  width  = 7,
  height = 5,
  dpi    = 600
)
ggsave(
  paste0(tables_figures, "/formatted/its_withoutABC_spline_scaled.pdf"),
  plot   = p_its,
  width  = 7,
  height = 5,
  dpi    = 600
)
```

# Section 1.5: update ITS covid


## Step 1: import policy and IR
```{r}

ir_semi_monthly <- records_time_agg_covid %>%
  dplyr::select(year, month, total_records, total_person_years,
                ir1000, lower95_1000, upper95_1000, label,date)

# 1. fill in 2023-12-31 for those end time NA
policy_time$end_time[is.na(policy_time$end_time)] <- "2023-12-31"


# 2. convert to date variable
policy_time$start_time <- as.Date(policy_time$start_time)
policy_time$end_time <- as.Date(policy_time$end_time)

```

## Step 2: re aggregate the data 
time since policy replaced interaction term as it in equavalent to the outcome of time*policy
```{r}
## 0. Make sure dates are Date class ---------------------------------------


policy_time_covid <- policy_time %>%
  mutate(
    start_time = as.Date(start_time),
    end_time   = as.Date(end_time)
  ) %>%
  # ---- ONLY CHANGE: keep policies starting in or after 2019 ----
  filter(start_time >= as.Date("2019-01-01")) %>%
  arrange(start_time)


## 1. Semi-month IR data (COVID) ------------------------------------------
semi_start <- function(x) {
  x <- as.Date(x)
  if_else(day(x) <= 15,
          make_date(year(x), month(x), 1),
          make_date(year(x), month(x), 16))
}

## step to next semi-month start (1st -> 16th; 16th -> 1st next month)
semi_next <- function(d) {
  d <- as.Date(d)
  if_else(day(d) == 1,
          d + days(15),  # 1st -> 16th
          make_date(year(d %m+% months(1)), month(d %m+% months(1)), 1)) # 16th -> 1st next month
}

## sequence of semi-month starts from a to b (inclusive)
semi_seq <- function(a, b) {
  a <- as.Date(a); b <- as.Date(b)
  out <- c(a)
  if (a == b) return(out)
  cur <- a
  while (cur < b) {
    cur <- semi_next(cur)
    out <- c(out, cur)
  }
  out[out <= b]
}

## subtract one semi-month from a semi-month start
semi_prev <- function(d) {
  d <- as.Date(d)
  if_else(day(d) == 16,
          d - days(15),  # 16th -> 1st same month
          make_date(year(d %m-% months(1)), month(d %m-% months(1)), 16)) # 1st -> 16th prev month
}



records_time_agg_covid <- ir_semi_monthly%>%
  arrange(date)

## 2A. Expand policies into SEMI-MONTHLY rows for 0/1 (use original end_time) ---

policy_semimonth_dummy_covid <- policy_time_covid %>%
  mutate(
    start_period = semi_start(start_time),
    end_period   = semi_start(end_time)
  ) %>%
  rowwise() %>%
  mutate(date = list(semi_seq(start_period, end_period))) %>%
  unnest(date) %>%
  ungroup()

## 3A. Create 0/1 wide policy-active matrix (unchanged logic) -------------

policy_wide_covid <- policy_semimonth_dummy_covid %>%
  mutate(active = 1L) %>%
  dplyr::select(date, policy_label, active) %>%
  pivot_wider(
    names_from  = policy_label,
    values_from = active,
    values_fill = 0L
  )

## 4A. Join 0/1 policy-active to semi-month IR (COVID) --------------------

records_time_agg_covid <- records_time_agg_covid %>%
  left_join(policy_wide_covid, by = "date") %>%
  mutate(across(starts_with("policy_"), ~ replace_na(., 0L))) %>%
  mutate(across(starts_with("gap_"), ~ replace_na(., 0L))) %>%
  arrange(date)

## 5. Create time index (1, 2, 3, ..., per semi-month) --------------------

records_time_agg_covid <- records_time_agg_covid %>%
  mutate(time_point = row_number())

## 2B. Prepare policy table for SLOPES (end at next policy start - 1 period)

policy_for_slope_covid <- policy_time_covid %>%
  arrange(start_time) %>%
  mutate(
    next_start   = lead(start_time),
    start_period = semi_start(start_time),
    eff_end = semi_start(end_time)
      
      
      #case_when(
      #!is.na(next_start) ~ semi_prev(semi_start(next_start)),  # next policy start - 1 semi-month
      #TRUE              ~ semi_start(end_time)
    #)
  )

## 3B. Expand into SEMI-MONTHLY rows for slopes (start_period -> eff_end) --

policy_semimonth_slope_covid <- policy_for_slope_covid %>%
  rowwise() %>%
  mutate(date = list(semi_seq(start_period, eff_end))) %>%
  unnest(date) %>%
  ungroup()

## 4B. Create slope variables (0,1,2,... only while active for slope) -----

policy_semimonth_slope_covid <- policy_semimonth_slope_covid %>%
  group_by(policy_label) %>%
  arrange(date, .by_group = TRUE) %>%
  mutate(slope = row_number() - 1L) %>%
  ungroup()

policy_slope_wide_covid <- policy_semimonth_slope_covid %>%
  dplyr::select(date, policy_label, slope) %>%
  pivot_wider(
    names_from  = policy_label,
    values_from = slope,
    names_glue  = "timeSince_{policy_label}",
    values_fill = 0L
  )

## 5B. Join slope columns into semi-month aggregated data (COVID) ---------

records_time_agg_covid <- records_time_agg_covid %>%
  left_join(policy_slope_wide_covid, by = "date") %>%
  mutate(across(starts_with("timeSince_"), ~ replace_na(., 0L)))

## 6. create interaction variable (unchanged) -----------------------------

interaction_vars <- c(
  A = "policy_A",
  B = "policy_B",
  C = "policy_C",
  D = "policy_D",
  E = "policy_E",
  `F` = "policy_F",
  G = "policy_G",
  gap_FG = "gap_FG"

)

for (new in names(interaction_vars)) {
  old <- interaction_vars[[new]]
  if (old %in% names(records_time_agg_covid)) {
    records_time_agg_covid[[paste0("interaction_", new)]] <-
      records_time_agg_covid$time_point * records_time_agg_covid[[old]]
  }
}



```


### Step 2.1: negative binomial ITS model (COVID)
```{r}

records_time_agg$year <- as.integer(records_time_agg$year )
annual_ir_scaled$year <- as.integer(annual_ir_scaled$year )

records_time_agg_covid <- records_time_agg_covid %>%
  arrange(date) %>%
  mutate(
    year2 = lubridate::year(date)
  ) %>%
  left_join(
    annual_ir_scaled %>%
      dplyr::select(year, weight),
    by = c("year2" = "year")
  ) %>%
  mutate(
    ir1000_scaled = ir1000 * weight,
    lower95_scaled = lower95_1000 * weight,
    upper95_scaled = upper95_1000 * weight
  )

# 1. fit the model with all policy time 

nb_its_crude_covid <- glm.nb(
  ir1000_scaled ~ time_point +
  policy_F + timeSince_policy_F +
  policy_G + timeSince_policy_G ,
  data = records_time_agg_covid
)

summary(nb_its_crude_covid)





```
### Step 2.2: plot
```{r}
#---------------------------------------------------------
# 1) Add fitted values from your NB model
#---------------------------------------------------------
records_time_agg_covid <- records_time_agg_covid |>
  arrange(date) |>
  mutate(
    fit_nb = predict(nb_its_crude_covid, type = "response")
  )

#---------------------------------------------------------
# 2) Ensure policy_time dates are Date class + keep only E, F, gap EF (COVID)
#---------------------------------------------------------
policy_time_filtered_covid <- policy_time_covid |>
  mutate(
    start_time = as.Date(start_time),
    end_time   = as.Date(end_time)
  ) |>
  filter(policy_label %in% c("policy_F", "policy_G", "gap_FG")) |>
  mutate(
    cut_id = dplyr::row_number()
  )

#---------------------------------------------------------
# 3) Plot: grey dots + orange fit line + selected policy lines
#---------------------------------------------------------

p_covid <- ggplot(records_time_agg_covid, aes(x = date)) +
  
  # vertical policy lines for E, F, gap EF
  geom_vline(
    data = policy_time_filtered_covid,
    aes(xintercept = start_time),
    linetype = "dashed",
    linewidth = 0.6,
    color = "grey30"
  ) +
  
  geom_text(
    data = policy_time_filtered_covid,
    aes(
      x = start_time,
      label = paste0(cut_id)
    ),
    y = Inf,
    angle = 90,
    vjust = -0.5,
    size = 3
  ) +
  
  # grey observed dots
  geom_point(aes(y = ir1000_scaled), alpha = 0.7, color = "grey50") +
  
  # fitted NB line (orange)
  geom_line(aes(y = fit_nb), linewidth = 1, color = "#E34000") +
  
  # ---- YEARLY x-axis (same as first plot) ----
  scale_x_date(
    breaks = seq(
      as.Date(format(min(records_time_agg_covid$date), "%Y-01-01")),
      as.Date(format(max(records_time_agg_covid$date), "%Y-01-01")),
      by = "1 year"
    ),
    date_labels = "%Y"
  ) +
  
  # -------------------------------
  # FIX Y SCALE TO 0–25
  # -------------------------------
  scale_y_continuous(
    limits = c(0, 15),
    breaks = seq(0, 15, by = 5)
  ) +
  
  labs(
    x = "Year",
    y = "Semi-monthly IR per 1000 PY"
  ) +
  
  theme_minimal() +
theme(
  plot.title = element_text(hjust = 0.5, face = "bold"),
axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5, size = 8),
  # add lined axes + ticks
  axis.line = element_line(linewidth = 0.4, colour = "grey30"),
  axis.ticks = element_line(linewidth = 0.4, colour = "grey30"),
  axis.ticks.length = unit(0.7, "mm"),

  # minimal theme removes ticks by default; turn them on
  axis.ticks.x = element_line(),
  axis.ticks.y = element_line(),

  panel.grid.major.x = element_blank(),
  panel.grid.minor.x = element_blank()
)

#---------------------------------------------------------
# 4) Save
#---------------------------------------------------------
ggsave(
  paste0(tables_figures, "/formatted/its_withoutABC_crude_covid_scaled.jpeg"),
  plot  = p_covid,
  width = 5,
  height = 5,
  dpi = 600
)

ggsave(
  paste0(tables_figures, "/formatted/its_withoutABC_crude_covid_scaled.pdf"),
  plot  = p_covid,
  width = 5,
  height = 5,
  dpi = 600
)
```

### Step 2.3: combine ACF and export (figure S4)
```{r}
########################################################
library(ggplot2)
library(forecast)   # ggAcf()
library(patchwork)

# -----------------------------
# Helper: ACF plot with fixed y-scale
# -----------------------------
make_acf_plot <- function(x, title, lag.max = 36) {
  forecast::ggAcf(x, lag.max = lag.max) +
    labs(title = title, x = "Lag", y = "ACF") +
    coord_cartesian(ylim = c(-1, 1)) +   # <<< FIXED Y SCALE
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold")
    )
}

# -----------------------------
# Build 4 ACF plots
# -----------------------------
pA <- make_acf_plot(
  residuals(nb_its_crude, type = "pearson"),
  "(A) ACF of Pearson residuals: NB ITS crude"
)
ggsave(pA, filename =paste0(tables_figures,"/acf_4panel_a.jpeg"), width = 12, height = 8, dpi = 600)

pB <- make_acf_plot(
  residuals(m_nb_spline, type = "pearson"),
  "(B) ACF of Pearson residuals: NB ITS + spline"
)
pB
ggsave(
  filename = paste0(tables_figures, "/acf_4panel_b.jpeg"),
  plot = pB,
  width = 12,
  height = 8,
  dpi = 600
)

pC <- make_acf_plot(
  residuals(nb_its_crude_covid, type = "pearson"),
  "(C) ACF of Pearson residuals: NB ITS crude (COVID)"
)
pC
pD <- make_acf_plot(
  residuals(m_nb_spline_covid, type = "pearson"),
  "(D) ACF of Pearson residuals: NB ITS + spline (COVID)"
)
pD
# -----------------------------
# Patch into 2x2 grid
# -----------------------------
library(patchwork)

p_ABC <- (pA | pB) /
            (pC | plot_spacer())

ggsave(
  filename = file.path(tables_figures, "formatted", "Figure S4. ACF and partial ACF for ITS analysis.jpeg"),
  plot = p_ABC,
  width = 12,
  height = 8,
  dpi = 600
)


```



### Step 2.4: create table 
```{r}
extract_nb_its_irr <- function(model,
                              keep = c("policy", "policy_slope"),
                              include_time_point = FALSE,
                              ci_type = c("wald", "profile"),
                              model_name = NULL) {

  ci_type <- match.arg(ci_type)
  if (is.null(model_name)) model_name <- deparse(substitute(model))

  est <- stats::coef(model)

  # --- CIs on beta scale ---
  if (ci_type == "wald") {
    se <- sqrt(diag(stats::vcov(model)))
    lci_beta <- est - 1.96 * se
    uci_beta <- est + 1.96 * se
    p_value  <- 2 * stats::pnorm(abs(est / se), lower.tail = FALSE)
  } else {
    ci_beta <- suppressMessages(stats::confint(model))  # profile CI for glm.nb
    lci_beta <- ci_beta[, 1]
    uci_beta <- ci_beta[, 2]
    se <- sqrt(diag(stats::vcov(model)))
    p_value <- 2 * stats::pnorm(abs(est / se), lower.tail = FALSE)
  }

  res <- dplyr::tibble(
    model = model_name,
    term  = names(est),
    beta  = unname(est),
    lci_beta = unname(lci_beta),
    uci_beta = unname(uci_beta),
    IRR     = exp(beta),
    lci_IRR = exp(lci_beta),
    uci_IRR = exp(uci_beta),
    p_value = unname(p_value)
  )

  # --- decide which terms to keep (based on term text; robust to ordering) ---
  keep_idx <- rep(FALSE, nrow(res))

  if ("policy" %in% keep) {
    keep_idx <- keep_idx | grepl("^policy_[DEFG]$", res$term)
  }
  if ("policy_slope" %in% keep) {
    keep_idx <- keep_idx | grepl("^(timeSince_policy_[DEFG])$", res$term)
  }
  if (include_time_point) {
    keep_idx <- keep_idx | (res$term == "time_point")
  }

res <- res |>
  dplyr::filter(keep_idx) |>
  dplyr::mutate(
    effect_type = dplyr::case_when(
      term == "time_point" ~ "Baseline trend (per time unit)",
      grepl("^policy_[DEFG]$", term)  ~ "Level change",
        grepl("^timeSince_policy_[DEFG]$", term) ~ "Slope change",
      TRUE ~ "Other"
    )
  ,
      policy = dplyr::case_when(
        grepl("policy_D", term) ~ "policy_D",
        grepl("policy_E", term) ~ "policy_E",
        grepl("policy_F", term) ~ "policy_F",
        grepl("policy_G", term) ~ "policy_G",
        term == "time_point"    ~ "time_point",
        TRUE ~ term
      ),
      IRR_CI = sprintf("%.2f (%.2f–%.2f)", IRR, lci_IRR, uci_IRR)
    ) |>
    dplyr::select(model, policy, effect_type, term, beta, IRR, lci_IRR, uci_IRR, IRR_CI, p_value)

  res
}

# crude
res_crude_alltime <- extract_nb_its_irr(nb_its_crude, ci_type = "wald", model_name = "All-time crude")

# spline/seasonality adjusted
res_spline_alltime <- extract_nb_its_irr(m_nb_spline, ci_type = "wald", model_name = "All-time + spline")
# crude
res_crude_covid <- extract_nb_its_irr(nb_its_crude_covid, ci_type = "wald", model_name = "COVID crude")


its_combined <- dplyr::bind_rows(res_crude_alltime,  res_spline_alltime,
                                 res_crude_covid)

write_xlsx(its_combined,paste0(tables_figures,"/its_results_v6.xlsx") )
```

####add yearly IRR and CI
```{r}
extract_nb_its_irr <- function(model,
                              keep = c("policy", "policy_slope", "gap", "gap_slope"),
                              include_time_point = FALSE,
                              ci_type = c("wald", "profile"),
                              model_name = NULL) {

  ci_type <- match.arg(ci_type)
  if (is.null(model_name)) model_name <- deparse(substitute(model))

  est <- stats::coef(model)
  se  <- sqrt(diag(stats::vcov(model)))

  # --- CIs on beta scale ---
  if (ci_type == "wald") {
    lci_beta <- est - 1.96 * se
    uci_beta <- est + 1.96 * se
    p_value  <- 2 * stats::pnorm(abs(est / se), lower.tail = FALSE)
  } else {
    ci_beta <- suppressMessages(stats::confint(model))  # profile CI for glm.nb
    lci_beta <- ci_beta[, 1]
    uci_beta <- ci_beta[, 2]
    p_value  <- 2 * stats::pnorm(abs(est / se), lower.tail = FALSE)
  }

  res <- dplyr::tibble(
    model    = model_name,
    term     = names(est),
    beta     = unname(est),
    se       = unname(se),
    lci_beta = unname(lci_beta),
    uci_beta = unname(uci_beta),
    IRR      = exp(beta),
    lci_IRR  = exp(lci_beta),
    uci_IRR  = exp(uci_beta),
    p_value  = unname(p_value)
  )

  # --- decide which terms to keep (based on term text; robust to ordering) ---
  keep_idx <- rep(FALSE, nrow(res))

  if ("policy" %in% keep) {
    keep_idx <- keep_idx | grepl("^policy_[DEFG]$", res$term)
  }
  if ("policy_slope" %in% keep) {
    keep_idx <- keep_idx | grepl("^(timeSince_policy_[DEFG])$", res$term)
  }
  if (include_time_point) {
    keep_idx <- keep_idx | (res$term == "time_point")
  }

  res <- res |>
    dplyr::filter(keep_idx) |>
    dplyr::mutate(
      effect_type = dplyr::case_when(
        term == "time_point" ~ "Baseline trend (per time unit)",
        grepl("^policy_[DEFG]$", term)  ~ "Level change",
          grepl("^timeSince_policy_[DEFG]$", term) ~ "Slope change",
        TRUE ~ "Other"
      ),
      policy = dplyr::case_when(
        grepl("policy_D", term) ~ "policy_D",
        grepl("policy_E", term) ~ "policy_E",
        grepl("policy_F", term) ~ "policy_F",
        grepl("policy_G", term) ~ "policy_G",
        term == "time_point"    ~ "time_point",
        TRUE ~ term
      ),
      IRR_CI = sprintf("%.2f (%.2f–%.2f)", IRR, lci_IRR, uci_IRR),

      # ---- annualise only trend / slope terms using beta and SE first ----
      year_beta = dplyr::case_when(
        effect_type %in% c("Baseline trend (per time unit)", "Slope change") ~ beta * 12,
        TRUE ~ NA_real_
      ),
      year_se = dplyr::case_when(
        effect_type %in% c("Baseline trend (per time unit)", "Slope change") ~ se * 12,
        TRUE ~ NA_real_
      ),

      year_lci_beta = dplyr::case_when(
        effect_type %in% c("Baseline trend (per time unit)", "Slope change") ~ year_beta - 1.96 * year_se,
        TRUE ~ NA_real_
      ),
      year_uci_beta = dplyr::case_when(
        effect_type %in% c("Baseline trend (per time unit)", "Slope change") ~ year_beta + 1.96 * year_se,
        TRUE ~ NA_real_
      ),

      year_IRR = exp(year_beta),
      year_lci_IRR = exp(year_lci_beta),
      year_uci_IRR = exp(year_uci_beta),

      yearly_IRR_CI = dplyr::case_when(
        effect_type %in% c("Baseline trend (per time unit)", "Slope change") ~
          sprintf("%.2f (%.2f–%.2f)", year_IRR, year_lci_IRR, year_uci_IRR),
        TRUE ~ NA_character_
      )
    ) |>
    dplyr::select(
      model, policy, effect_type, term,
      beta, se, lci_beta, uci_beta,
      IRR, lci_IRR, uci_IRR, IRR_CI,
      year_beta, year_se, year_lci_beta, year_uci_beta,
      year_IRR, year_lci_IRR, year_uci_IRR, yearly_IRR_CI,
      p_value
    )

  res
}

# crude
res_crude_alltime <- extract_nb_its_irr(
  nb_its_crude,
  ci_type = "wald",
  model_name = "All-time crude"
)

# spline/seasonality adjusted
res_spline_alltime <- extract_nb_its_irr(
  m_nb_spline,
  ci_type = "wald",
  model_name = "All-time + spline"
)

# crude covid
res_crude_covid <- extract_nb_its_irr(
  nb_its_crude_covid,
  ci_type = "wald",
  model_name = "COVID crude"
)

its_combined_scaled <- dplyr::bind_rows(
  res_crude_alltime,
  res_spline_alltime,
  res_crude_covid
)
res_crude_alltime_scaled_formatted <- res_crude_alltime %>%
  mutate(
    `Crude IRR (95% CI) from 1998-2023` = ifelse(
      effect_type == "Level change" & is.na(yearly_IRR_CI),
      IRR_CI,
      yearly_IRR_CI
    )
  ) %>%
  rename(
    Policy = policy,
    `Effect type` = effect_type
  )

res_spline_alltime_scaled_formatted <- res_spline_alltime %>%
  mutate(
    `Adjusted IRR (95% CI) from 1998-2023` = ifelse(
      effect_type == "Level change" & is.na(yearly_IRR_CI),
      IRR_CI,
      yearly_IRR_CI
    )
  ) %>%
  rename(
    Policy = policy,
    `Effect type` = effect_type
  )

res_crude_covid_scaled_formatted <- res_crude_covid %>%
  mutate(
    `Crude IRR (95% CI) from 2019-2023` = ifelse(
      effect_type == "Level change" & is.na(yearly_IRR_CI),
      IRR_CI,
      yearly_IRR_CI
    )
  ) %>%
  rename(
    Policy = policy,
    `Effect type` = effect_type
  )

its_combined_scaled_formatted <- res_crude_alltime_scaled_formatted %>%
  left_join(
    res_spline_alltime_scaled_formatted %>%
      rename(p_spline = p_value),
    by = c("Policy", "Effect type")
  ) %>%
  left_join(
    res_crude_covid_scaled_formatted %>%
      rename(p_covid = p_value),
    by = c("Policy", "Effect type")
  ) %>%
  mutate(
    `P crude 1998-2023` = case_when(
      is.na(p_value)    ~ NA_character_,
      p_value < 0.0001  ~ "<0.0001",
      p_value < 0.001   ~ sprintf("%.5f", p_value),
      p_value < 0.01    ~ sprintf("%.4f", p_value),
      p_value < 0.1     ~ sprintf("%.3f", p_value),
      TRUE              ~ sprintf("%.2f", p_value)
    ),

    `P adjusted 1998-2023` = case_when(
      is.na(p_spline)    ~ NA_character_,
      p_spline < 0.0001  ~ "<0.0001",
      p_spline < 0.001   ~ sprintf("%.5f", p_spline),
      p_spline < 0.01    ~ sprintf("%.4f", p_spline),
      p_spline < 0.1     ~ sprintf("%.3f", p_spline),
      TRUE               ~ sprintf("%.2f", p_spline)
    ),

    `P crude 2019-2023` = case_when(
      is.na(p_covid)    ~ NA_character_,
      p_covid < 0.0001  ~ "<0.0001",
      p_covid < 0.001   ~ sprintf("%.5f", p_covid),
      p_covid < 0.01    ~ sprintf("%.4f", p_covid),
      p_covid < 0.1     ~ sprintf("%.3f", p_covid),
      TRUE              ~ sprintf("%.2f", p_covid)
    ),

    across(
      c(
        `P crude 1998-2023`,
        `P adjusted 1998-2023`,
        `P crude 2019-2023`
      ),
      ~ gsub("\\.", "·", .x)
    )
  ) %>%
  dplyr::select(
    Policy,
    `Effect type`,
    `Crude IRR (95% CI) from 1998-2023`,
    `P crude 1998-2023`,
    `Adjusted IRR (95% CI) from 1998-2023`,
    `P adjusted 1998-2023`,
    `Crude IRR (95% CI) from 2019-2023`,
    `P crude 2019-2023`
  )
its_combined_scaled_formatted <- its_combined_scaled_formatted%>%
  mutate(
    `Crude IRR (95% CI) from 1998-2023` = gsub("\\.","·",`Crude IRR (95% CI) from 1998-2023`),
     `Adjusted IRR (95% CI) from 1998-2023` = gsub("\\.","·",`Adjusted IRR (95% CI) from 1998-2023`),    
                                                `Crude IRR (95% CI) from 2019-2023` = gsub("\\.","·",`Crude IRR (95% CI) from 2019-2023`))

write_xlsx(
  its_combined_scaled_formatted,
  paste0(tables_figures, "/formatted/Table S13.xlsx")
)
```


# Section 1.6: use scaled regional ir for figure
## 1. prepare the data 
```{r}
regional_ir_scaled <-ir_summary_table_scaled%>%
  filter(variable=="Region")
```

## Step 2: regional incidence by map
```{r}
# 1. Load map
uk_map <- st_read(
  "georef-united-kingdom-county-unitary-authority.geojson",
  quiet = TRUE
)

# 2. Prepare map names and join IR data
uk_map <- uk_map %>%

  left_join(
    regional_ir_scaled,
    by = c("rgn_name" = "level")
  )

# Check whether values were matched
table(is.na(uk_map$`IR per 1000py`))

# 3. Colour palette
tailwind_cols <- pink_to_red

# 4. Plot
ir_region_uk_map <- ggplot(uk_map) +
  geom_sf(
    aes(fill = `IR per 1000py`),
    color = "white",
    linewidth = 0.2
  ) +
  scale_fill_gradientn(
    colours = tailwind_cols,
    na.value = "grey90"
  ) +
  labs(
    fill = "IR (per 1000 PY)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  ) +
  coord_sf(
    xlim = c(-6.5, 2),
    ylim = c(49.5, 56),
    expand = FALSE
  )

# 5. Save plot
ggsave(
  filename = paste0(
    tables_figures,
    "/Figure 1. Regional incidence rates of GP consultation for dental problems in England.pdf"
  ),
  plot = ir_region_uk_map,
  width = 8,
  height = 6,
  dpi = 600
)

# 7. add lable manually and convert to jpeg
pdf_convert(
  pdf = paste0(tables_figures, "/Figure 1. Regional incidence rates of GP consultation for dental problems in England.pdf"),
  format = "jpeg",
  dpi = 600,
  filenames = paste0(tables_figures, "/formatted/Figure 1. Regional incidence rates of GP consultation for dental problems .jpeg"
))
```

# Section 1.7: heatmap table and heatmap
## Step 1: combine and calculate weight
```{r}

#--------------------------------------------------------------------
# 1. Dental problem category × Region
#--------------------------------------------------------------------

dcat_region_scaled <- dcat_region %>% 
  rename(cases = total_cases) %>%
  left_join(
    ir_dcat_region_overall,
    by = c("Dental problem categories", "Region")
  ) %>%
  mutate(
    weight = cases / total_cases,
    ir1000_scaled = ir1000 * weight,
    lower95_scaled = lower95 * weight,
    upper95_scaled = upper95 * weight
  )


#--------------------------------------------------------------------
# 2. Dental problem category × Sex
#--------------------------------------------------------------------

dcat_sex_scaled <- dcat_sex %>% 
  rename(cases = total_cases) %>%
  left_join(
    ir_dcat_sex_overall,
    by = c("Dental problem categories", "sex")
  ) %>%
  mutate(
    weight = cases / total_cases,
    ir1000_scaled = ir1000 * weight,
    lower95_scaled = lower95 * weight,
    upper95_scaled = upper95 * weight
  )


#--------------------------------------------------------------------
# 3. Dental problem category × Age band
#--------------------------------------------------------------------

dcat_age2_scaled <- dcat_age2 %>%
  rename(cases = total_cases) %>%
  left_join(
    ir_age2_dcat_overall,
    by = c("Dental problem categories", "age_band2")
 ) %>%
  mutate(
        weight = cases / total_cases,

    ir1000_scaled = ir1000 * weight
  ) %>%
  rowwise() %>%
  mutate(
    tmp = list(
      epi.conf(
        cbind(cases, total_person_year),
        ctype = "inc.rate",
        method = "exact"
      ) * 1000
    ),
    lower95 = tmp$lower,
    upper95 = tmp$upper
  ) %>%
  dplyr::select(-tmp) %>%
  ungroup() %>%
  mutate(
    lower95_scaled = lower95 * weight,
    upper95_scaled = upper95 * weight
  )

#--------------------------------------------------------------------
# 4. Age band × Region
#--------------------------------------------------------------------

age2_rgn_scaled <- age2_rgn %>% 
  rename(cases = total_cases) %>%
  left_join(
    ir_ab2_rgn_overall,
    by = c(
      "Region" = "region",
      "age_band2"
    )
  ) %>%
  mutate(
    weight = cases / total_cases,
    ir1000_scaled = ir1000 * weight
  ) %>%
  rowwise() %>%
  mutate(
    tmp = list(
      epi.conf(
        cbind(cases, total_person_year),
        ctype = "inc.rate",
        method = "exact"
      ) * 1000
    ),
    lower95 = tmp$lower,
    upper95 = tmp$upper
  ) %>%
  dplyr::select(-tmp) %>%
  ungroup() %>%
  mutate(
    lower95_scaled = lower95 * weight,
    upper95_scaled = upper95 * weight
  )


#--------------------------------------------------------------------
# 5. Age band × Sex
#--------------------------------------------------------------------

age2_sex_scaled <- ir_age_gender_scaled 


#--------------------------------------------------------------------
# 6. Region × Sex
#--------------------------------------------------------------------

region_sex_scaled <- region_sex %>% 
  rename(cases = total_cases) %>%
  left_join(
    ir_region_sex_overall,
    by = c(
      "Region" = "region",
      "Sex" = "gender"
    )
  ) %>%
  mutate(
    weight = cases / total_cases,
    ir1000_scaled = ir1000 * weight
  ) %>%
  rowwise() %>%
  mutate(
    tmp = list(
      epi.conf(
        cbind(cases, total_person_year),
        ctype = "inc.rate",
        method = "exact"
      ) * 1000
    ),
    lower95 = tmp$lower,
    upper95 = tmp$upper
  ) %>%
  dplyr::select(-tmp) %>%
  ungroup() %>%
  mutate(
    lower95_scaled = lower95 * weight,
    upper95_scaled = upper95 * weight
  )


#====================================================================
# Ethnicity combinations

#====================================================================


#--------------------------------------------------------------------
# 7. Dental problem category × Ethnicity

#--------------------------------------------------------------------

dcat_eth_scaled <- dcat_eth %>%
   rename(cases = total_cases) %>%
 
  left_join(
    ir_dcat_overall,
    by = c("Dental problem categories" = "level")
  ) %>%
  mutate(
    ir1000_scaled = ir1000 * weight,
    lower95_scaled = lower95 * weight,
    upper95_scaled = upper95 * weight
  )


#--------------------------------------------------------------------
# 8. Region × Ethnicity

#--------------------------------------------------------------------

eth_region_scaled <- eth_region %>%
  rename(cases = total_cases) %>%
  left_join(
    ir_region_overall,
    by = c("Region" = "level")
  ) %>%
  mutate(
    ir1000_scaled = ir1000 * weight
  ) %>%
  rowwise() %>%
  mutate(
    tmp = list(
      epi.conf(
        cbind(cases, total_person_year),
        ctype = "inc.rate",
        method = "exact"
      ) * 1000
    ),
    lower95 = tmp$lower,
    upper95 = tmp$upper
  ) %>%
  dplyr::select(-tmp) %>%
  ungroup() %>%
  mutate(
    lower95_scaled = lower95 * weight,
    upper95_scaled = upper95 * weight
  )


#--------------------------------------------------------------------
# 9. Sex × Ethnicity

#--------------------------------------------------------------------

eth_sex_scaled <- eth_sex %>%
  rename(cases = total_cases) %>%
  left_join(
    ir_gender_overall,
    by = c("Sex" = "level")
  ) %>%
  mutate(
    ir1000_scaled = ir1000 * weight
  ) %>%
  rowwise() %>%
  mutate(
    tmp = list(
      epi.conf(
        cbind(cases, total_person_year),
        ctype = "inc.rate",
        method = "exact"
      ) * 1000
    ),
    lower95 = tmp$lower,
    upper95 = tmp$upper
  ) %>%
  dplyr::select(-tmp) %>%
  ungroup() %>%
  mutate(
    lower95_scaled = lower95 * weight,
    upper95_scaled = upper95 * weight
  )


#--------------------------------------------------------------------
# 10. Age band × Ethnicity

#--------------------------------------------------------------------

age2_eth_scaled <- age2_eth %>%
  rename(cases = total_cases) %>%
  left_join(
    ir_age_overall,
    by = c("age_band2" = "level")
  ) %>%
  mutate(
    ir1000_scaled = ir1000 * weight
  ) %>%
  rowwise() %>%
  mutate(
    tmp = list(
      epi.conf(
        cbind(cases, total_person_year),
        ctype = "inc.rate",
        method = "exact"
      ) * 1000
    ),
    lower95 = tmp$lower,
    upper95 = tmp$upper
  ) %>%
  dplyr::select(-tmp) %>%
  ungroup() %>%
  mutate(
    lower95_scaled = lower95 * weight,
    upper95_scaled = upper95 * weight
  )


#====================================================================
# IMD combinations
#====================================================================


#--------------------------------------------------------------------
# 11. Dental problem category × IMD
#--------------------------------------------------------------------

dcat_imd_scaled <- dcat_imd %>%
    rename(cases = total_cases) %>%

  left_join(
    ir_dcat_overall,
    by = c("Dental problem categories" = "level")
  ) %>%
  mutate(
    ir1000_scaled = ir1000 * weight,
    lower95_scaled = lower95 * weight,
    upper95_scaled = upper95 * weight
  )


#--------------------------------------------------------------------
# 12. Region × IMD
#--------------------------------------------------------------------

imd_region_scaled <- imd_region %>%
  rename(cases = total_cases) %>%
  left_join(
    ir_region_overall,
    by = c("Region" = "level")
  ) %>%
  mutate(
    ir1000_scaled = ir1000 * weight
  ) %>%
  rowwise() %>%
  mutate(
    tmp = list(
      epi.conf(
        cbind(cases, total_person_year),
        ctype = "inc.rate",
        method = "exact"
      ) * 1000
    ),
    lower95 = tmp$lower,
    upper95 = tmp$upper
  ) %>%
  dplyr::select(-tmp) %>%
  ungroup() %>%
  mutate(
    lower95_scaled = lower95 * weight,
    upper95_scaled = upper95 * weight
  )


#--------------------------------------------------------------------
# 13. Sex × IMD
#--------------------------------------------------------------------

imd_sex_scaled <- imd_sex %>%
  rename(cases = total_cases) %>%
  left_join(
    ir_gender_overall,
    by = c("Sex" = "level")
  ) %>%
  mutate(
    ir1000_scaled = ir1000 * weight
  ) %>%
  rowwise() %>%
  mutate(
    tmp = list(
      epi.conf(
        cbind(cases, total_person_year),
        ctype = "inc.rate",
        method = "exact"
      ) * 1000
    ),
    lower95 = tmp$lower,
    upper95 = tmp$upper
  ) %>%
  dplyr::select(-tmp) %>%
  ungroup() %>%
  mutate(
    lower95_scaled = lower95 * weight,
    upper95_scaled = upper95 * weight
  )


#--------------------------------------------------------------------
# 14. Age band × IMD

#--------------------------------------------------------------------

age2_imd_scaled <- age2_imd %>%
  rename(cases = total_cases) %>%
  left_join(
    ir_age_overall,
    by = c("age_band2" = "level")
  ) %>%
  mutate(
    ir1000_scaled = ir1000 * weight
  ) %>%
  rowwise() %>%
  mutate(
    tmp = list(
      epi.conf(
        cbind(cases, total_person_year),
        ctype = "inc.rate",
        method = "exact"
      ) * 1000
    ),
    lower95 = tmp$lower,
    upper95 = tmp$upper
  ) %>%
  dplyr::select(-tmp) %>%
  ungroup() %>%
  mutate(
    lower95_scaled = lower95 * weight,
    upper95_scaled = upper95 * weight
  )


#--------------------------------------------------------------------
# 15. Ethnicity × IMD
#--------------------------------------------------------------------

eth_imd_scaled <- eth_imd %>%
  rename(cases = total_cases) %>%
  mutate(
    ir1000_scaled = ir1000 * overall_weight
  ) %>%
  rowwise() %>%
  mutate(
    tmp = list(
      epi.conf(
        cbind(cases, total_person_year),
        ctype = "inc.rate",
        method = "exact"
      ) * 1000
    ),
    lower95 = tmp$lower,
    upper95 = tmp$upper
  ) %>%
  dplyr::select(-tmp) %>%
  ungroup() %>%
  mutate(
    lower95_scaled = lower95 * overall_weight,
    upper95_scaled = upper95 * overall_weight
  )
```

## Step 2: make heatmap
```{r}
############################################################
## SCALED TRIANGLE HEATMAP
############################################################

library(dplyr)
library(ggplot2)
library(stringr)
library(scales)
library(cowplot)
library(grid)
library(gtable)

############################
# 0) Helpers
############################

ETH_LEVELS <- c(
  "White",
  "Asian",
  "Black",
  "Mixed or Multiple",
  "Other",
  "Not specified"
)

short_eth <- function(x) {
  x %>%
    str_replace("^Black, Black British, Caribbean or African$", "Black") %>%
    str_replace("^Mixed or multiple ethnic groups$", "Mixed or Multiple") %>%
    str_replace("^Mixed ethnicity$", "Mixed or Multiple") %>%
    str_replace("^Ethnicity not specified$", "Not specified") %>%
    str_replace("^Not specified$", "Not specified") %>%
    str_replace("^Other ethnic group$", "Other") %>%
    str_replace("^Other ethnicity$", "Other") %>%
    str_replace("^Asian or Asian British$", "Asian")
}

as_eth_factor <- function(x) {
  factor(
    short_eth(as.character(x)),
    levels = ETH_LEVELS
  )
}

PAD_WIDTH <- 22

pad_lab <- function(x) {
  str_pad(
    as.character(x),
    width = PAD_WIDTH,
    side = "right"
  )
}

label_plot <- function(
  txt,
  size = 14,
  face = "bold",
  hjust = 0.5,
  vjust = 0.5,
  angle = 0,
  x = 0,
  y = 0,
  mar = margin(0, 0, 0, 0, unit = "pt")
) {

  ggplot() +
    annotate(
      "text",
      x = x,
      y = y,
      label = txt,
      fontface = face,
      hjust = hjust,
      vjust = vjust,
      angle = angle,
      size = size / ggplot2::.pt
    ) +
    theme_void() +
    theme(
      plot.margin = mar,
      panel.background = element_rect(
        fill = "transparent",
        colour = NA
      ),
      plot.background = element_rect(
        fill = "transparent",
        colour = NA
      )
    ) +
    xlim(-1, 1) +
    ylim(-1, 1) +
    coord_cartesian(clip = "off")
}

blank_plot <- function() {
  ggplot() +
    theme_void() +
    theme(
      plot.margin = margin(0, 0, 0, 0, unit = "pt")
    )
}

spacer_plot <- function() {
  ggplot() +
    theme_void() +
    theme(
      plot.margin = margin(0, 0, 0, 0, unit = "pt")
    )
}


############################
# Tile plot
############################

tile_plot <- function(
  df,
  xvar,
  yvar,
  show_x = FALSE,
  show_y = FALSE,
  fill_limits = NULL
) {

  x_top_text <- if (show_x) {

    element_text(
      angle = 45,
      hjust = 0,
      vjust = 0,
      size = 12,
      margin = margin(b = 2, unit = "pt")
    )

  } else {

    element_blank()
  }

  y_text <- if (show_y) {

    element_text(
      size = 12,
      margin = margin(r = 2, unit = "pt")
    )

  } else {

    element_blank()
  }

  ggplot(
    df,
    aes(
      x = {{ xvar }},
      y = {{ yvar }},
      fill = ir1000_scaled
    )
  ) +
    geom_tile(color = NA) +
    geom_text(
      aes(
        label = sprintf(
          "%.2f",
          ir1000_scaled
        )
      ),
      size = 3.7,
      na.rm = TRUE
    ) +
    scale_x_discrete(
      position = "top",
      expand = c(0, 0)
    ) +
    scale_y_discrete(
      expand = c(0, 0)
    ) +
    scale_fill_gradientn(
      colours = tailwind_cols,
      limits = fill_limits,
      oob = scales::squish,
      name = "Scaled IR per 1,000 PY"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      panel.grid = element_blank(),
      axis.title = element_blank(),
      axis.ticks = element_blank(),

      axis.text.x.top = x_top_text,
      axis.text.x.bottom = element_blank(),
      axis.text.y = y_text,

      legend.position = "none",

      panel.background = element_rect(
        fill = "transparent",
        colour = NA
      ),

      plot.background = element_rect(
        fill = "transparent",
        colour = NA
      ),

      plot.margin = margin(
        0, 0, 0, 0,
        unit = "pt"
      ),

      panel.spacing = unit(
        0,
        "pt"
      )
    )
}


align_heatmap_column <- function(
  heatmaps,
  rel_heights
) {

  aligned <- cowplot::align_plots(
    plotlist = heatmaps,
    align = "v",
    axis = "lr"
  )

  ROW_GAP <- 0.10

  plotlist2 <- list()
  rel2 <- c()

  for (i in seq_along(aligned)) {

    plotlist2 <- c(
      plotlist2,
      list(aligned[[i]])
    )

    rel2 <- c(
      rel2,
      rel_heights[i]
    )

    if (i < length(aligned)) {

      plotlist2 <- c(
        plotlist2,
        list(spacer_plot())
      )

      rel2 <- c(
        rel2,
        ROW_GAP
      )
    }
  }

  cowplot::plot_grid(
    plotlist = plotlist2,
    ncol = 1,
    rel_heights = rel2,
    align = "v",
    axis = "lr",
    greedy = TRUE
  )
}


make_column <- function(
  header_plot,
  heatmaps,
  rel_heights,
  header_height = 0.45
) {

  hm_col <- align_heatmap_column(
    heatmaps,
    rel_heights
  )

  cowplot::plot_grid(
    header_plot,
    hm_col,
    ncol = 1,
    rel_heights = c(
      header_height,
      sum(rel_heights)
    ),
    align = "v",
    axis = "lr",
    greedy = TRUE
  )
}


equalise_axis_top_height <- function(plots) {

  grobs <- lapply(
    plots,
    ggplotGrob
  )

  axis_rows_list <- lapply(
    grobs,
    function(g) {
      unique(
        g$layout$t[
          g$layout$name %in% c(
            "axis-t",
            "xlab-t"
          )
        ]
      )
    }
  )

  axis_rows <- axis_rows_list[[1]]

  max_heights <- grobs[[1]]$heights[
    axis_rows
  ]

  if (length(grobs) > 1) {

    for (i in 2:length(grobs)) {

      max_heights <- grid::unit.pmax(
        max_heights,
        grobs[[i]]$heights[
          axis_rows
        ]
      )
    }
  }

  grobs <- lapply(
    grobs,
    function(g) {

      g$heights[
        axis_rows
      ] <- max_heights

      g
    }
  )

  grobs
}


############################
# 1) Prepare SCALED tables
############################


#----------------------------
# DCAT × IMD
#----------------------------

dcat_imd_plot <- dcat_imd_scaled %>%
  mutate(
    dcat = factor(
      `Dental problem categories`,
      levels = sort(
        unique(
          `Dental problem categories`
        )
      )
    ),

    IMD_raw = as.character(imd),

    IMD = case_when(
      IMD_raw == "1" ~ "1 (Least Deprived)",
      IMD_raw == "2" ~ "2",
      IMD_raw == "3" ~ "3",
      IMD_raw == "4" ~ "4",
      IMD_raw == "5" ~ "5 (Most Deprived)",
      TRUE ~ "Unknown"
    ),

    IMD = factor(
      IMD,
      levels = c(
        "1 (Least Deprived)",
        "2",
        "3",
        "4",
        "5 (Most Deprived)",
        "Unknown"
      )
    ),

    IMD_lab = factor(
      pad_lab(IMD),
      levels = pad_lab(
        levels(IMD)
      )
    )
  ) %>%
  filter(IMD != "Unknown")


#----------------------------
# DCAT × Ethnicity
#----------------------------

dcat_eth_plot <- dcat_eth_scaled %>%
  mutate(
    dcat = factor(
      `Dental problem categories`,
      levels = sort(
        unique(
          `Dental problem categories`
        )
      )
    ),

    Ethnicity = as_eth_factor(
      Ethnicity
    ),

    Eth_lab = factor(
      pad_lab(Ethnicity),
      levels = pad_lab(
        levels(Ethnicity)
      )
    )
  ) %>%
  filter(
    Ethnicity != "Not specified"
  )


#----------------------------
# DCAT × Region
#----------------------------

dcat_region_plot <- dcat_region_scaled %>%
  mutate(
    dcat = factor(
      `Dental problem categories`,
      levels = sort(
        unique(
          `Dental problem categories`
        )
      )
    ),

    Region = factor(
      Region,
      levels = sort(
        unique(
          as.character(Region)
        )
      )
    ),

    Rgn_lab = factor(
      pad_lab(Region),
      levels = pad_lab(
        levels(Region)
      )
    )
  )


#----------------------------
# DCAT × Sex
#----------------------------

dcat_sex_plot <- dcat_sex_scaled %>%
  mutate(
    dcat = factor(
      `Dental problem categories`,
      levels = sort(
        unique(
          `Dental problem categories`
        )
      )
    ),

    Sex = factor(
      sex,
      levels = sort(
        unique(
          as.character(sex)
        )
      )
    ),

    Sex_lab = factor(
      pad_lab(Sex),
      levels = pad_lab(
        levels(Sex)
      )
    )
  )


#----------------------------
# DCAT × Age
#----------------------------

dcat_age2_plot <- dcat_age2_scaled %>%
  mutate(
    dcat = factor(
      `Dental problem categories`,
      levels = sort(
        unique(
          `Dental problem categories`
        )
      )
    ),

    Age2_raw = as.character(
      `Age group`
    ),

    Age2 = factor(
      Age2_raw,
      levels = c(
        "0-4",
        "5-9",
        "10-14",
        "15-19",
        "20-24",
        "25-29",
        "30-34",
        "35-39",
        "40-44",
        "45-49",
        "50-54",
        "55-59",
        "60-64",
        "65-69",
        "70-74",
        "75-79",
        "80-84",
        "85-89",
        "90+"
      )
    ),

    Age2_lab = factor(
      pad_lab(Age2),
      levels = pad_lab(
        levels(Age2)
      )
    )
  )


############################
# Non-DCAT tables
############################


#----------------------------
# Ethnicity × IMD
#----------------------------

eth_imd_plot <- eth_imd_scaled %>%
  mutate(
    Ethnicity = as_eth_factor(
      Ethnicity
    ),

    IMD = factor(
      as.character(IMD),
      levels = sort(
        unique(
          as.character(IMD)
        )
      )
    )
  ) %>%
  filter(
    Ethnicity != "Not specified"
  ) %>%
  filter(
    IMD != "Unknown"
  )


#----------------------------
# Ethnicity × Region
#----------------------------

eth_region_plot <- eth_region_scaled %>%
  mutate(
    Ethnicity = as_eth_factor(
      Ethnicity
    ),

    Region = factor(
      Region,
      levels = sort(
        unique(
          as.character(Region)
        )
      )
    )
  ) %>%
  filter(
    Ethnicity != "Not specified"
  )


#----------------------------
# Ethnicity × Sex
#----------------------------

eth_sex_plot <- eth_sex_scaled %>%
  mutate(
    Ethnicity = as_eth_factor(
      Ethnicity
    ),

    Sex = factor(
      Sex,
      levels = sort(
        unique(
          as.character(Sex)
        )
      )
    )
  ) %>%
  filter(
    Ethnicity != "Not specified"
  )


#----------------------------
# IMD × Region
#----------------------------

imd_region_plot <- imd_region_scaled %>%
  mutate(
    IMD = factor(
      as.character(IMD),
      levels = sort(
        unique(
          as.character(IMD)
        )
      )
    ),

    Region = factor(
      Region,
      levels = sort(
        unique(
          as.character(Region)
        )
      )
    )
  ) %>%
  filter(
    IMD != "Unknown"
  )


#----------------------------
# IMD × Sex
#----------------------------

imd_sex_plot <- imd_sex_scaled %>%
  mutate(
    IMD = factor(
      as.character(IMD),
      levels = sort(
        unique(
          as.character(IMD)
        )
      )
    ),

    Sex = factor(
      Sex,
      levels = sort(
        unique(
          as.character(Sex)
        )
      )
    )
  ) %>%
  filter(
    IMD != "Unknown"
  )


#----------------------------
# Region × Sex
#----------------------------

region_sex_plot <- region_sex_scaled %>%
  mutate(
    Region = factor(
      Region,
      levels = sort(
        unique(
          as.character(Region)
        )
      )
    ),

    Sex = factor(
      Sex,
      levels = sort(
        unique(
          as.character(Sex)
        )
      )
    )
  )


############################
# Age combinations
############################


#----------------------------
# Age × Sex
#----------------------------

age2_sex_plot <- age2_sex_scaled %>%
  mutate(
    Age2 = factor(
      as.character(age_band2),
      levels = c(
        "0-4",
        "5-9",
        "10-14",
        "15-19",
        "20-24",
        "25-29",
        "30-34",
        "35-39",
        "40-44",
        "45-49",
        "50-54",
        "55-59",
        "60-64",
        "65-69",
        "70-74",
        "75-79",
        "80-84",
        "85-89",
        "90+"
      )
    ),

    Sex = factor(
      gender,
      levels = sort(
        unique(
          as.character(gender)
        )
      )
    )
  )


#----------------------------
# Age × Region
#----------------------------

age2_rgn_plot <- age2_rgn_scaled %>%
  mutate(
    Age2 = factor(
      as.character(age_band2),
      levels = c(
        "0-4",
        "5-9",
        "10-14",
        "15-19",
        "20-24",
        "25-29",
        "30-34",
        "35-39",
        "40-44",
        "45-49",
        "50-54",
        "55-59",
        "60-64",
        "65-69",
        "70-74",
        "75-79",
        "80-84",
        "85-89",
        "90+"
      )
    ),

    Region = factor(
      Region,
      levels = sort(
        unique(
          as.character(Region)
        )
      )
    )
  )


#----------------------------
# Age × Ethnicity
#----------------------------

age2_eth_plot <- age2_eth_scaled %>%
  mutate(
    Age2 = factor(
      as.character(age_band2),
      levels = c(
        "0-4",
        "5-9",
        "10-14",
        "15-19",
        "20-24",
        "25-29",
        "30-34",
        "35-39",
        "40-44",
        "45-49",
        "50-54",
        "55-59",
        "60-64",
        "65-69",
        "70-74",
        "75-79",
        "80-84",
        "85-89",
        "90+"
      )
    ),

    Ethnicity = as_eth_factor(
      Ethnicity
    )
  ) %>%
  filter(
    Ethnicity != "Not specified"
  )


#----------------------------
# Age × IMD
#----------------------------

age2_imd_plot <- age2_imd_scaled %>%
  mutate(
    Age2 = factor(
      as.character(age_band2),
      levels = c(
        "0-4",
        "5-9",
        "10-14",
        "15-19",
        "20-24",
        "25-29",
        "30-34",
        "35-39",
        "40-44",
        "45-49",
        "50-54",
        "55-59",
        "60-64",
        "65-69",
        "70-74",
        "75-79",
        "80-84",
        "85-89",
        "90+"
      )
    ),

    IMD = factor(
      IMD,
      levels = c(
        "1 (Least Deprived)",
        "2",
        "3",
        "4",
        "5 (Most Deprived)",
        "Unknown"
      )
    )
  ) %>%
  filter(
    IMD != "Unknown"
  )


############################
# 2) Shared colour limits
#
# IMPORTANT:
# use ir1000_scaled everywhere
############################

fill_limits <- range(
  dcat_sex_plot$ir1000_scaled,
  dcat_region_plot$ir1000_scaled,
  dcat_eth_plot$ir1000_scaled,
  dcat_imd_plot$ir1000_scaled,
  dcat_age2_plot$ir1000_scaled,

  imd_sex_plot$ir1000_scaled,
  imd_region_plot$ir1000_scaled,

  eth_sex_plot$ir1000_scaled,
  eth_region_plot$ir1000_scaled,
  eth_imd_plot$ir1000_scaled,

  region_sex_plot$ir1000_scaled,

  age2_sex_plot$ir1000_scaled,
  age2_rgn_plot$ir1000_scaled,
  age2_eth_plot$ir1000_scaled,
  age2_imd_plot$ir1000_scaled,

  na.rm = TRUE
)


############################
# 3) Build triangle heatmaps
############################


#----------------------------
# DCAT column
#----------------------------

p_dcat_age2 <- tile_plot(
  dcat_age2_plot,
  dcat,
  Age2_lab,
  show_x = TRUE,
  show_y = TRUE,
  fill_limits = fill_limits
)

p_dcat_rgn <- tile_plot(
  dcat_region_plot,
  dcat,
  Rgn_lab,
  show_x = FALSE,
  show_y = TRUE,
  fill_limits = fill_limits
)

p_dcat_eth <- tile_plot(
  dcat_eth_plot,
  dcat,
  Eth_lab,
  show_x = FALSE,
  show_y = TRUE,
  fill_limits = fill_limits
)

p_dcat_imd <- tile_plot(
  dcat_imd_plot,
  dcat,
  IMD_lab,
  show_x = FALSE,
  show_y = TRUE,
  fill_limits = fill_limits
)

p_dcat_sex <- tile_plot(
  dcat_sex_plot,
  dcat,
  Sex_lab,
  show_x = FALSE,
  show_y = TRUE,
  fill_limits = fill_limits
)


#----------------------------
# Sex column
#----------------------------

p_sex_age2 <- tile_plot(
  age2_sex_plot,
  Sex,
  Age2,
  show_x = TRUE,
  show_y = FALSE,
  fill_limits = fill_limits
)

p_sex_rgn <- tile_plot(
  region_sex_plot,
  Sex,
  Region,
  show_x = FALSE,
  show_y = FALSE,
  fill_limits = fill_limits
)

p_sex_eth <- tile_plot(
  eth_sex_plot,
  Sex,
  Ethnicity,
  show_x = FALSE,
  show_y = FALSE,
  fill_limits = fill_limits
)

p_sex_imd <- tile_plot(
  imd_sex_plot,
  Sex,
  IMD,
  show_x = FALSE,
  show_y = FALSE,
  fill_limits = fill_limits
)

p_sex_sex <- blank_plot()


#----------------------------
# IMD column
#----------------------------

p_imd_age2 <- tile_plot(
  age2_imd_plot,
  IMD,
  Age2,
  show_x = TRUE,
  show_y = FALSE,
  fill_limits = fill_limits
)

p_imd_rgn <- tile_plot(
  imd_region_plot,
  IMD,
  Region,
  show_x = FALSE,
  show_y = FALSE,
  fill_limits = fill_limits
)

p_imd_eth <- tile_plot(
  eth_imd_plot,
  IMD,
  Ethnicity,
  show_x = FALSE,
  show_y = FALSE,
  fill_limits = fill_limits
)

p_imd_imd <- blank_plot()
p_imd_sex <- blank_plot()


#----------------------------
# Ethnicity column
#----------------------------

p_eth_age2 <- tile_plot(
  age2_eth_plot,
  Ethnicity,
  Age2,
  show_x = TRUE,
  show_y = FALSE,
  fill_limits = fill_limits
)

p_eth_rgn <- tile_plot(
  eth_region_plot,
  Ethnicity,
  Region,
  show_x = FALSE,
  show_y = FALSE,
  fill_limits = fill_limits
)

p_eth_eth <- blank_plot()
p_eth_imd <- blank_plot()
p_eth_sex <- blank_plot()


#----------------------------
# Region column
#----------------------------

p_rgn_age2 <- tile_plot(
  age2_rgn_plot,
  Region,
  Age2,
  show_x = TRUE,
  show_y = FALSE,
  fill_limits = fill_limits
)

p_rgn_rgn <- blank_plot()
p_rgn_eth <- blank_plot()
p_rgn_imd <- blank_plot()
p_rgn_sex <- blank_plot()


############################
# Equalise top axes
############################

top_row_fixed <- equalise_axis_top_height(
  list(
    p_dcat_age2,
    p_sex_age2,
    p_imd_age2,
    p_eth_age2,
    p_rgn_age2
  )
)

p_dcat_age2 <- top_row_fixed[[1]]
p_sex_age2  <- top_row_fixed[[2]]
p_imd_age2  <- top_row_fixed[[3]]
p_eth_age2  <- top_row_fixed[[4]]
p_rgn_age2  <- top_row_fixed[[5]]


############################
# 4) Row heights
############################

heatmap_row_heights <- c(
  Age2 = 10.5,
  Region = 4,
  Ethnicity = 2.6,
  IMD = 2.6,
  Sex = 0.9
)


############################
# 5) Headers + row labels
############################

h_row <- label_plot("")

h_dcat <- label_plot(
  "Dental problem",
  hjust = 0,
  mar = margin(
    0, 0, 0, 25,
    unit = "pt"
  )
)

h_sex <- label_plot("Sex")
h_imd <- label_plot("IMD")
h_eth <- label_plot("Ethnicity")
h_rgn <- label_plot("English region")


r_age2 <- label_plot(
  "Age",
  face = "bold",
  angle = 90,
  x = 0.98,
  hjust = 1
)

r_rgn <- label_plot(
  "English region",
  face = "bold",
  angle = 90,
  x = 0.98,
  y = 0.6,
  hjust = 1
)

r_eth <- label_plot(
  "Ethnicity",
  face = "bold",
  angle = 90,
  x = 0.98,
  y = 0.6,
  hjust = 1
)

r_imd <- label_plot(
  "IMD",
  face = "bold",
  angle = 90,
  x = 0.98,
  y = 0.3,
  hjust = 1
)

r_sex <- label_plot(
  "Sex",
  face = "bold",
  angle = 90,
  x = 0.98,
  y = 0.5,
  hjust = 1
)


col_rows <- cowplot::plot_grid(
  h_row,

  cowplot::plot_grid(
    r_age2,
    r_rgn,
    r_eth,
    r_imd,
    r_sex,
    ncol = 1,
    rel_heights = heatmap_row_heights
  ),

  ncol = 1,
  rel_heights = c(
    0.45,
    sum(heatmap_row_heights)
  ),

  align = "v",
  axis = "lr",
  greedy = TRUE
)


############################
# 6) Build columns
############################

col_dcat <- make_column(
  header_plot = h_dcat,

  heatmaps = list(
    p_dcat_age2,
    p_dcat_rgn,
    p_dcat_eth,
    p_dcat_imd,
    p_dcat_sex
  ),

  rel_heights = heatmap_row_heights,
  header_height = 0.45
)


col_sex <- make_column(
  header_plot = h_sex,

  heatmaps = list(
    p_sex_age2,
    p_sex_rgn,
    p_sex_eth,
    p_sex_imd,
    p_sex_sex
  ),

  rel_heights = heatmap_row_heights,
  header_height = 0.45
)


col_imd <- make_column(
  header_plot = h_imd,

  heatmaps = list(
    p_imd_age2,
    p_imd_rgn,
    p_imd_eth,
    p_imd_imd,
    p_imd_sex
  ),

  rel_heights = heatmap_row_heights,
  header_height = 0.45
)


col_eth <- make_column(
  header_plot = h_eth,

  heatmaps = list(
    p_eth_age2,
    p_eth_rgn,
    p_eth_eth,
    p_eth_imd,
    p_eth_sex
  ),

  rel_heights = heatmap_row_heights,
  header_height = 0.45
)


col_rgn <- make_column(
  header_plot = h_rgn,

  heatmaps = list(
    p_rgn_age2,
    p_rgn_rgn,
    p_rgn_eth,
    p_rgn_imd,
    p_rgn_sex
  ),

  rel_heights = heatmap_row_heights,
  header_height = 0.45
)


############################
# 7) Shared legend
#
# IMPORTANT:
# use ir1000_scaled
############################

legend_plot <- ggplot(
  dcat_sex_plot,
  aes(
    x = dcat,
    y = Sex_lab,
    fill = ir1000_scaled
  )
) +
  geom_tile() +
  scale_fill_gradientn(
    colours = tailwind_cols,
    limits = fill_limits,
    oob = scales::squish,
    name = "Scaled IR (per 1000 PY)"
  ) +
  theme_void() +
  theme(
    legend.position = "right",

    legend.title = element_text(
      face = "bold",
      size = 13
    ),

    legend.text = element_text(
      size = 11
    ),

    legend.margin = margin(
      12, 20, 12, 20,
      unit = "pt"
    ),

    legend.box.margin = margin(
      12, 20, 12, 20,
      unit = "pt"
    ),

    legend.key.height = unit(
      22,
      "pt"
    ),

    legend.key.width = unit(
      18,
      "pt"
    ),

    plot.margin = margin(
      26, 55, 26, 22,
      unit = "pt"
    )
  )


legend_only <- cowplot::get_legend(
  legend_plot
)


legend_box <- cowplot::plot_grid(
  spacer_plot(),

  cowplot::ggdraw(
    legend_only
  ),

  ncol = 1,
  rel_heights = c(
    0.35,
    1
  )
)


############################
# 8) Combine
############################

COL_GAP <- 0.005

final_scaled <- cowplot::plot_grid(
  col_rows,

  spacer_plot(),
  col_dcat,

  spacer_plot(),
  col_sex,

  spacer_plot(),
  col_imd,

  spacer_plot(),
  col_eth,

  spacer_plot(),
  col_rgn,

  spacer_plot(),
  legend_box,

  nrow = 1,

  rel_widths = c(
    0.10,
    COL_GAP,
    0.75,
    COL_GAP,
    0.15,
    0.01,
    0.40,
    COL_GAP,
    0.40,
    COL_GAP,
    0.6,
    COL_GAP,
    0.20
  ),

  align = "h",
  axis = "tb",
  greedy = TRUE
)


final_scaled


############################
# 9) Save scaled heatmap
############################

ggsave(
  filename = paste0(
    tables_figures,
    "/formatted/Figure 2. Heatmap of incidence rate by sociodemographic factors.jpeg"
  ),

  plot = final_scaled,

  width = 21,
  height = 14,
  units = "in",
  dpi = 1200
)

ggsave(
  filename = paste0(
    tables_figures,
    "/formatted/Figure 2. Heatmap of incidence rate by sociodemographic factors.pdf"
  ),

  plot = final_scaled,

  width = 21,
  height = 14,
  units = "in",
  dpi = 1200
)
```


## format heatmap table
```{r}
#---------------------------------------------------------
# 1. Scaled dataframes
#---------------------------------------------------------

loaded_objs <- c(
  "dcat_imd_scaled", "dcat_eth_scaled", "dcat_region_scaled", "dcat_sex_scaled",
  "eth_imd_scaled", "eth_region_scaled", "eth_sex_scaled",
  "imd_region_scaled", "imd_sex_scaled", "region_sex_scaled",
  "dcat_age2_scaled", "age2_sex_scaled", "age2_rgn_scaled",
  "age2_eth_scaled", "age2_imd_scaled"
)


#---------------------------------------------------------
# 2. Variable mapping
#---------------------------------------------------------

var_map <- list(

  dcat_imd_scaled =
    c(var1 = "Dental problem categories", var2 = "IMD"),

  dcat_eth_scaled =
    c(var1 = "Dental problem categories", var2 = "Ethnicity"),

  dcat_region_scaled =
    c(var1 = "Dental problem categories", var2 = "Region"),

  dcat_sex_scaled =
    c(var1 = "Dental problem categories", var2 = "Sex"),

  eth_imd_scaled =
    c(var1 = "Ethnicity", var2 = "IMD"),

  eth_region_scaled =
    c(var1 = "Ethnicity", var2 = "Region"),

  eth_sex_scaled =
    c(var1 = "Ethnicity", var2 = "Sex"),

  imd_region_scaled =
    c(var1 = "IMD", var2 = "Region"),

  imd_sex_scaled =
    c(var1 = "IMD", var2 = "Sex"),

  region_sex_scaled =
    c(var1 = "Region", var2 = "Sex"),

  dcat_age2_scaled =
    c(var1 = "Dental problem categories", var2 = "Age2"),

  # age2_sex_scaled uses the original variable names
  age2_sex_scaled =
    c(var1 = "age_band2", var2 = "gender"),

  age2_rgn_scaled =
    c(var1 = "Age2", var2 = "Region"),

  age2_eth_scaled =
    c(var1 = "Age2", var2 = "Ethnicity"),

  age2_imd_scaled =
    c(var1 = "Age2", var2 = "IMD")
)


#---------------------------------------------------------
# 3. Prefix levels
#---------------------------------------------------------

prefix_level <- function(var, level) {

  dplyr::case_when(

    var == "Age2" ~
      paste("Age", level),

    var == "age_band2" ~
      paste("Age", level),

    var == "Sex" ~
      paste("Sex", level),

    var == "gender" ~
      paste("Sex", level),

    var == "Ethnicity" ~
      paste("Ethnicity", level),

    var == "IMD" ~
      paste("IMD", level),

    var == "Region" ~
      paste("English region", level),

    var == "Dental problem categories" ~
      paste("Dental problem", level),

    TRUE ~
      paste(var, level)
  )
}


#---------------------------------------------------------
# 4. Standardise scaled IR tables
#---------------------------------------------------------

standardise_ir_table <- function(df, df_name) {

  this_map <- var_map[[df_name]]

  df %>%
    dplyr::mutate(

      variable1 = unname(this_map["var1"]),
      variable2 = unname(this_map["var2"]),

      level1_raw = .data[[unname(this_map["var1"])]],
      level2_raw = .data[[unname(this_map["var2"])]],

      level1 = prefix_level(
        variable1,
        as.character(level1_raw)
      ),

      level2 = prefix_level(
        variable2,
        as.character(level2_raw)
      ),

      `IR per 1000 PY (95%CI)` = sprintf(
        "%.2f (%.2f-%.2f)",
        ir1000_scaled,
        lower95_scaled,
        upper95_scaled
      )
    ) %>%

    dplyr::select(
      variable1,
      level1,
      variable2,
      level2,
      cases,
      total_person_year,
      ir1000_scaled,
      lower95_scaled,
      upper95_scaled,
      `IR per 1000 PY (95%CI)`
    )
}


#---------------------------------------------------------
# 5. Combine all scaled tables
#---------------------------------------------------------

combined_ir_tables_scaled <- purrr::map_dfr(
  loaded_objs,
  ~ standardise_ir_table(get(.x), .x),
  .id = "source_df"
)


#---------------------------------------------------------
# 6. View result
#---------------------------------------------------------

View(combined_ir_tables_scaled)



```

## fromat table
```{r}
combined_ir_tables_scaled_formatted <- combined_ir_tables_scaled %>%
  mutate(
    total_person_year = as.integer(round(total_person_year))
  ) %>%
  rename(
    `Variable 1` = variable1,
    `Level 1` = level1,
    `Variable 2` = variable2,
    `Level 2` = level2,
    `N events` = cases,
    `Total person years` = total_person_year
  ) %>%
  dplyr::select(
    `Level 1`,
    `Level 2`,
    `N events`,
    `Total person years`,
    `IR per 1000 PY (95%CI)`
  )%>%
  mutate(
  `N events` = formatC(`N events`, format = "f", digits = 0, big.mark = ","),
`Total person years` = formatC(
  `Total person years`,format = "f", digits = 0,big.mark = ","),
    `IR per 1000 PY (95%CI)` = gsub("\\.","·",`IR per 1000 PY (95%CI)`)
)


write_xlsx(
  combined_ir_tables_scaled_formatted,
  paste0(
    tables_figures,
    "/formatted/Table S8. Intersectional IR and confidence interval in heatmap.xlsx"
  )
)
```

